---
title: "AMSTROM Cohort Analysis"
author: "Morgan Virolan"
date: "2026-04-20"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: cosmo
---

# Overview

This notebook pools the per-run downstream R-derived parquet tables stored in:

- `results/R_ANALYSIS/<RUN_ID>/`

and writes cohort-level pooled outputs to:

- `results/R_ANALYSIS/cohort/`

The notebook preserves the existing downstream contracts and does not change the
underlying statistical logic. It only stacks matching run-level derived tables,
adds run metadata where needed, and exports the pooled cohort tables.

# 1. Setup

## 1.1 Load packages

```{r setup_load_packages, message=FALSE, warning=FALSE}
library(arrow)
library(dplyr)
library(ggplot2)
library(tidyr)

options(dplyr.summarise.inform = FALSE)
```

# 2. Discover available runs

## 2.1 Define analysis paths

```{r discover_define_paths}
analysis_root <- file.path("results", "R_ANALYSIS")
cohort_output_root <- file.path(analysis_root, "cohort")
```

## 2.2 List run folders

Only per-run folders are kept. The existing cohort folder is excluded from pooling.

```{r discover_list_run_folders}
run_ids <- list.dirs(analysis_root, recursive = FALSE, full.names = FALSE)
run_ids <- setdiff(run_ids, "cohort")

run_ids
```

# 3. Pool run-level derived tables

## 3.1 Run metadata

```{r pool_run_metadata}
cohort_run_metadata_table <- dplyr::bind_rows(
  lapply(run_ids, function(run_id) {
    read_parquet(file.path(analysis_root, run_id, "run_metadata.parquet"))
  })
) |>
  dplyr::arrange(RUN_ID)

cohort_run_metadata_table
```

## 3.2 Reference-level feature table

```{r pool_reference_level_feature_table_norm}
cohort_reference_level_feature_table_norm <- dplyr::bind_rows(
  lapply(run_ids, function(run_id) {
    read_parquet(file.path(analysis_root, run_id, "reference_level_feature_table_norm.parquet"))
  })
) |>
  dplyr::arrange(RUN_ID, channel, reference_level_id)
```

## 3.3 EX_DYN middle-1s feature table

```{r pool_ex_dyn_mid1s_feature_table_norm}
cohort_ex_dyn_mid1s_feature_table_norm <- dplyr::bind_rows(
  lapply(run_ids, function(run_id) {
    read_parquet(file.path(analysis_root, run_id, "ex_dyn_mid1s_feature_table_norm.parquet"))
  })
) |>
  dplyr::arrange(RUN_ID, channel, VC_count)
```

## 3.4 EX_STA middle-1s feature table

```{r pool_ex_sta_mid1s_feature_table_norm}
cohort_ex_sta_mid1s_feature_table_norm <- dplyr::bind_rows(
  lapply(run_ids, function(run_id) {
    read_parquet(file.path(analysis_root, run_id, "ex_sta_mid1s_feature_table_norm.parquet"))
  })
) |>
  dplyr::arrange(RUN_ID, channel, window_id)
```

## 3.5 EX_DYN EMG-to-torque ratio table

```{r pool_ex_dyn_emg_torque_ratio_table}
cohort_ex_dyn_emg_torque_ratio_table <- dplyr::bind_rows(
  lapply(run_ids, function(run_id) {
    read_parquet(file.path(analysis_root, run_id, "ex_dyn_emg_torque_ratio_table.parquet"))
  })
) |>
  dplyr::arrange(RUN_ID, channel, VC_count)
```

## 3.6 EX_STA EMG-to-torque ratio table

```{r pool_ex_sta_emg_torque_ratio_table}
cohort_ex_sta_emg_torque_ratio_table <- dplyr::bind_rows(
  lapply(run_ids, function(run_id) {
    read_parquet(file.path(analysis_root, run_id, "ex_sta_emg_torque_ratio_table.parquet"))
  })
) |>
  dplyr::arrange(RUN_ID, channel, window_id)
```

## 3.7 EMG-to-torque ratio slope summary table

```{r pool_emg_torque_ratio_slope_summary_table}
cohort_emg_torque_ratio_slope_summary_table <- dplyr::bind_rows(
  lapply(run_ids, function(run_id) {
    read_parquet(file.path(analysis_root, run_id, "emg_torque_ratio_slope_summary_table.parquet"))
  })
) |>
  dplyr::arrange(SEQ, channel, RUN_ID)
```

## 3.8 Reference calibration table in normalized space

```{r pool_reference_calibration_table_norm}
cohort_reference_calibration_table_norm <- dplyr::bind_rows(
  lapply(run_ids, function(run_id) {
    read_parquet(file.path(analysis_root, run_id, "reference_calibration_table_norm.parquet"))
  })
) |>
  dplyr::arrange(channel, RUN_ID)
```

# 4. Add metadata-enriched cohort tables

These tables keep the original pooled outputs and also create metadata-enriched
versions for cohort-facing analyses.

## 4.1 EMG-to-torque ratio slope summary with metadata

```{r enrich_emg_torque_ratio_slope_summary_with_metadata}
cohort_emg_torque_ratio_slope_summary_with_metadata <- cohort_emg_torque_ratio_slope_summary_table |>
  dplyr::left_join(
    cohort_run_metadata_table,
    by = c("RUN_ID", "participant_id")
  ) |>
  dplyr::arrange(SEQ, channel, RUN_ID)
```

## 4.2 Reference calibration table with metadata

```{r enrich_reference_calibration_table_norm_with_metadata}
cohort_reference_calibration_table_norm_with_metadata <- cohort_reference_calibration_table_norm |>
  dplyr::left_join(
    cohort_run_metadata_table,
    by = c("RUN_ID", "participant_id")
  ) |>
  dplyr::arrange(channel, RUN_ID)
```

# 5. Write cohort outputs

## 5.1 Export control

Set this to `TRUE` only when you want to refresh the cohort parquet exports on disk.
For routine notebook work, keeping it `FALSE` avoids unnecessary writes and rereads.

```{r write_export_control}
EXPORT_COHORT_PARQUET <- FALSE
```

## 5.2 Create output folder when export is enabled

```{r write_create_output_folder}
if (isTRUE(EXPORT_COHORT_PARQUET)) {
  dir.create(cohort_output_root, recursive = TRUE, showWarnings = FALSE)
}
```

## 5.3 Export pooled cohort parquet tables

```{r write_export_cohort_parquet_tables}
if (isTRUE(EXPORT_COHORT_PARQUET)) {
  write_parquet(cohort_run_metadata_table,
                sink = file.path(cohort_output_root, "cohort_run_metadata.parquet"))
  write_parquet(cohort_reference_level_feature_table_norm,
                sink = file.path(cohort_output_root, "cohort_reference_level_feature_table_norm.parquet"))
  write_parquet(cohort_ex_dyn_mid1s_feature_table_norm,
                sink = file.path(cohort_output_root, "cohort_ex_dyn_mid1s_feature_table_norm.parquet"))
  write_parquet(cohort_ex_sta_mid1s_feature_table_norm,
                sink = file.path(cohort_output_root, "cohort_ex_sta_mid1s_feature_table_norm.parquet"))
  write_parquet(cohort_ex_dyn_emg_torque_ratio_table,
                sink = file.path(cohort_output_root, "cohort_ex_dyn_emg_torque_ratio_table.parquet"))
  write_parquet(cohort_ex_sta_emg_torque_ratio_table,
                sink = file.path(cohort_output_root, "cohort_ex_sta_emg_torque_ratio_table.parquet"))
  write_parquet(cohort_emg_torque_ratio_slope_summary_table,
                sink = file.path(cohort_output_root, "cohort_emg_torque_ratio_slope_summary_table.parquet"))
  write_parquet(cohort_reference_calibration_table_norm,
                sink = file.path(cohort_output_root, "cohort_reference_calibration_table_norm.parquet"))
  write_parquet(cohort_emg_torque_ratio_slope_summary_with_metadata,
                sink = file.path(cohort_output_root, "cohort_emg_torque_ratio_slope_summary_with_metadata.parquet"))
  write_parquet(cohort_reference_calibration_table_norm_with_metadata,
                sink = file.path(cohort_output_root, "cohort_reference_calibration_table_norm_with_metadata.parquet"))
}
```

# 6. Cohort analysis scaffold

This section is intentionally kept generic so the same notebook can support
future pooled analyses beyond the current EMG-derived tables. The idea is to
use the pooled cohort parquet outputs as the stable analysis inputs, then build
feature-specific analyses below.

## 6.1 Define cohort analysis inputs

```{r cohort_analysis_define_inputs}
cohort_analysis_input_root <- cohort_output_root

cohort_file_run_metadata <- file.path(cohort_analysis_input_root, "cohort_run_metadata.parquet")
cohort_file_reference_level_feature_table_norm <- file.path(cohort_analysis_input_root, "cohort_reference_level_feature_table_norm.parquet")
cohort_file_ex_dyn_mid1s_feature_table_norm <- file.path(cohort_analysis_input_root, "cohort_ex_dyn_mid1s_feature_table_norm.parquet")
cohort_file_ex_sta_mid1s_feature_table_norm <- file.path(cohort_analysis_input_root, "cohort_ex_sta_mid1s_feature_table_norm.parquet")
cohort_file_ex_dyn_emg_torque_ratio_table <- file.path(cohort_analysis_input_root, "cohort_ex_dyn_emg_torque_ratio_table.parquet")
cohort_file_ex_sta_emg_torque_ratio_table <- file.path(cohort_analysis_input_root, "cohort_ex_sta_emg_torque_ratio_table.parquet")
cohort_file_emg_torque_ratio_slope_summary_table <- file.path(cohort_analysis_input_root, "cohort_emg_torque_ratio_slope_summary_table.parquet")
cohort_file_reference_calibration_table_norm <- file.path(cohort_analysis_input_root, "cohort_reference_calibration_table_norm.parquet")
cohort_file_emg_torque_ratio_slope_summary_with_metadata <- file.path(cohort_analysis_input_root, "cohort_emg_torque_ratio_slope_summary_with_metadata.parquet")
cohort_file_reference_calibration_table_norm_with_metadata <- file.path(cohort_analysis_input_root, "cohort_reference_calibration_table_norm_with_metadata.parquet")
```

## 6.2 Read pooled cohort tables for downstream analysis

```{r cohort_analysis_read_inputs}
if (isTRUE(EXPORT_COHORT_PARQUET)) {
  cohort_run_metadata_analysis <- read_parquet(cohort_file_run_metadata)
  cohort_reference_level_feature_table_norm_analysis <- read_parquet(cohort_file_reference_level_feature_table_norm)
  cohort_ex_dyn_mid1s_feature_table_norm_analysis <- read_parquet(cohort_file_ex_dyn_mid1s_feature_table_norm)
  cohort_ex_sta_mid1s_feature_table_norm_analysis <- read_parquet(cohort_file_ex_sta_mid1s_feature_table_norm)
  cohort_ex_dyn_emg_torque_ratio_table_analysis <- read_parquet(cohort_file_ex_dyn_emg_torque_ratio_table)
  cohort_ex_sta_emg_torque_ratio_table_analysis <- read_parquet(cohort_file_ex_sta_emg_torque_ratio_table)
  cohort_emg_torque_ratio_slope_summary_table_analysis <- read_parquet(cohort_file_emg_torque_ratio_slope_summary_with_metadata)
  cohort_reference_calibration_table_norm_analysis <- read_parquet(cohort_file_reference_calibration_table_norm_with_metadata)
} else {
  cohort_run_metadata_analysis <- cohort_run_metadata_table
  cohort_reference_level_feature_table_norm_analysis <- cohort_reference_level_feature_table_norm
  cohort_ex_dyn_mid1s_feature_table_norm_analysis <- cohort_ex_dyn_mid1s_feature_table_norm
  cohort_ex_sta_mid1s_feature_table_norm_analysis <- cohort_ex_sta_mid1s_feature_table_norm
  cohort_ex_dyn_emg_torque_ratio_table_analysis <- cohort_ex_dyn_emg_torque_ratio_table
  cohort_ex_sta_emg_torque_ratio_table_analysis <- cohort_ex_sta_emg_torque_ratio_table
  cohort_emg_torque_ratio_slope_summary_table_analysis <- cohort_emg_torque_ratio_slope_summary_with_metadata
  cohort_reference_calibration_table_norm_analysis <- cohort_reference_calibration_table_norm_with_metadata
}
```

## 6.3 Quick cohort data check

Use this block to verify that the pooled cohort tables were created as expected
before running any actual pooled analysis.

```{r cohort_analysis_quick_data_check, eval=FALSE}
tibble::tibble(
  table_name = c(
    "cohort_run_metadata_analysis",
    "cohort_reference_level_feature_table_norm_analysis",
    "cohort_ex_dyn_mid1s_feature_table_norm_analysis",
    "cohort_ex_sta_mid1s_feature_table_norm_analysis",
    "cohort_ex_dyn_emg_torque_ratio_table_analysis",
    "cohort_ex_sta_emg_torque_ratio_table_analysis",
    "cohort_emg_torque_ratio_slope_summary_table_analysis",
    "cohort_reference_calibration_table_norm_analysis"
  ),
  n_rows = c(
    nrow(cohort_run_metadata_analysis),
    nrow(cohort_reference_level_feature_table_norm_analysis),
    nrow(cohort_ex_dyn_mid1s_feature_table_norm_analysis),
    nrow(cohort_ex_sta_mid1s_feature_table_norm_analysis),
    nrow(cohort_ex_dyn_emg_torque_ratio_table_analysis),
    nrow(cohort_ex_sta_emg_torque_ratio_table_analysis),
    nrow(cohort_emg_torque_ratio_slope_summary_table_analysis),
    nrow(cohort_reference_calibration_table_norm_analysis)
  )
)
```

## 6.4 Cohort EMG analysis checks

This section replaces the earlier placeholders with simple pooled EMG tables and
plots that use the cohort parquet outputs directly. The goal is not to change
any single-run logic, only to verify that the pooled workflow produces readable
cohort-level results.

### 6.4.1 Analysis selection

```{r cohort_analysis_feature_selection}
selected_analysis_family <- "EMG_ratio"
selected_sequences <- c("EX_DYN", "EX_STA")
selected_channels <- c("emg6", "emg8", "emg10")
SHOW_COHORT_TABLES <- FALSE
SHOW_COHORT_PLOTS <- TRUE

selected_analysis_family
selected_sequences
selected_channels
SHOW_COHORT_TABLES
SHOW_COHORT_PLOTS
```

### 6.4.2 Included runs and participants

```{r cohort_analysis_included_runs_table, eval=FALSE}
cohort_included_runs_table <- cohort_run_metadata_analysis |>
  dplyr::arrange(participant_id, RUN_ID) |>
  dplyr::select(
    RUN_ID,
    participant_id,
    sex,
    dominant_side,
    practice_level,
    height_cm,
    mass_kg,
    mvc_ref
  )

if (isTRUE(SHOW_COHORT_TABLES)) {
  print(cohort_included_runs_table)
}
```

### 6.4.3 Reference-level descriptive table

```{r cohort_analysis_reference_level_summary_table, eval=FALSE}
cohort_reference_level_summary_table <- cohort_reference_level_feature_table_norm_analysis |>
  dplyr::filter(channel %in% selected_channels) |>
  dplyr::group_by(channel, reference_level) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    n_runs = dplyr::n_distinct(RUN_ID),
    mean_torque_norm = mean(torque_norm, na.rm = TRUE),
    sd_torque_norm = sd(torque_norm, na.rm = TRUE),
    mean_rms_norm = mean(rms_norm, na.rm = TRUE),
    sd_rms_norm = sd(rms_norm, na.rm = TRUE),
    mean_mdf_hz = mean(mdf_mid_1s_hz, na.rm = TRUE),
    sd_mdf_hz = sd(mdf_mid_1s_hz, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel, reference_level)

if (isTRUE(SHOW_COHORT_TABLES)) {
  print(cohort_reference_level_summary_table)
}
```

### 6.4.4 EX_DYN ratio descriptive table

```{r cohort_analysis_ex_dyn_ratio_summary_table, eval=FALSE}
cohort_ex_dyn_ratio_summary_table <- cohort_ex_dyn_emg_torque_ratio_table_analysis |>
  dplyr::filter(
    SEQ %in% selected_sequences,
    channel %in% selected_channels
  ) |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    n_runs = dplyr::n_distinct(RUN_ID),
    mean_ratio = mean(emg_torque_ratio, na.rm = TRUE),
    sd_ratio = sd(emg_torque_ratio, na.rm = TRUE),
    mean_expected_ratio = mean(expected_ratio, na.rm = TRUE),
    mean_torque_norm = mean(torque_norm, na.rm = TRUE),
    mean_rms_norm = mean(rms_norm, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(SEQ, channel)

if (isTRUE(SHOW_COHORT_TABLES)) {
  print(cohort_ex_dyn_ratio_summary_table)
}
```

### 6.4.5 EX_STA ratio descriptive table

```{r cohort_analysis_ex_sta_ratio_summary_table, eval=FALSE}
cohort_ex_sta_ratio_summary_table <- cohort_ex_sta_emg_torque_ratio_table_analysis |>
  dplyr::filter(
    SEQ %in% selected_sequences,
    channel %in% selected_channels
  ) |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    n_runs = dplyr::n_distinct(RUN_ID),
    mean_ratio = mean(emg_torque_ratio, na.rm = TRUE),
    sd_ratio = sd(emg_torque_ratio, na.rm = TRUE),
    mean_expected_ratio = mean(expected_ratio, na.rm = TRUE),
    mean_torque_norm = mean(torque_norm, na.rm = TRUE),
    mean_rms_norm = mean(rms_norm, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(SEQ, channel)

if (isTRUE(SHOW_COHORT_TABLES)) {
  print(cohort_ex_sta_ratio_summary_table)
}
```

### 6.4.6 Ratio slope summary table

```{r cohort_analysis_ratio_slope_summary_table, eval=FALSE}
cohort_ratio_slope_summary_table <- cohort_emg_torque_ratio_slope_summary_table_analysis |>
  dplyr::filter(
    SEQ %in% selected_sequences,
    channel %in% selected_channels
  ) |>
  dplyr::arrange(SEQ, channel, participant_id, RUN_ID) |>
  dplyr::select(
    RUN_ID,
    participant_id,
    sex,
    SEQ,
    channel,
    time_index_variable,
    slope_estimate,
    intercept,
    r_squared,
    n_points,
    fit_type
  )

if (isTRUE(SHOW_COHORT_TABLES)) {
  print(cohort_ratio_slope_summary_table)
}
```

### 6.4.7 Cohort-ready plotting tables

The cohort plots avoid participant labels so they remain readable when the
sample grows. Individuals are shown with transparent lines, and thicker lines
show cohort means. This chunk only prepares plotting inputs; all figures are
kept together below in a dedicated plotting subsection.

```{r cohort_analysis_plot_styling_tables, eval=FALSE}
cohort_ex_dyn_ratio_plot_table <- cohort_ex_dyn_emg_torque_ratio_table_analysis |>
  dplyr::filter(
    SEQ == "EX_DYN",
    channel %in% selected_channels
  )

cohort_ex_dyn_ratio_mean_table <- cohort_ex_dyn_ratio_plot_table |>
  dplyr::group_by(channel, VC_count) |>
  dplyr::summarise(
    n_runs = sum(is.finite(emg_torque_ratio)),
    mean_emg_torque_ratio  = mean(emg_torque_ratio,  na.rm = TRUE),
    sd_emg_torque_ratio    = sd(emg_torque_ratio,    na.rm = TRUE),
    sem_emg_torque_ratio   = sd_emg_torque_ratio / sqrt(n_runs),
    ci_low_emg_torque_ratio  = mean_emg_torque_ratio - 1.96 * sem_emg_torque_ratio,
    ci_high_emg_torque_ratio = mean_emg_torque_ratio + 1.96 * sem_emg_torque_ratio,
    mean_expected_ratio    = mean(expected_ratio,     na.rm = TRUE),
    .groups = "drop"
  )

cohort_ex_dyn_rms_plot_table <- cohort_ex_dyn_mid1s_feature_table_norm_analysis |>
  dplyr::filter(
    SEQ == "EX_DYN",
    channel %in% selected_channels
  )

cohort_ex_dyn_rms_mean_table <- cohort_ex_dyn_rms_plot_table |>
  dplyr::group_by(channel, VC_count) |>
  dplyr::summarise(
    n_runs = sum(is.finite(rms_norm)),
    mean_rms_norm = mean(rms_norm, na.rm = TRUE),
    sd_rms_norm = sd(rms_norm, na.rm = TRUE),
    sem_rms_norm = sd_rms_norm / sqrt(n_runs),
    ci_low_rms_norm = mean_rms_norm - 1.96 * sem_rms_norm,
    ci_high_rms_norm = mean_rms_norm + 1.96 * sem_rms_norm,
    .groups = "drop"
  )

cohort_ex_dyn_mdf_plot_table <- cohort_ex_dyn_mid1s_feature_table_norm_analysis |>
  dplyr::filter(
    SEQ == "EX_DYN",
    channel %in% selected_channels
  )

cohort_ex_dyn_mdf_mean_table <- cohort_ex_dyn_mdf_plot_table |>
  dplyr::group_by(channel, VC_count) |>
  dplyr::summarise(
    n_runs = sum(is.finite(mdf_mid_1s_hz)),
    mean_mdf_mid_1s_hz = mean(mdf_mid_1s_hz, na.rm = TRUE),
    sd_mdf_mid_1s_hz = sd(mdf_mid_1s_hz, na.rm = TRUE),
    sem_mdf_mid_1s_hz = sd_mdf_mid_1s_hz / sqrt(n_runs),
    ci_low_mdf_mid_1s_hz = mean_mdf_mid_1s_hz - 1.96 * sem_mdf_mid_1s_hz,
    ci_high_mdf_mid_1s_hz = mean_mdf_mid_1s_hz + 1.96 * sem_mdf_mid_1s_hz,
    .groups = "drop"
  )

cohort_ex_sta_ratio_plot_table <- cohort_ex_sta_emg_torque_ratio_table_analysis |>
  dplyr::filter(
    SEQ == "EX_STA",
    channel %in% selected_channels
  ) |>
  dplyr::group_by(RUN_ID, participant_id, channel) |>
  dplyr::arrange(window_mid_s, .by_group = TRUE) |>
  dplyr::mutate(
    window_time_from_start_s = window_mid_s - min(window_mid_s, na.rm = TRUE)
  ) |>
  dplyr::ungroup()

cohort_ex_sta_ratio_mean_table <- cohort_ex_sta_ratio_plot_table |>
  dplyr::group_by(channel, window_time_from_start_s) |>
  dplyr::summarise(
    n_runs = sum(is.finite(emg_torque_ratio)),
    mean_emg_torque_ratio = mean(emg_torque_ratio, na.rm = TRUE),
    sd_emg_torque_ratio = sd(emg_torque_ratio, na.rm = TRUE),
    sem_emg_torque_ratio = sd_emg_torque_ratio / sqrt(n_runs),
    ci_low_emg_torque_ratio = mean_emg_torque_ratio - 1.96 * sem_emg_torque_ratio,
    ci_high_emg_torque_ratio = mean_emg_torque_ratio + 1.96 * sem_emg_torque_ratio,
    .groups = "drop"
  )

cohort_ex_sta_rms_plot_table <- cohort_ex_sta_mid1s_feature_table_norm_analysis |>
  dplyr::left_join(
    cohort_ex_sta_emg_torque_ratio_table_analysis |>
      dplyr::select(RUN_ID, participant_id, channel, window_id, window_mid_s),
    by = c("RUN_ID", "participant_id", "channel", "window_id")
  ) |>
  dplyr::filter(
    SEQ == "EX_STA",
    channel %in% selected_channels
  ) |>
  dplyr::group_by(RUN_ID, participant_id, channel) |>
  dplyr::arrange(window_mid_s, .by_group = TRUE) |>
  dplyr::mutate(
    window_time_from_start_s = window_mid_s - min(window_mid_s, na.rm = TRUE)
  ) |>
  dplyr::ungroup()

cohort_ex_sta_rms_mean_table <- cohort_ex_sta_rms_plot_table |>
  dplyr::group_by(channel, window_time_from_start_s) |>
  dplyr::summarise(
    n_runs = sum(is.finite(rms_norm)),
    mean_rms_norm = mean(rms_norm, na.rm = TRUE),
    sd_rms_norm = sd(rms_norm, na.rm = TRUE),
    sem_rms_norm = sd_rms_norm / sqrt(n_runs),
    ci_low_rms_norm = mean_rms_norm - 1.96 * sem_rms_norm,
    ci_high_rms_norm = mean_rms_norm + 1.96 * sem_rms_norm,
    .groups = "drop"
  )

cohort_ex_sta_mdf_plot_table <- cohort_ex_sta_rms_plot_table

cohort_ex_sta_mdf_mean_table <- cohort_ex_sta_mdf_plot_table |>
  dplyr::group_by(channel, window_time_from_start_s) |>
  dplyr::summarise(
    n_runs = sum(is.finite(mdf_mid_1s_hz)),
    mean_mdf_mid_1s_hz = mean(mdf_mid_1s_hz, na.rm = TRUE),
    sd_mdf_mid_1s_hz = sd(mdf_mid_1s_hz, na.rm = TRUE),
    sem_mdf_mid_1s_hz = sd_mdf_mid_1s_hz / sqrt(n_runs),
    ci_low_mdf_mid_1s_hz = mean_mdf_mid_1s_hz - 1.96 * sem_mdf_mid_1s_hz,
    ci_high_mdf_mid_1s_hz = mean_mdf_mid_1s_hz + 1.96 * sem_mdf_mid_1s_hz,
    .groups = "drop"
  )

cohort_ex_dyn_boxplot_table <- cohort_ex_dyn_mid1s_feature_table_norm_analysis |>
  dplyr::left_join(
    cohort_ex_dyn_emg_torque_ratio_table_analysis |>
      dplyr::select(RUN_ID, participant_id, channel, VC_count, emg_torque_ratio),
    by = c("RUN_ID", "participant_id", "channel", "VC_count")
  ) |>
  dplyr::filter(
    SEQ == "EX_DYN",
    channel %in% selected_channels
  ) |>
  dplyr::group_by(RUN_ID, participant_id, channel) |>
  dplyr::arrange(VC_count, .by_group = TRUE) |>
  dplyr::mutate(
    phase_id = dplyr::ntile(VC_count, 3),
    phase = dplyr::case_when(
      phase_id == 1 ~ "Early",
      phase_id == 2 ~ "Middle",
      phase_id == 3 ~ "Late"
    ),
    phase = factor(phase, levels = c("Early", "Middle", "Late"))
  ) |>
  dplyr::ungroup()

cohort_ex_sta_boxplot_table <- cohort_ex_sta_rms_plot_table |>
  dplyr::left_join(
    cohort_ex_sta_emg_torque_ratio_table_analysis |>
      dplyr::select(RUN_ID, participant_id, channel, window_id, emg_torque_ratio),
    by = c("RUN_ID", "participant_id", "channel", "window_id")
  ) |>
  dplyr::group_by(RUN_ID, participant_id, channel) |>
  dplyr::arrange(window_time_from_start_s, .by_group = TRUE) |>
  dplyr::mutate(
    phase_id = dplyr::ntile(window_time_from_start_s, 3),
    phase = dplyr::case_when(
      phase_id == 1 ~ "Early",
      phase_id == 2 ~ "Middle",
      phase_id == 3 ~ "Late"
    ),
    phase = factor(phase, levels = c("Early", "Middle", "Late"))
  ) |>
  dplyr::ungroup()
```

### 6.4.8 Cohort plots

All cohort figures are grouped below so the notebook is easier to scroll and
review after the summary and preparation chunks above.

#### 6.4.8.1 EX_DYN cohort ratio trajectories - all participants

```{r cohort_analysis_ex_dyn_ratio_plot_all, eval=FALSE, fig.width=10, fig.height=6}
cohort_ex_dyn_ratio_plot_all <- ggplot(
  cohort_ex_dyn_ratio_mean_table,
  aes(x = VC_count, y = mean_emg_torque_ratio)
) +
  geom_ribbon(
    aes(ymin = ci_low_emg_torque_ratio, ymax = ci_high_emg_torque_ratio),
    fill = "#bdbdbd",
    alpha = 0.35
  ) +
  geom_line(
    color = "#1a1a1a",
    linewidth = 1.1
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title    = "Cohort EX_DYN - EMG/torque ratio over repetitions",
    subtitle = "Black = cohort mean; grey ribbon = 95% CI",
    x = "Repetition (VC_count)",
    y = "EMG/torque ratio (RMS_norm / torque_norm)"
  )

```

#### 6.4.8.2 EX_STA cohort ratio trajectories - all participants

For EX_STA, the cohort plot uses time since the first retained window within
that run. This avoids mixing subject-specific absolute run clocks.

```{r cohort_analysis_ex_sta_ratio_plot_all, eval=FALSE, fig.width=10, fig.height=6}
cohort_ex_sta_ratio_plot_all <- ggplot(
  cohort_ex_sta_ratio_mean_table,
  aes(x = window_time_from_start_s, y = mean_emg_torque_ratio)
) +
  geom_ribbon(
    aes(ymin = ci_low_emg_torque_ratio, ymax = ci_high_emg_torque_ratio),
    fill = "#bdbdbd",
    alpha = 0.35
  ) +
  geom_line(
    color = "#1a1a1a",
    linewidth = 1.1
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "Cohort EX_STA - EMG/torque ratio over relative time",
    subtitle = "Black line = cohort mean; grey ribbon = 95% CI",
    x = "Time since first retained EX_STA window (s)",
    y = "EMG/torque ratio (RMS_norm / torque_norm)"
  )

```

#### 6.4.8.3 EX_DYN cohort RMS trajectories - all participants

```{r cohort_analysis_ex_dyn_rms_plot_all, eval=FALSE, fig.width=10, fig.height=6}
cohort_ex_dyn_rms_plot_all <- ggplot(
  cohort_ex_dyn_rms_mean_table,
  aes(x = VC_count, y = mean_rms_norm)
) +
  geom_ribbon(
    aes(ymin = ci_low_rms_norm, ymax = ci_high_rms_norm),
    fill = "#bdbdbd",
    alpha = 0.35
  ) +
  geom_line(
    color = "#1a1a1a",
    linewidth = 1.1
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "Cohort EX_DYN - Normalized RMS across repetitions",
    subtitle = "Black line = cohort mean; grey ribbon = 95% CI",
    x = "Repetition (VC_count)",
    y = "Normalized RMS"
  )

```

#### 6.4.8.4 EX_STA cohort RMS trajectories - all participants

```{r cohort_analysis_ex_sta_rms_plot_all, eval=FALSE, fig.width=10, fig.height=6}
cohort_ex_sta_rms_plot_all <- ggplot(
  cohort_ex_sta_rms_mean_table,
  aes(x = window_time_from_start_s, y = mean_rms_norm)
) +
  geom_ribbon(
    aes(ymin = ci_low_rms_norm, ymax = ci_high_rms_norm),
    fill = "#bdbdbd",
    alpha = 0.35
  ) +
  geom_line(
    color = "#1a1a1a",
    linewidth = 1.1
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "Cohort EX_STA - Normalized RMS over relative time",
    subtitle = "Black line = cohort mean; grey ribbon = 95% CI",
    x = "Time since first retained EX_STA window (s)",
    y = "Normalized RMS"
  )

```

#### 6.4.8.5 EX_DYN cohort MDF trajectories - all participants

```{r cohort_analysis_ex_dyn_mdf_plot_all, eval=FALSE, fig.width=10, fig.height=6}
cohort_ex_dyn_mdf_plot_all <- ggplot(
  cohort_ex_dyn_mdf_mean_table,
  aes(x = VC_count, y = mean_mdf_mid_1s_hz)
) +
  geom_ribbon(
    aes(ymin = ci_low_mdf_mid_1s_hz, ymax = ci_high_mdf_mid_1s_hz),
    fill = "#bdbdbd",
    alpha = 0.35
  ) +
  geom_line(
    color = "#1a1a1a",
    linewidth = 1.1
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "Cohort EX_DYN - MDF across repetitions",
    subtitle = "Black line = cohort mean; grey ribbon = 95% CI",
    x = "Repetition (VC_count)",
    y = "Median frequency (Hz)"
  )

```

#### 6.4.8.6 EX_STA cohort MDF trajectories - all participants

```{r cohort_analysis_ex_sta_mdf_plot_all, eval=FALSE, fig.width=10, fig.height=6}
cohort_ex_sta_mdf_plot_all <- ggplot(
  cohort_ex_sta_mdf_mean_table,
  aes(x = window_time_from_start_s, y = mean_mdf_mid_1s_hz)
) +
  geom_ribbon(
    aes(ymin = ci_low_mdf_mid_1s_hz, ymax = ci_high_mdf_mid_1s_hz),
    fill = "#bdbdbd",
    alpha = 0.35
  ) +
  geom_line(
    color = "#1a1a1a",
    linewidth = 1.1
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "Cohort EX_STA - MDF over relative time",
    subtitle = "Black line = cohort mean; grey ribbon = 95% CI",
    x = "Time since first retained EX_STA window (s)",
    y = "Median frequency (Hz)"
  )

```

#### 6.4.8.7 EX_DYN combined fatigue trajectory (RMS, MDF, ratio)

```{r cohort_analysis_ex_dyn_combined_line, eval=FALSE, fig.width=11, fig.height=8}
# Stack the three already-computed mean/CI tables into a single long frame.
cohort_ex_dyn_combined_long <- dplyr::bind_rows(
  cohort_ex_dyn_rms_mean_table |>
    dplyr::transmute(
      channel, VC_count,
      metric    = "Normalized RMS",
      mean_val  = mean_rms_norm,
      ci_low    = ci_low_rms_norm,
      ci_high   = ci_high_rms_norm
    ),
  cohort_ex_dyn_mdf_mean_table |>
    dplyr::transmute(
      channel, VC_count,
      metric    = "MDF (Hz)",
      mean_val  = mean_mdf_mid_1s_hz,
      ci_low    = ci_low_mdf_mid_1s_hz,
      ci_high   = ci_high_mdf_mid_1s_hz
    ),
  cohort_ex_dyn_ratio_mean_table |>
    dplyr::transmute(
      channel, VC_count,
      metric    = "EMG/torque ratio",
      mean_val  = mean_emg_torque_ratio,
      ci_low    = ci_low_emg_torque_ratio,
      ci_high   = ci_high_emg_torque_ratio
    )
) |>
  dplyr::mutate(metric = factor(metric, levels = c("Normalized RMS", "MDF (Hz)", "EMG/torque ratio")))

cohort_ex_dyn_combined_fatigue_plot <- ggplot(
  cohort_ex_dyn_combined_long,
  aes(x = VC_count, y = mean_val)
) +
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high), fill = "#bdbdbd", alpha = 0.35) +
  geom_line(color = "#1a1a1a", linewidth = 1.1) +
  facet_grid(metric ~ channel, scales = "free_y") +
  labs(
    title    = "Cohort EX_DYN - Fatigue progression: RMS, MDF, EMG/torque ratio",
    subtitle = "Black line = cohort mean; grey ribbon = 95% CI",
    x        = "Repetition (VC_count)",
    y        = NULL
  )

```

#### 6.4.8.8 EX_STA combined fatigue trajectory (RMS, MDF, ratio)

```{r cohort_analysis_ex_sta_combined_line, eval=FALSE, fig.width=11, fig.height=8}
cohort_ex_sta_combined_long <- dplyr::bind_rows(
  cohort_ex_sta_rms_mean_table |>
    dplyr::transmute(
      channel, window_time_from_start_s,
      metric    = "Normalized RMS",
      mean_val  = mean_rms_norm,
      ci_low    = ci_low_rms_norm,
      ci_high   = ci_high_rms_norm
    ),
  cohort_ex_sta_mdf_mean_table |>
    dplyr::transmute(
      channel, window_time_from_start_s,
      metric    = "MDF (Hz)",
      mean_val  = mean_mdf_mid_1s_hz,
      ci_low    = ci_low_mdf_mid_1s_hz,
      ci_high   = ci_high_mdf_mid_1s_hz
    ),
  cohort_ex_sta_ratio_mean_table |>
    dplyr::transmute(
      channel, window_time_from_start_s,
      metric    = "EMG/torque ratio",
      mean_val  = mean_emg_torque_ratio,
      ci_low    = ci_low_emg_torque_ratio,
      ci_high   = ci_high_emg_torque_ratio
    )
) |>
  dplyr::mutate(metric = factor(metric, levels = c("Normalized RMS", "MDF (Hz)", "EMG/torque ratio")))

cohort_ex_sta_combined_fatigue_plot <- ggplot(
  cohort_ex_sta_combined_long,
  aes(x = window_time_from_start_s, y = mean_val)
) +
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high), fill = "#bdbdbd", alpha = 0.35) +
  geom_line(color = "#1a1a1a", linewidth = 1.1) +
  facet_grid(metric ~ channel, scales = "free_y") +
  labs(
    title    = "Cohort EX_STA - Fatigue progression: RMS, MDF, EMG/torque ratio",
    subtitle = "Black line = cohort mean; grey ribbon = 95% CI",
    x        = "Time since first retained window (s)",
    y        = NULL
  )

```

#### 6.4.8.9 Ratio slope overview plot

Each run's fitted slope is shown as a line over its observed x range.
The blue line is the cohort mean slope (mean intercept + mean slope × x).

```{r cohort_analysis_ratio_slope_plot, eval=FALSE, fig.width=10, fig.height=5}
# EX_DYN only — EX_STA slopes are near-flat (constant-load hold) and add little.
# Build individual slope lines: two endpoints per run (straight lines).
slope_lines_df <- cohort_ratio_slope_summary_table |>
  dplyr::filter(SEQ == "EX_DYN") |>
  dplyr::mutate(x_min = 1.0, x_max = as.numeric(n_points)) |>
  tidyr::pivot_longer(c(x_min, x_max), names_to = "endpoint", values_to = "x") |>
  dplyr::mutate(y = intercept + slope_estimate * x) |>
  dplyr::select(RUN_ID, channel, x, y)

# Pooled slope: mean intercept + mean slope per channel.
pooled_slope_params <- cohort_ratio_slope_summary_table |>
  dplyr::filter(SEQ == "EX_DYN") |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    mean_intercept = mean(intercept,      na.rm = TRUE),
    mean_slope     = mean(slope_estimate, na.rm = TRUE),
    x_max          = max(n_points),
    .groups        = "drop"
  ) |>
  dplyr::mutate(x_min = 1.0)

pooled_slope_df <- pooled_slope_params |>
  tidyr::pivot_longer(c(x_min, x_max), names_to = "endpoint", values_to = "x") |>
  dplyr::mutate(y = mean_intercept + mean_slope * x) |>
  dplyr::select(channel, x, y)

cohort_ratio_slope_plot <- ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#4d4d4d", linewidth = 0.5) +
  geom_line(
    data  = slope_lines_df,
    aes(x = x, y = y, group = RUN_ID),
    color = "#4d4d4d", alpha = 0.35, linewidth = 0.7
  ) +
  geom_line(
    data  = pooled_slope_df,
    aes(x = x, y = y),
    color = "#1b6ca8", linewidth = 1.4
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title    = "Cohort EX_DYN - EMG/torque ratio slopes",
    subtitle = "Thin grey = individual run slopes; solid blue = cohort mean slope",
    x        = "Repetition (VC_count)",
    y        = "Predicted ratio (intercept + slope \u00d7 VC_count)"
  )

```

#### 6.4.8.10 Cohort RMS across intensity levels

Boxplot distribution across runs per intensity level, with individual run
trajectories shown in alpha so the spread and the direction of each
participant's calibration curve are both visible.
Dashed line = MVC_REF = 1 (normalization anchor).

```{r cohort_rms_calibration_boxplot, eval=FALSE, fig.width=10, fig.height=5}
cohort_rms_calib_plot_table <- cohort_reference_level_feature_table_norm_analysis |>
  dplyr::filter(channel %in% selected_channels) |>
  dplyr::mutate(
    reference_level = factor(reference_level, levels = c("30%", "50%", "70%", "100%"))
  )

cohort_rms_calibration_boxplot <- ggplot(
  cohort_rms_calib_plot_table,
  aes(x = reference_level, y = rms_norm)
) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "#888888", linewidth = 0.55) +
  geom_line(
    aes(group = RUN_ID),
    color = "#4d4d4d", alpha = 0.25, linewidth = 0.7
  ) +
  geom_point(
    aes(group = RUN_ID),
    color = "#4d4d4d", alpha = 0.35, size = 1.5
  ) +
  stat_summary(
    fun = mean, geom = "line", aes(group = 1),
    color = "#1b6ca8", linewidth = 1.2
  ) +
  stat_summary(
    fun = mean, geom = "point", aes(group = 1),
    color = "#1b6ca8", size = 2.5
  ) +
  facet_wrap(~ channel) +
  labs(
    title    = "Cohort - Normalized RMS across intensity levels",
    subtitle = "Solid blue = cohort mean; grey lines = individual run trajectories; dashed = MVC_REF = 1",
    x        = "Reference intensity level",
    y        = "Normalized RMS (MVC_REF = 1)"
  )

print(cohort_rms_calibration_boxplot)
```

#### 6.4.8.11 Cohort MDF across intensity levels

MDF in raw Hz (not normalized). Each channel has a different Hz range so
scales are free across facets.

```{r cohort_mdf_calibration_boxplot, eval=FALSE, fig.width=10, fig.height=5}
cohort_mdf_calib_plot_table <- cohort_reference_level_feature_table_norm_analysis |>
  dplyr::filter(channel %in% selected_channels) |>
  dplyr::mutate(
    reference_level = factor(reference_level, levels = c("30%", "50%", "70%", "100%"))
  )

cohort_mdf_calibration_boxplot <- ggplot(
  cohort_mdf_calib_plot_table,
  aes(x = reference_level, y = mdf_mid_1s_hz)
) +
  geom_line(
    aes(group = RUN_ID),
    color = "#4d4d4d", alpha = 0.25, linewidth = 0.7
  ) +
  geom_point(
    aes(group = RUN_ID),
    color = "#4d4d4d", alpha = 0.35, size = 1.5
  ) +
  stat_summary(
    fun = mean, geom = "line", aes(group = 1),
    color = "#d17a22", linewidth = 1.2
  ) +
  stat_summary(
    fun = mean, geom = "point", aes(group = 1),
    color = "#d17a22", size = 2.5
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title    = "Cohort - MDF across intensity levels",
    subtitle = "Solid orange = cohort mean; grey lines = individual run trajectories",
    x        = "Reference intensity level",
    y        = "Median frequency (Hz)"
  )

print(cohort_mdf_calibration_boxplot)
```

#### 6.4.8.12 Cohort plot gallery

This final chunk prints all cohort figures together so they can be reviewed in
one continuous visual block.

```{r cohort_analysis_plot_gallery, eval=FALSE, fig.width=11, fig.height=7}
if (isTRUE(SHOW_COHORT_PLOTS)) {
  print(cohort_rms_calibration_boxplot)
  print(cohort_mdf_calibration_boxplot)
  print(cohort_ex_dyn_ratio_plot_all)
  print(cohort_ex_sta_ratio_plot_all)
  print(cohort_ex_dyn_rms_plot_all)
  print(cohort_ex_sta_rms_plot_all)
  print(cohort_ex_dyn_mdf_plot_all)
  print(cohort_ex_sta_mdf_plot_all)
  print(cohort_ex_dyn_combined_fatigue_plot)
  print(cohort_ex_sta_combined_fatigue_plot)
  print(cohort_ratio_slope_plot)
}
```

### 6.4.9 Simple cohort-level checks for future extensions

```{r cohort_analysis_extension_note, eval=FALSE}
message(
  "Cohort EMG sanity-check plots completed. ",
  "Future modality-specific analyses should follow the same pattern: ",
  "read exported pooled parquet, build explicit summary tables, then plot."
)
```


# 7. Summary

## 7.1 Export report

```{r summary_export_report}
message("Runs combined: ", length(run_ids))
message("Cohort parquet export enabled: ", EXPORT_COHORT_PARQUET)
if (isTRUE(EXPORT_COHORT_PARQUET)) {
  message("Cohort outputs written to: ", normalizePath(cohort_output_root, winslash = "/", mustWork = FALSE))
} else {
  message("Cohort outputs were kept in memory only for this notebook run.")
}
```
