---
title: "EMG Fatigue Analysis Pipeline"
author: "Morgan Virolan"
date: "2026-02-20"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: cosmo


---

# Scientific question and Objectives
This is the companion notebook for the AMSTROM Campain, where we analyze the sEMG and torque data from the fatigue protocol. It's main objectives is (at this point) to quntifify the fatigue-induced changes in sEMG features (RMS, MDF) during the dynamic (EX_DYN) and static (EX_STA) fatigue sequences, and to see how they relate to the torque loss.
It's the first step of the broader study which will also include the analysis of other physiological signals recorded during the same experimental protocole (BIA, NIRS, Mecanical values) + stats & ML.

Dataset structure
Signal assumptions (delsys hardware band-pass ?? "slow" ramp up -> cut middle sequence + fist windows on STA)

# Data and Preprocessing Framework
```{r setup, message=FALSE, warning=FALSE}
library(arrow)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(signal)   # butter(), filtfilt()
library(glue)


options(dplyr.summarise.inform = FALSE)
```

```{r load_data}
RUN_ID <- "101SaCl_20251127"
subject_directory <- "results/DATA_EXPORT" |> 
  file.path(RUN_ID) |>
  glue::glue()

library(arrow)
library(dplyr)

master_file_path <- file.path(subject_directory, "02_master.parquet")
emg_file_path    <- file.path(subject_directory, "02_emg_compact.parquet")
torque_file_path <- file.path(subject_directory, "02_torque_compact.parquet")

master_table <- read_parquet(master_file_path)
emg_table    <- read_parquet(emg_file_path)
torque_table <- read_parquet(torque_file_path)

analysis_table <- master_table |>
  inner_join(
    emg_table |> select(-VC, -VC_count, -SEQ, -SEQ_index),
    by = "time_index"
  ) |>
  left_join(
    torque_table |> select(-VC, -VC_count, -SEQ, -SEQ_index),
    by = "time_index"
  )

cat("nrow(master_table):", nrow(master_table), "\n")
cat("nrow(emg_table):   ", nrow(emg_table), "\n")
cat("nrow(torque_table):", nrow(torque_table), "\n")
cat("nrow(analysis_table):", nrow(analysis_table), "\n")

## master & emg must share the same hz, torque is expected to be about 1/5 of that and do contain n/a
analysis_table %>%
  summarise(
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est_hz = n() / duration_s,
    torque_na_frac = mean(is.na(torque_raw))
  )
```




```{r mvc_svc_load}
# MVC_REF = 3 to 4 reps of maximal voluntary contraction (MVC) at the start of the session
mvc_ref_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_REF",
              .data[["VC"]] == 1) |> # only keep actual reps (voluntary contraction = 1) / no rest
  dplyr::arrange(.data[["time_index"]])


#SVC_REF = submaximal voluntary contractions at 30/50/70% of MVC, used for calibration of torque and EMG
svc_ref_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "SVC_REF",
              .data[["VC"]] == 1) |> 
  dplyr::arrange(.data[["time_index"]])


#MVC_RECOV_DYN = MVC checkup after the dynamic fatigue sequence (EX_DYN) + 5 mins rest -  used to assess recovery oafter fatigue

mvc_recov_dyn_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_RECOV_DYN",
              .data[["VC"]] == 1) |> 
  dplyr::arrange(.data[["time_index"]])


#MVC_RECOV_STA = MVC checkup after the static fatigue sequence (EX_STA) + 5 mins rest -  used to assess recovery after fatigue

mvc_recov_sta_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_RECOV_STA",
              .data[["VC"]] == 1) |> 
  dplyr::arrange(.data[["time_index"]])

```


```{r ex_dyn_ex_sta_load}
## EX_DYN = dynamic fatigue sequence (62 reps, 3s efforts 2s recup, max effort each rep)
ex_dyn_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "EX_DYN",
                .data[["VC"]] == 1) |>
  dplyr::arrange(.data[["time_index"]])

## EX_STA = static fatigue sequence (1 rep, 60s effort, 70% of baseline mvc)
ex_sta_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "EX_STA",
                .data[["VC"]] == 1) |> #just in case sequencing is loose (everything should be VC==1)
  dplyr::arrange(.data[["time_index"]])

```
## Dataset Structure

## Torque Calibration
```{r torque_calibration}
#set values, from the biodex calibration
torque_offset_volt <- 0.106
torque_gain_Nm_per_volt <- 145.30

calc_torque_nm <- function(df) {
  df %>% dplyr::mutate(torque_Nm = torque_gain_Nm_per_volt * (torque_raw - torque_offset_volt))
}

ex_dyn_table       <- calc_torque_nm(ex_dyn_table)
ex_sta_table       <- calc_torque_nm(ex_sta_table)
mvc_ref_table      <- calc_torque_nm(mvc_ref_table)
svc_ref_table      <- calc_torque_nm(svc_ref_table)
mvc_recov_dyn_table <- calc_torque_nm(mvc_recov_dyn_table)
mvc_recov_sta_table <- calc_torque_nm(mvc_recov_sta_table)

#print the 10 lines of each + 10 lines of the raw table to see the difference
print(head(ex_dyn_table |> dplyr::select(time_ref_s, torque_raw, torque_Nm), 10))

```

CLEANUP

```{r cleanup_raw_torque}
#cleanup raw torque (keep only the calibrated torque in Nm to save memory)
# ex_dyn_table <- ex_dyn_table |> dplyr::select(-torque_raw)
# ex_sta_table <- ex_sta_table |> dplyr::select(-torque_raw)
# mvc_ref_table <- mvc_ref_table |> dplyr::select(-torque_raw)
# svc_ref_table <- svc_ref_table |> dplyr::select(-torque_raw)
# mvc_recov_dyn_table <- mvc_recov_dyn_table |> dplyr::select(-torque_raw)
# mvc_recov_sta_table <- mvc_recov_sta_table |> dplyr::select(-torque_raw)

print(head(ex_dyn_table, 10))                                           

```
```{r cleanup_analysis_table}
#remove the analysis table  - we extracted what we need for now
rm(analysis_table)
rm(emg_table)
rm(torque_table)
gc()
```

## Signal Processing 

```{r EMG_filtering}
# CONSTANTS
emg_sampling_frequency_hz <- 2148.149

emg_bandpass_low_hz <- 20
emg_bandpass_high_hz <- 450
emg_filter_order <- 4


# BANDPASS FILTER FUNCTION -----------------------------------
# clean artefacts, movement related noises, high freq noise.
emg_bandpass_filter <- function(emg_signal, sampling_frequency_hz, 
                                low_hz, high_hz, filter_order) {
  
  # Centers signal around zero by subtracting the mean.
  # This prevents a large "spike" at 0Hz in the frequency domain (dc offset removal)
  emg_centered <- emg_signal - mean(emg_signal, na.rm = TRUE)
  
  # Nyquist Normalization : cutoffs needed for digital filters
  # ratio of the Nyquist frequency (half the sampling rate).
  normalized_cutoffs <- c(low_hz, high_hz) / (sampling_frequency_hz / 2)
  
  # Butterworth filter 
  butter_filter <- signal::butter(filter_order, normalized_cutoffs, type = "pass")
  
  # Zero-Phase Filtering
  # filter twice (forward and backward) -> no time-shift (phase lag
  signal::filtfilt(butter_filter, emg_centered)
}


```


# Definition of the Analysis Window

Adds markdown later to justify windows (comparable, avoid artefacts dues to ramps, enough hz to still run everything, etc)

## Contraction Labels


## Middle 1 s Window Extraction

```{r extract_middle_1s_window_function}
extract_middle_1s_window <- function(rep_df, window_length_s = 1.0) {

  rep_df |>
    dplyr::group_by(.data$VC_count) |>
    dplyr::mutate(
      t_mid = (min(.data$time_ref_s) + max(.data$time_ref_s)) / 2,
      mid_start = .data$t_mid - window_length_s / 2,
      mid_end   = .data$t_mid + window_length_s / 2
    ) |>
    dplyr::filter(
      .data$time_ref_s >= mid_start,
      .data$time_ref_s < mid_end
    ) |>
    dplyr::select(-.data$t_mid, -.data$mid_start, -.data$mid_end) |>
    dplyr::ungroup()
}
```

```{r build_1s_table}
#mvc_ref
mvc_ref_mid1s_table <- extract_middle_1s_window(mvc_ref_table)

mvc_ref_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

#svc_ref
svc_ref_mid1s_table <- extract_middle_1s_window(svc_ref_table)

svc_ref_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

#mvc_recov_dyn
mvc_recov_dyn_mid1s_table <- extract_middle_1s_window(mvc_recov_dyn_table)

mvc_recov_dyn_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

#mvc_recov_sta
mvc_recov_sta_mid1s_table <- extract_middle_1s_window(mvc_recov_sta_table)

mvc_recov_sta_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

#ex_dyn
ex_dyn_mid1s_table <- extract_middle_1s_window(ex_dyn_table)

ex_dyn_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

```



```{r extract_ex_sta_middle_1s_window}
# EX_STA window grid (1 s, non-overlap)

window_length_s <- 1.0 #must match
window_step_s   <- 1.0

t_start <- min(ex_sta_table$time_ref_s)
t_end   <- max(ex_sta_table$time_ref_s)

window_start_s <- seq(t_start, t_end - window_length_s, by = window_step_s)

ex_sta_window_table <- tibble::tibble(
  window_id      = seq_along(window_start_s),
  window_start_s = window_start_s,
  window_end_s   = window_start_s + window_length_s,
  window_mid_s   = window_start_s + window_length_s / 2
)

nrow(ex_sta_window_table)
ex_sta_window_table |> dplyr::slice(1:5)

## attach data

ex_sta_mid1s_table <- ex_sta_table |>
  dplyr::mutate(
    window_id = floor((time_ref_s - t_start) / window_step_s) + 1L
  ) |>
  dplyr::filter(
    window_id >= 1L,
    window_id <= nrow(ex_sta_window_table)
  ) |>
  dplyr::left_join(
    ex_sta_window_table,
    by = "window_id"
  ) |>
  dplyr::select(
    time_index, time_ref_s, SEQ_index, SEQ, VC, VC_count,
    window_id, window_start_s, window_end_s, window_mid_s,
    emg6, emg8, emg10, torque_raw, torque_Nm
  )

# quick check: should show ~1.0 s duration per window
ex_sta_mid1s_table |>
  dplyr::group_by(window_id) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    .groups = "drop"
  ) |>
  dplyr::slice(1:10) |>
  print()
```

# Feature Extraction

Markdown explications + features etc. 
All EMG features are computed on previously defined 1 s windows.


## EMG Amplitude Descriptor (RMS)



```{r RMS_function}
# RMS FUNCTION -----------------------------------
# Squares every point (-> everything becomes positive), finds average, takes the root -> power of the signal (same unit). amplitude/intensity of the contraction.
emg_rms <- function(emg_signal) {
  sqrt(mean(emg_signal^2, na.rm = TRUE))
}
```


## EMG Spectral Descriptor (Median Frequency)

markdown explications 

```{r MDF_function}
# MDF

emg_median_frequency_hz <- function(emg_signal, sampling_frequency_hz) {
  emg_centered <- emg_signal - mean(emg_signal) #recompute, last one is nested and global var is a dangerous world.
  n_samples <- length(emg_centered)
# Hann window to reduce spectral leakage 
  hann_window <- 0.5 - 0.5 * cos(2 * pi * (0:(n_samples - 1)) / (n_samples - 1))
  emg_windowed <- emg_centered * hann_window

# Single-sided power spectrum (FFT)
  fft_complex <- stats::fft(emg_windowed)
  n_half <- floor(n_samples / 2) + 1
  fft_half <- fft_complex[1:n_half]
  
# Power spectrum (magnitude squared of the FFT, normalized by the number of samples)
  power_spectrum <- (Mod(fft_half) ^ 2) / n_samples
  frequency_axis_hz <- (0:(n_half - 1)) * (sampling_frequency_hz / n_samples)
  
  #drop 0hz bin
  power_spectrum <- power_spectrum[-1]
frequency_axis_hz <- frequency_axis_hz[-1]

  # Cumulative power to find the median frequency
  cumulative_power <- cumsum(power_spectrum)
  median_index <- which(cumulative_power >= 0.5 * sum(power_spectrum))[1]
  
  frequency_axis_hz[median_index]
  }
```


# Construction of Feature Tables




## MVC Reference Features

```{r compute_feature_mvc_ref_mid1s_ref_table}
channels <- c("emg6", "emg8", "emg10")

mvc_ref_split <- split(mvc_ref_mid1s_table, mvc_ref_mid1s_table$VC_count)

mvc_ref_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_ref_split), function(vc_id) {

  block <- mvc_ref_split[[vc_id]]

  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_REF",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

print(mvc_ref_mid1s_feature_table)
```

```{r compute_feature_svc_ref_mid1s_table}

channels <- c("emg6", "emg8", "emg10")

svc_ref_split <- split(svc_ref_mid1s_table, svc_ref_mid1s_table$VC_count)

svc_ref_mid1s_feature_table <- dplyr::bind_rows(lapply(names(svc_ref_split), function(vc_id) {

  block <- svc_ref_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "SVC_REF",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

print(svc_ref_mid1s_feature_table)
svc_ref_mid1s_feature_table |> dplyr::count(channel) |> print()
```

```{r compute_feature_mvc_recov_dyn_mid1s_table}
mvc_recov_dyn_split <- split(mvc_recov_dyn_mid1s_table, mvc_recov_dyn_mid1s_table$VC_count)

mvc_recov_dyn_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_recov_dyn_split), function(vc_id) {

  block <- mvc_recov_dyn_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_RECOV_DYN",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

print(mvc_recov_dyn_mid1s_feature_table)
```

```{r compute_feature_mvc_recov_sta_mid1s_table}
mvc_recov_sta_split <- split(mvc_recov_sta_mid1s_table, mvc_recov_sta_mid1s_table$VC_count)

mvc_recov_sta_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_recov_sta_split), function(vc_id) {

  block <- mvc_recov_sta_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_RECOV_STA",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

print(mvc_recov_sta_mid1s_feature_table)
```

```{r combine_feature_mid1s_tables}

reference_mid1s_feature_table <- dplyr::bind_rows(
  mvc_ref_mid1s_feature_table,
  svc_ref_mid1s_feature_table,
  mvc_recov_dyn_mid1s_feature_table,
  mvc_recov_sta_mid1s_feature_table
) |>
  dplyr::arrange(SEQ, channel, VC_count)

print(reference_mid1s_feature_table)
```

```{r mvc_level_summary}
mvc_level_summary <- reference_mid1s_feature_table |>
  dplyr::filter(SEQ %in% c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")) |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    n_reps = dplyr::n(),
    torque_mean_Nm = mean(torque_mean_Nm, na.rm = TRUE),
    rms_mid_1s_mean = mean(rms_mid_1s, na.rm = TRUE),
    mdf_mid_1s_hz_mean = mean(mdf_mid_1s_hz, na.rm = TRUE),
    .groups = "drop"
  )

mvc_level_summary

###mvc_ref_baseline <- mvc_level_summary |>
mvc_ref_baseline <- mvc_level_summary |>
  dplyr::filter(SEQ == "MVC_REF") |>
  dplyr::select(
    channel,
    torque_ref_Nm = torque_mean_Nm,
    rms_ref = rms_mid_1s_mean,
    mdf_ref_hz = mdf_mid_1s_hz_mean
  )

mvc_ref_baseline
```


## EX_DYN Feature Table

feature table per rep per channel

```{r build_ex_dyn_1mids_feature_table}

channels <- c("emg6", "emg8", "emg10")

ex_dyn_split <- split(ex_dyn_mid1s_table, ex_dyn_mid1s_table$VC_count)

ex_dyn_mid1s_feature_table <- dplyr::bind_rows(lapply(names(ex_dyn_split), function(vc_id) {

  block <- ex_dyn_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "EX_DYN",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

```


## EX_STA Feature Table

```{r compute_feature_ex_sta_mid1s_table}

channels <- c("emg6", "emg8", "emg10")

ex_sta_split <- split(ex_sta_mid1s_table, ex_sta_mid1s_table$window_id)

ex_sta_mid1s_feature_table <- dplyr::bind_rows(lapply(names(ex_sta_split), function(win_id) {

  block <- ex_sta_split[[win_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "EX_STA",
      window_id = as.integer(win_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, window_id)

print(ex_sta_mid1s_feature_table)


```


# Baseline Normalization


## MVC Baseline Computation
```{r compute_mvc_ref_mid1s_rms_baseline_table}
mvc_ref_mid1s_rms_baseline_table <- mvc_ref_mid1s_feature_table |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    rms_mvc_ref_mid1s_baseline = mean(rms_mid_1s, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)

print(mvc_ref_mid1s_rms_baseline_table)
```

```{r mvc_torque_plot_summary}
mvc_torque_plot_df <- mvc_level_summary |>
  dplyr::distinct(SEQ, torque_mean_Nm) |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

ggplot(mvc_torque_plot_df, aes(x = SEQ, y = torque_mean_Nm)) +
  geom_point(size = 3) +
  geom_line(aes(group = 1)) +
  labs(
    title = "MVC — Torque baseline vs recovery",
    x = NULL,
    y = "Torque (Nm)"
  )
```

```{r mvc_recov_plot_summary}
mvc_rms_plot_df <- mvc_level_summary |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

ggplot(mvc_rms_plot_df, aes(x = SEQ, y = rms_mid_1s_mean)) +
  geom_point(size = 2) +
  geom_line(aes(group = channel)) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "MVC — RMS baseline vs recovery (middle 1s)",
    x = NULL,
    y = "RMS (a.u.)"
  )
```

```{r mvc_mdf_plot_summary}
mvc_mdf_plot_df <- mvc_level_summary |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

ggplot(mvc_mdf_plot_df, aes(x = SEQ, y = mdf_mid_1s_hz_mean)) +
  geom_point(size = 2) +
  geom_line(aes(group = channel)) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "MVC — MDF baseline vs recovery (middle 1s)",
    x = NULL,
    y = "MDF (Hz)"
  )
```

```{r ex_dyn_peak_extract}
#turn it dynamic later, if we keep ?
ex_dyn_peak <- ex_dyn_mid1s_feature_table |>
  dplyr::slice_max(torque_mean_Nm, n = 3) |>
  dplyr::mutate(SEQ = "EX_DYN_PEAK")

ex_dyn_peak

# mvc_dyn_recovery_table}

mvc_dyn_recovery_df <- dplyr::bind_rows(
  mvc_level_summary,
  ex_dyn_peak
) |>
  dplyr::filter(
    SEQ %in% c("MVC_REF", "EX_DYN_PEAK", "MVC_RECOV_DYN", "MVC_RECOV_STA")
  )

mvc_dyn_recovery_df
```





```{r mvc_dyn_recovery_plot}

# ex_dynpeak (not dynamic yet)
ex_dyn_peak_h <- ex_dyn_mid1s_feature_table |>
  dplyr::slice_max(torque_mean_Nm, n = 3) |>
  dplyr::mutate(
    SEQ = "EX_DYN_PEAK",
    n_reps = 1L,
    rms_mid_1s_mean = rms_mid_1s,
    mdf_mid_1s_hz_mean = mdf_mid_1s_hz
  ) |>
  dplyr::select(
    SEQ, channel, n_reps,
    torque_mean_Nm,
    rms_mid_1s_mean,
    mdf_mid_1s_hz_mean
  )

# traj
mvc_dyn_recovery_df <- dplyr::bind_rows(
  mvc_level_summary |>
    dplyr::select(SEQ, channel, n_reps, torque_mean_Nm, rms_mid_1s_mean, mdf_mid_1s_hz_mean),
  ex_dyn_peak_h
) |>
  dplyr::filter(SEQ %in% c("MVC_REF", "EX_DYN_PEAK", "MVC_RECOV_DYN", "MVC_RECOV_STA")) |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "EX_DYN_PEAK", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

# build table
plot_long <- mvc_dyn_recovery_df |>
  dplyr::group_by(SEQ) |>
  dplyr::mutate(torque_mean_Nm = mean(torque_mean_Nm)) |>
  dplyr::ungroup() |>
  tidyr::pivot_longer(
    cols = c(torque_mean_Nm, rms_mid_1s_mean, mdf_mid_1s_hz_mean),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_mean_Nm = "Torque (Nm)",
      rms_mid_1s_mean = "RMS (a.u.)",
      mdf_mid_1s_hz_mean = "MDF (Hz)"
    ),
    metric = factor(metric, levels = c("Torque (Nm)", "RMS (a.u.)", "MDF (Hz)"))
  ) |>
  dplyr::filter(is.finite(value))

# Plot
ggplot(plot_long, aes(x = SEQ, y = value)) +
  geom_point(size = 2) +
  geom_line(aes(group = channel)) +
  facet_grid(metric ~ channel, scales = "free_y") +
  scale_x_discrete(
    labels = c(
      MVC_REF = "MVC",
      EX_DYN_PEAK = "EX_DYN peak",
      MVC_RECOV_DYN = "RECOV dyn",
      MVC_RECOV_STA = "RECOV sta"
    )
  ) +
  labs(
    title = "REF -> DYN -> R_DYN -> R_STA",
    x = NULL,
    y = NULL
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

check (emg 10 ??)

```{r check_emg_10}
# table
mvc_dyn_recovery_check_wide <- mvc_dyn_recovery_df |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "EX_DYN_PEAK", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  ) |>
  dplyr::arrange(channel, SEQ) |>
  dplyr::select(SEQ, channel, torque_mean_Nm, rms_mid_1s_mean, mdf_mid_1s_hz_mean) |>
  tidyr::pivot_wider(
    names_from = SEQ,
    values_from = c(torque_mean_Nm, rms_mid_1s_mean, mdf_mid_1s_hz_mean),
    names_glue = "{.value}__{SEQ}"
  )

mvc_dyn_recovery_check_wide
```


## RMS Normalization


normalize ex_dyn

```{r ex_dyn_mid1s_rms_normalization}
ex_dyn_mid1s_feature_table_norm <- ex_dyn_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm = rms_mid_1s / rms_mvc_ref_mid1s_baseline
  )

print(ex_dyn_mid1s_feature_table_norm)
```


normalize ex_sta
```{r ex_sta_mid1s_rms_normalization}
ex_sta_mid1s_feature_table_norm <- ex_sta_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm = rms_mid_1s / rms_mvc_ref_mid1s_baseline
  )

print(ex_sta_mid1s_feature_table_norm)
```


# Fatigue Trend Modeling

## MVC/SVC Trends

```{r svc_ref_vc_plots_RMS_MF}
svc_df <- svc_ref_mid1s_feature_table |>
  dplyr::mutate(
    VC_count  = as.integer(VC_count),
    intensity = factor(VC_count,
                       levels = c(1,2,3),
                       labels = c("30% ","50% ","70%")),
    torque_Nm = as.numeric(torque_mean_Nm)
  ) |>
  dplyr::filter(
    is.finite(VC_count),
    is.finite(torque_Nm),
    is.finite(rms_mid_1s),
    is.finite(mdf_mid_1s_hz)
  )

p_mdf <- ggplot(svc_df, aes(x = intensity, y = mdf_mid_1s_hz, group = 1)) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "SVC_REF - MDF (middle 1s)",
    x = "Intensity (30/50/70% MVC)",
    y = "MDF (Hz)"
  )

p_mdf

p_rms <- ggplot(svc_df, aes(x = intensity, y = rms_mid_1s, group = 1)) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "SVC_REF - RMS across intensities (middle 1s)",
    x = "Intensity (30/50/70% MVC)",
    y = "RMS (a.u.)"
  )

p_rms
```

```{r percent_emg_vs_percent_force_svc_ref}
# MVC torque baseline 
torque_mvc_ref_baseline <- mvc_ref_mid1s_feature_table |>
  dplyr::summarise(torque_mvc = mean(torque_mean_Nm, na.rm = TRUE)) |>
  dplyr::pull(torque_mvc)

# %EMG and %Force ()
svc_pct_df <- svc_ref_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    torque_pct = 100 * torque_mean_Nm / torque_mvc_ref_baseline,
    emg_pct    = 100 * rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    intensity  = factor(VC_count, levels = c(1,2,3), labels = c("30%","50%","70%"))
  )

# %EMG vs %Force 
ggplot(svc_pct_df, aes(x = torque_pct, y = emg_pct)) +
  geom_point(aes(shape = intensity), size = 2) +
  geom_line(aes(group = 1)) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "SVC_REF - %EMG vs %Force (normalized to MVC_REF)",
    x = "% Force (Torque / MVC_REF torque)",
    y = "% EMG (RMS / MVC_REF RMS)"
  )
```

```{r force_vs_emg_%}
svc_pct_summary_table <- svc_pct_df |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    `Slope (%EMG / %Force)` = coef(lm(emg_pct ~ torque_pct))[2],
    `Pearson r (emg_pct, torque_pct)` = cor(emg_pct, torque_pct, use = "complete.obs"),
    `R² (emg_pct ~ torque_pct)` = summary(lm(emg_pct ~ torque_pct))$r.squared,
    .groups = "drop"
  )

svc_pct_summary_table
```



## EX_DYN Trends

```{r ex_dyn_torque_vs_rep}
ggplot(ex_dyn_mid1s_feature_table_norm,
       aes(x = VC_count, y = torque_mean_Nm)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - Torque across repetitions",
    x = "Repetition (VC_count)",
    y = "Torque (Nm)"
  )
```


```{r ex_dyn_rms_vs_torque}
ggplot(ex_dyn_mid1s_feature_table_norm,
       aes(x = torque_mean_Nm, y = rms_norm)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - EMG activation vs produced torque",
    x = "Torque (Nm)",
    y = "Normalized RMS"
  )
```  


```{r ex_dyn_mdf_vs_rep}
ggplot(ex_dyn_mid1s_feature_table_norm,
       aes(x = VC_count, y = mdf_mid_1s_hz)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - Median frequency across repetitions",
    x = "Repetition (VC_count)",
    y = "Median frequency (Hz)"
  )
```

```{r ex_dyn_mdf_vs_torque}

ggplot(ex_dyn_mid1s_feature_table_norm,
       aes(x = torque_mean_Nm, y = mdf_mid_1s_hz)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - Median frequency vs torque",
    x = "Torque (Nm)",
    y = "Median frequency (Hz)"
  )
```  

```{r ex_dyn_rms_vs_repetition}

ggplot(ex_dyn_mid1s_feature_table_norm,
       aes(x = VC_count, y = rms_norm)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - Normalized RMS across repetitions",
    x = "Repetition (VC_count)",
    y = "Normalized RMS"
  )
```
```{r ex_dyn_combined_fatigue_plot}
ex_dyn_plot_long <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::select(channel, VC_count, torque_mean_Nm, rms_norm, mdf_mid_1s_hz) |>
  tidyr::pivot_longer(
    cols = c(torque_mean_Nm, rms_norm, mdf_mid_1s_hz),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_mean_Nm = "Torque (Nm)",
      rms_norm = "RMS_norm (a.u.)",
      mdf_mid_1s_hz = "MDF (Hz)"
    ),
    metric = factor(metric, levels = c("Torque (Nm)", "RMS_norm (a.u.)", "MDF (Hz)"))
  )

ggplot(ex_dyn_plot_long, aes(x = VC_count, y = value)) +
  geom_point(alpha = 0.6, size = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(metric ~ channel, scales = "free_y") +
  labs(
    title = "EX_DYN - Fatigue progression (middle 1s): Torque, RMS_norm, MDF",
    x = "Repetition (VC_count)",
    y = NULL
  )
```

## EX_STA Trends

```{r ex_sta_triple_panel_fatigue}

ex_sta_plot_long <- ex_sta_mid1s_feature_table_norm |>
  dplyr::select(channel, window_id, torque_mean_Nm, rms_norm, mdf_mid_1s_hz) |>
  tidyr::pivot_longer(
    cols = c(torque_mean_Nm, rms_norm, mdf_mid_1s_hz),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_mean_Nm = "Torque (Nm)",
      rms_norm = "RMS_norm (a.u.)",
      mdf_mid_1s_hz = "MDF (Hz)"
    ),
    metric = factor(metric, levels = c("Torque (Nm)", "RMS_norm (a.u.)", "MDF (Hz)"))
  )

ggplot(ex_sta_plot_long, aes(x = window_id, y = value)) +
  geom_point(alpha = 0.6, size = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(metric ~ channel, scales = "free_y") +
  labs(
    title = "EX_STA - Fatigue progression (1 s windows)",
    x = "Time (s)",
    y = NULL
  )
```


## Trend Summary Table

EX_DYN


```{r ex_dyn_summary_slopes_with_units}

ex_dyn_summary_table <-
  ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::summarise(

    `Torque slope (Nm / rep)` =
      coef(lm(torque_mean_Nm ~ VC_count))[2],

    `RMS_norm slope (Δ RMS_norm / rep)` =
      coef(lm(rms_norm ~ VC_count))[2],

    `MDF slope (Hz / rep)` =
      coef(lm(mdf_mid_1s_hz ~ VC_count))[2],

    `RMS_norm vs Torque slope (Δ RMS_norm / Nm)` =
      coef(lm(rms_norm ~ torque_mean_Nm))[2],

    `MDF vs Torque slope (Hz / Nm)` =
      coef(lm(mdf_mid_1s_hz ~ torque_mean_Nm))[2],

    .groups = "drop"
  )

ex_dyn_summary_table

```



# Results

## Torque Fatigue Profile


## RMS Evolution

## MDF Evolution

# Technical Limitations




