---
title: "EMG Fatigue Analysis Pipeline"
author: "Morgan Virolan"
date: "2026-04-17"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: cosmo


---

# Scientific question and Objectives
This is the companion notebook for the AMSTROM Campain, where we analyze the sEMG and torque data from the fatigue protocol. It's main objectives is (at this point) to quantifify the fatigue-induced changes in sEMG features (RMS, MDF) during the dynamic (EX_DYN) and static (EX_STA) fatigue sequences, and to see how they relate to the torque loss.
It's the first step of the broader study which will also include the analysis of other physiological signals recorded during the same experimental protocole (BIA, NIRS, Mecanical values) + stats & ML.
 
It's automatic, set the RUN_ID on the chunk below (must match existing subject) and hit "Run all". If doing a QC/verif check, the "Results" section at the end will print the plots to visualize the fatigue trends.

# Data and Preprocessing Framework
```{r setup, message=FALSE, warning=FALSE}
library(arrow)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(signal)   # butter(), filtfilt()
library(glue)


options(dplyr.summarise.inform = FALSE)
```

```{r load_data}

# Set RUN_ID to the participant folder name (format: XXXNnPp_YYYYMMDD)
RUN_ID <- "123DuPi_20251015"
export_root <- file.path("results", "DATA_EXPORT")
run_directory <- file.path(export_root, RUN_ID)

master_file_path      <- file.path(run_directory, "master_index_grid.parquet")
emg_file_path         <- file.path(run_directory, "preprocessed_emg.parquet")
torque_file_path      <- file.path(run_directory, "preprocessed_torque.parquet")
participant_file_path <- file.path(run_directory, "participant.parquet")

master_table <- read_parquet(
  master_file_path,
  col_select = c("time_index", "time_ref_s", "SEQ_index", "SEQ", "VC", "VC_count")
)
emg_table <- read_parquet(
  emg_file_path,
  col_select = c("time_index", "emg6", "emg8", "emg10")
)
torque_table <- read_parquet(
  torque_file_path,
  col_select = c("time_index", "torque_raw")
)
participant_table <- read_parquet(participant_file_path) |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, .before = 1)

participant_id <- participant_table$participant_id[[1]]

if (nrow(master_table) != nrow(emg_table)) {
  stop("master_table and emg_table do not have the same number of rows.")
}

if (!identical(master_table$time_index, emg_table$time_index)) {
  stop("master_table and emg_table are not aligned on time_index.")
}

analysis_table <- dplyr::bind_cols(
  master_table,
  emg_table |> dplyr::select(-time_index)
) |>
  left_join(torque_table, by = "time_index")

cat("RUN_ID:", RUN_ID, "\n")
cat("participant_id:", participant_id, "\n")
cat("nrow(master_table):", nrow(master_table), "\n")
cat("nrow(emg_table):   ", nrow(emg_table), "\n")
cat("nrow(torque_table):", nrow(torque_table), "\n")
cat("nrow(analysis_table):", nrow(analysis_table), "\n")

## master & emg must share the same hz, torque is expected to be about 1/5 of that and do contain n/a
analysis_table %>%
  summarise(
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est_hz = n() / duration_s,
    torque_na_frac = mean(is.na(torque_raw))
  )
```




```{r mvc_svc_load}
# MVC_REF = 3 to 4 reps of maximal voluntary contraction (MVC) at the start of the session
mvc_ref_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_REF",
              .data[["VC"]] == 1) |> # only keep actual reps (voluntary contraction = 1) / no rest
  dplyr::arrange(.data[["time_index"]])


#SVC_REF = submaximal voluntary contractions at 30/50/70% of MVC, used for calibration of torque and EMG
svc_ref_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "SVC_REF",
              .data[["VC"]] == 1) |> 
  dplyr::arrange(.data[["time_index"]])


#MVC_RECOV_DYN = MVC checkup after the dynamic fatigue sequence (EX_DYN) + 5 mins rest -  used to assess recovery oafter fatigue

mvc_recov_dyn_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_RECOV_DYN",
              .data[["VC"]] == 1) |> 
  dplyr::arrange(.data[["time_index"]])


#MVC_RECOV_STA = MVC checkup after the static fatigue sequence (EX_STA) + 5 mins rest -  used to assess recovery after fatigue

mvc_recov_sta_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_RECOV_STA",
              .data[["VC"]] == 1) |> 
  dplyr::arrange(.data[["time_index"]])

```


```{r ex_dyn_ex_sta_load}
## EX_DYN = dynamic fatigue sequence (62 reps, 3s efforts 2s recup, max effort each rep)
ex_dyn_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "EX_DYN",
                .data[["VC"]] == 1) |>
  dplyr::arrange(.data[["time_index"]])

## EX_STA = static fatigue sequence (1 rep, 60s effort, 70% of baseline mvc)
ex_sta_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "EX_STA",
                .data[["VC"]] == 1) |> #just in case sequencing is loose (everything should be VC==1)
  dplyr::arrange(.data[["time_index"]])

```
## Dataset Structure

## Torque Calibration
```{r torque_calibration}
#set values, from the biodex calibration
torque_offset_volt <- 0.106
torque_gain_Nm_per_volt <- 145.30

calc_torque_nm <- function(df) {
  df %>% dplyr::mutate(torque_Nm = torque_gain_Nm_per_volt * (torque_raw - torque_offset_volt))
}

ex_dyn_table       <- calc_torque_nm(ex_dyn_table)
ex_sta_table       <- calc_torque_nm(ex_sta_table)
mvc_ref_table      <- calc_torque_nm(mvc_ref_table)
svc_ref_table      <- calc_torque_nm(svc_ref_table)
mvc_recov_dyn_table <- calc_torque_nm(mvc_recov_dyn_table)
mvc_recov_sta_table <- calc_torque_nm(mvc_recov_sta_table)

```

```{r cleanup_analysis_table}
#remove the analysis table  - we extracted what we need for now
rm(analysis_table)
rm(emg_table)
rm(torque_table)
gc()
```

## Signal Processing 

```{r EMG_filtering}
# CONSTANTS
emg_sampling_frequency_hz <- 2148.149

emg_bandpass_low_hz <- 20
emg_bandpass_high_hz <- 450
emg_filter_order <- 4


# BANDPASS FILTER FUNCTION -----------------------------------
# clean artefacts, movement related noises, high freq noise.
emg_bandpass_filter <- function(emg_signal, sampling_frequency_hz, 
                                low_hz, high_hz, filter_order) {
  
  # Centers signal around zero by subtracting the mean.
  # This prevents a large "spike" at 0Hz in the frequency domain (dc offset removal)
  emg_centered <- emg_signal - mean(emg_signal, na.rm = TRUE)
  
  # Nyquist Normalization : cutoffs needed for digital filters
  # ratio of the Nyquist frequency (half the sampling rate).
  normalized_cutoffs <- c(low_hz, high_hz) / (sampling_frequency_hz / 2)
  
  # Butterworth filter 
  butter_filter <- signal::butter(filter_order, normalized_cutoffs, type = "pass")
  
  # Zero-Phase Filtering
  # filter twice (forward and backward) -> no time-shift (phase lag
  signal::filtfilt(butter_filter, emg_centered)
}


```

```{r apply_emg_filter}
# Apply the Butterworth bandpass filter to all EMG channels in every sequence table.
# Zero-phase (filtfilt) — no time shift. Applied after filter definition, before feature extraction.
emg_channels <- c("emg6", "emg8", "emg10")

apply_emg_filter <- function(df) {
  df |>
    dplyr::mutate(
      dplyr::across(
        dplyr::all_of(emg_channels),
        ~ emg_bandpass_filter(.x, emg_sampling_frequency_hz,
                              emg_bandpass_low_hz, emg_bandpass_high_hz, emg_filter_order)
      )
    )
}

ex_dyn_table        <- apply_emg_filter(ex_dyn_table)
ex_sta_table        <- apply_emg_filter(ex_sta_table)
mvc_ref_table       <- apply_emg_filter(mvc_ref_table)
svc_ref_table       <- apply_emg_filter(svc_ref_table)
mvc_recov_dyn_table <- apply_emg_filter(mvc_recov_dyn_table)
mvc_recov_sta_table <- apply_emg_filter(mvc_recov_sta_table)
```


# Definition of the Analysis Window

Adds markdown later to justify windows (comparable, avoid artefacts dues to ramps, enough hz to still run everything, etc)

## Contraction Labels


## Middle 1 s Window Extraction

```{r extract_middle_1s_window_function}
extract_middle_1s_window <- function(rep_df, window_length_s = 1.0) {

  rep_df |>
    dplyr::group_by(.data$VC_count) |>
    dplyr::mutate(
      t_mid = (min(.data$time_ref_s) + max(.data$time_ref_s)) / 2,
      mid_start = .data$t_mid - window_length_s / 2,
      mid_end   = .data$t_mid + window_length_s / 2
    ) |>
    dplyr::filter(
      .data$time_ref_s >= mid_start,
      .data$time_ref_s < mid_end
    ) |>
    dplyr::select(-.data$t_mid, -.data$mid_start, -.data$mid_end) |>
    dplyr::ungroup()
}
```

```{r build_1s_table}
#mvc_ref
mvc_ref_mid1s_table <- extract_middle_1s_window(mvc_ref_table)

mvc_ref_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

#svc_ref
svc_ref_mid1s_table <- extract_middle_1s_window(svc_ref_table)

svc_ref_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

#mvc_recov_dyn
mvc_recov_dyn_mid1s_table <- extract_middle_1s_window(mvc_recov_dyn_table)

mvc_recov_dyn_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

#mvc_recov_sta
mvc_recov_sta_mid1s_table <- extract_middle_1s_window(mvc_recov_sta_table)

mvc_recov_sta_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

#ex_dyn
ex_dyn_mid1s_table <- extract_middle_1s_window(ex_dyn_table)

ex_dyn_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

```



```{r extract_ex_sta_middle_1s_window}
# EX_STA window grid (1 s, non-overlap)

window_length_s <- 1.0 #must match
window_step_s   <- 1.0

t_start <- min(ex_sta_table$time_ref_s)
t_end   <- max(ex_sta_table$time_ref_s)

window_start_s <- seq(t_start, t_end - window_length_s, by = window_step_s)

ex_sta_window_table <- tibble::tibble(
  window_id      = seq_along(window_start_s),
  window_start_s = window_start_s,
  window_end_s   = window_start_s + window_length_s,
  window_mid_s   = window_start_s + window_length_s / 2
)

nrow(ex_sta_window_table)
ex_sta_window_table |> dplyr::slice(1:5)

## attach data

ex_sta_mid1s_table <- ex_sta_table |>
  dplyr::mutate(
    window_id = floor((time_ref_s - t_start) / window_step_s) + 1L
  ) |>
  dplyr::filter(
    window_id >= 1L,
    window_id <= nrow(ex_sta_window_table)
  ) |>
  dplyr::left_join(
    ex_sta_window_table,
    by = "window_id"
  ) |>
  dplyr::select(
    time_index, time_ref_s, SEQ_index, SEQ, VC, VC_count,
    window_id, window_start_s, window_end_s, window_mid_s,
    emg6, emg8, emg10, torque_raw, torque_Nm
  )

# quick check: should show ~1.0 s duration per window
ex_sta_mid1s_table |>
  dplyr::group_by(window_id) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    .groups = "drop"
  ) |>
  dplyr::slice(1:10) |>
  print()
```

# Feature Extraction

Markdown explications + features etc. 
All EMG features are computed on previously defined 1 s windows.


## EMG Amplitude Descriptor (RMS)



```{r RMS_function}
# RMS FUNCTION -----------------------------------
# Squares every point (-> everything becomes positive), finds average, takes the root -> power of the signal (same unit). amplitude/intensity of the contraction.
emg_rms <- function(emg_signal) {
  sqrt(mean(emg_signal^2, na.rm = TRUE))
}
```


## EMG Spectral Descriptor (Median Frequency)

markdown explications 

```{r MDF_function}
# MDF

emg_median_frequency_hz <- function(emg_signal, sampling_frequency_hz) {
  emg_centered <- emg_signal - mean(emg_signal) #recompute, last one is nested and global var is a dangerous world.
  n_samples <- length(emg_centered)
# Hann window to reduce spectral leakage 
  hann_window <- 0.5 - 0.5 * cos(2 * pi * (0:(n_samples - 1)) / (n_samples - 1))
  emg_windowed <- emg_centered * hann_window

# Single-sided power spectrum (FFT)
  fft_complex <- stats::fft(emg_windowed)
  n_half <- floor(n_samples / 2) + 1
  fft_half <- fft_complex[1:n_half]
  
# Power spectrum (magnitude squared of the FFT, normalized by the number of samples)
  power_spectrum <- (Mod(fft_half) ^ 2) / n_samples
  frequency_axis_hz <- (0:(n_half - 1)) * (sampling_frequency_hz / n_samples)
  
  #drop 0hz bin
  power_spectrum <- power_spectrum[-1]
frequency_axis_hz <- frequency_axis_hz[-1]

  # Cumulative power to find the median frequency
  cumulative_power <- cumsum(power_spectrum)
  median_index <- which(cumulative_power >= 0.5 * sum(power_spectrum))[1]
  
  frequency_axis_hz[median_index]
  }
```


# Construction of Feature Tables




## MVC Reference Features

```{r compute_feature_mvc_ref_mid1s_ref_table}
channels <- c("emg6", "emg8", "emg10")

mvc_ref_split <- split(mvc_ref_mid1s_table, mvc_ref_mid1s_table$VC_count)

mvc_ref_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_ref_split), function(vc_id) {

  block <- mvc_ref_split[[vc_id]]

  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_REF",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

print(mvc_ref_mid1s_feature_table)
```

```{r compute_feature_svc_ref_mid1s_table}
svc_ref_split <- split(svc_ref_mid1s_table, svc_ref_mid1s_table$VC_count)

svc_ref_mid1s_feature_table <- dplyr::bind_rows(lapply(names(svc_ref_split), function(vc_id) {

  block <- svc_ref_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "SVC_REF",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

print(svc_ref_mid1s_feature_table)
svc_ref_mid1s_feature_table |> dplyr::count(channel) |> print()
```

```{r compute_feature_mvc_recov_dyn_mid1s_table}
mvc_recov_dyn_split <- split(mvc_recov_dyn_mid1s_table, mvc_recov_dyn_mid1s_table$VC_count)

mvc_recov_dyn_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_recov_dyn_split), function(vc_id) {

  block <- mvc_recov_dyn_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_RECOV_DYN",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

print(mvc_recov_dyn_mid1s_feature_table)
```

```{r compute_feature_mvc_recov_sta_mid1s_table}
mvc_recov_sta_split <- split(mvc_recov_sta_mid1s_table, mvc_recov_sta_mid1s_table$VC_count)

mvc_recov_sta_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_recov_sta_split), function(vc_id) {

  block <- mvc_recov_sta_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_RECOV_STA",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

print(mvc_recov_sta_mid1s_feature_table)
```

```{r combine_feature_mid1s_tables}

reference_mid1s_feature_table <- dplyr::bind_rows(
  mvc_ref_mid1s_feature_table,
  svc_ref_mid1s_feature_table,
  mvc_recov_dyn_mid1s_feature_table,
  mvc_recov_sta_mid1s_feature_table
) |>
  dplyr::arrange(SEQ, channel, VC_count)

print(reference_mid1s_feature_table)
```

```{r mvc_level_summary}
mvc_level_summary <- reference_mid1s_feature_table |>
  dplyr::filter(SEQ %in% c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")) |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    n_reps = dplyr::n(),
    torque_mean_Nm = mean(torque_mean_Nm, na.rm = TRUE),
    rms_mid_1s_mean = mean(rms_mid_1s, na.rm = TRUE),
    mdf_mid_1s_hz_mean = mean(mdf_mid_1s_hz, na.rm = TRUE),
    .groups = "drop"
  )

mvc_level_summary

mvc_ref_baseline <- mvc_level_summary |>
  dplyr::filter(SEQ == "MVC_REF") |>
  dplyr::select(
    channel,
    torque_ref_Nm = torque_mean_Nm,
    rms_ref = rms_mid_1s_mean,
    mdf_ref_hz = mdf_mid_1s_hz_mean
  )

mvc_ref_baseline

# Shared torque baseline for downstream normalized torque analyses
torque_mvc_ref_baseline <- mean(mvc_ref_baseline$torque_ref_Nm, na.rm = TRUE)

torque_mvc_ref_baseline
```

## EX_DYN Feature Table

feature table per rep per channel

```{r build_ex_dyn_1mids_feature_table}

ex_dyn_split <- split(ex_dyn_mid1s_table, ex_dyn_mid1s_table$VC_count)

ex_dyn_mid1s_feature_table <- dplyr::bind_rows(lapply(names(ex_dyn_split), function(vc_id) {

  block <- ex_dyn_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "EX_DYN",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)

```


## EX_STA Feature Table

```{r compute_feature_ex_sta_mid1s_table}

ex_sta_split <- split(ex_sta_mid1s_table, ex_sta_mid1s_table$window_id)

ex_sta_mid1s_feature_table <- dplyr::bind_rows(lapply(names(ex_sta_split), function(win_id) {

  block <- ex_sta_split[[win_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "EX_STA",
      window_id = as.integer(win_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, window_id)

print(ex_sta_mid1s_feature_table)


```

```{r filter_ex_sta_onset_windows}
# Use the existing downstream 70% SVC reference torque as the EX_STA target source.
ex_sta_target_torque_Nm <- svc_ref_mid1s_feature_table |>
  dplyr::filter(VC_count == 3L) |>
  dplyr::summarise(target_torque_Nm = mean(torque_mean_Nm, na.rm = TRUE)) |>
  dplyr::pull(target_torque_Nm)

if (!is.finite(ex_sta_target_torque_Nm)) {
  stop("EX_STA onset filter failed: could not determine the 70% target torque from svc_ref_mid1s_feature_table.")
}

ex_sta_threshold_torque_Nm <- 0.95 * ex_sta_target_torque_Nm

ex_sta_window_torque_table <- ex_sta_mid1s_feature_table |>
  dplyr::distinct(window_id, torque_mean_Nm) |>
  dplyr::left_join(
    ex_sta_window_table |>
      dplyr::select(window_id, window_mid_s),
    by = "window_id"
  ) |>
  dplyr::arrange(window_id)

ex_sta_first_retained_window <- ex_sta_window_torque_table |>
  dplyr::filter(torque_mean_Nm >= ex_sta_threshold_torque_Nm) |>
  dplyr::slice(1)

if (nrow(ex_sta_first_retained_window) != 1) {
  stop(
    paste0(
      "EX_STA onset filter failed: no 1 s EX_STA window reached the threshold of ",
      round(ex_sta_threshold_torque_Nm, 3),
      " Nm."
    )
  )
}

ex_sta_first_retained_window_id <- ex_sta_first_retained_window$window_id[[1]]
ex_sta_first_retained_window_mid_s <- ex_sta_first_retained_window$window_mid_s[[1]]
ex_sta_number_of_excluded_initial_windows <- sum(
  ex_sta_window_torque_table$window_id < ex_sta_first_retained_window_id
)

ex_sta_onset_filter_log <- tibble::tibble(
  target_torque_used = ex_sta_target_torque_Nm,
  threshold_torque_used = ex_sta_threshold_torque_Nm,
  first_retained_window_id = ex_sta_first_retained_window_id,
  first_retained_window_mid_s = ex_sta_first_retained_window_mid_s,
  number_of_excluded_initial_windows = ex_sta_number_of_excluded_initial_windows
)

ex_sta_mid1s_feature_table <- ex_sta_mid1s_feature_table |>
  dplyr::filter(window_id >= ex_sta_first_retained_window_id)

print(ex_sta_onset_filter_log)
print(ex_sta_mid1s_feature_table)
```


# Baseline Normalization


## MVC Baseline Computation
```{r compute_mvc_ref_mid1s_rms_baseline_table}
mvc_ref_mid1s_rms_baseline_table <- mvc_ref_mid1s_feature_table |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    rms_mvc_ref_mid1s_baseline = mean(rms_mid_1s, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)

print(mvc_ref_mid1s_rms_baseline_table)
```


## RMS Normalization


normalize ex_dyn

```{r ex_dyn_mid1s_rms_normalization}
ex_dyn_mid1s_feature_table_norm <- ex_dyn_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm = rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    torque_norm = torque_mean_Nm / torque_mvc_ref_baseline
  )

print(ex_dyn_mid1s_feature_table_norm)
```


normalize ex_sta
```{r ex_sta_mid1s_rms_normalization}
ex_sta_mid1s_feature_table_norm <- ex_sta_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm = rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    torque_norm = torque_mean_Nm / torque_mvc_ref_baseline
  )

print(ex_sta_mid1s_feature_table_norm)
```

```{r mvc_recov_mid1s_rms_normalization}
mvc_recov_dyn_mid1s_feature_table_norm <- mvc_recov_dyn_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm         = rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    torque_norm      = torque_mean_Nm / torque_mvc_ref_baseline,
    emg_torque_ratio = rms_norm / torque_norm
  )

mvc_recov_sta_mid1s_feature_table_norm <- mvc_recov_sta_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm         = rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    torque_norm      = torque_mean_Nm / torque_mvc_ref_baseline,
    emg_torque_ratio = rms_norm / torque_norm
  )
```

```{r emg_torque_ratio_tables}
# Fatigue-oriented ratio: normalized EMG demand per unit normalized torque output
ex_dyn_emg_torque_ratio_table <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::mutate(
    emg_torque_ratio = rms_norm / torque_norm
  ) |>
  dplyr::select(
    SEQ, VC_count, channel,
    rms_norm, torque_norm, emg_torque_ratio
  )

ex_sta_emg_torque_ratio_table <- ex_sta_mid1s_feature_table_norm |>
  dplyr::left_join(
    ex_sta_window_table |>
      dplyr::select(window_id, window_mid_s),
    by = "window_id"
  ) |>
  dplyr::mutate(
    emg_torque_ratio = rms_norm / torque_norm
  ) |>
  dplyr::select(
    SEQ, window_id, window_mid_s, channel,
    rms_norm, torque_norm, emg_torque_ratio
  )

print(ex_dyn_emg_torque_ratio_table)
print(
  ex_sta_emg_torque_ratio_table |>
    dplyr::select(SEQ, window_mid_s, channel, rms_norm, torque_norm, emg_torque_ratio)
)
```

```{r emg_torque_ratio_slope_summary}
# Explicit per-channel linear slopes for the existing fatigue ratio tables.
ex_dyn_emg_torque_ratio_slope_summary_table <- ex_dyn_emg_torque_ratio_table |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    time_index_variable = "VC_count",
    fit_type = "lm(emg_torque_ratio ~ VC_count)",
    slope_estimate = unname(coef(lm(emg_torque_ratio ~ VC_count))[2]),
    intercept = unname(coef(lm(emg_torque_ratio ~ VC_count))[1]),
    r_squared = summary(lm(emg_torque_ratio ~ VC_count))$r.squared,
    n_points = dplyr::n(),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)

ex_sta_emg_torque_ratio_slope_summary_table <- ex_sta_emg_torque_ratio_table |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    time_index_variable = "window_mid_s",
    fit_type = "lm(emg_torque_ratio ~ window_mid_s)",
    slope_estimate = unname(coef(lm(emg_torque_ratio ~ window_mid_s))[2]),
    intercept = unname(coef(lm(emg_torque_ratio ~ window_mid_s))[1]),
    r_squared = summary(lm(emg_torque_ratio ~ window_mid_s))$r.squared,
    n_points = dplyr::n(),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)

emg_torque_ratio_slope_summary_table <- dplyr::bind_rows(
  ex_dyn_emg_torque_ratio_slope_summary_table,
  ex_sta_emg_torque_ratio_slope_summary_table
) |>
  dplyr::select(
    SEQ,
    channel,
    time_index_variable,
    slope_estimate,
    intercept,
    r_squared,
    n_points,
    fit_type
  ) |>
  dplyr::arrange(SEQ, channel)

print(emg_torque_ratio_slope_summary_table)
```


# Fatigue Trend Modeling

## MVC/SVC Trends

```{r build_reference_level_feature_table}
# Extend the existing 30/50/70 reference set with a 100% point from MVC_REF.
# The 100% row is an aggregated MVC_REF summary, not a true fourth SVC repetition.
# VC_count = 4L is kept only as a compatibility placeholder; plot/order logic should use reference_level_id.
reference_level_feature_table <- dplyr::bind_rows(
  svc_ref_mid1s_feature_table |>
    dplyr::mutate(
      reference_level_id = VC_count,
      reference_level = factor(
        VC_count,
        levels = c(1, 2, 3, 4),
        labels = c("30%", "50%", "70%", "100%")
      )
    ),
  mvc_ref_mid1s_feature_table |>
    dplyr::group_by(channel) |>
    dplyr::summarise(
      SEQ = "MVC_REF",
      VC_count = 4L,
      rms_mid_1s = mean(rms_mid_1s, na.rm = TRUE),
      mdf_mid_1s_hz = mean(mdf_mid_1s_hz, na.rm = TRUE),
      torque_mean_Nm = mean(torque_mean_Nm, na.rm = TRUE),
      n_samples = mean(n_samples, na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(
      reference_level_id = 4L,
      reference_level = factor(
        reference_level_id,
        levels = c(1, 2, 3, 4),
        labels = c("30%", "50%", "70%", "100%")
      )
    )
) |>
  dplyr::arrange(channel, reference_level_id)

print(reference_level_feature_table)
```

```{r build_reference_level_feature_table_norm}
# Keep the normalized reference workflow separate from the validated Nm-space table.
reference_level_feature_table_norm <- reference_level_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm = rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    torque_norm = torque_mean_Nm / torque_mvc_ref_baseline
  ) |>
  dplyr::select(
    channel,
    SEQ,
    VC_count,
    reference_level_id,
    reference_level,
    rms_mid_1s,
    mdf_mid_1s_hz,
    torque_mean_Nm,
    rms_mvc_ref_mid1s_baseline,
    rms_norm,
    torque_norm,
    n_samples
  ) |>
  dplyr::arrange(channel, reference_level_id)

print(reference_level_feature_table_norm)
```

```{r build_reference_calibration_table}
# Build a compact per-channel calibration fit from the unified 30/50/70/100 reference set.
# The fitted relationship maps calibrated torque to the expected channel-specific reference EMG feature.
reference_calibration_input_table <- reference_level_feature_table |>
  dplyr::select(
    channel,
    reference_level_id,
    reference_level,
    torque_mean_Nm,
    rms_mid_1s,
    mdf_mid_1s_hz
  ) |>
  tidyr::pivot_longer(
    cols = c(rms_mid_1s, mdf_mid_1s_hz),
    names_to = "response_variable",
    values_to = "response_value"
  ) |>
  dplyr::mutate(
    fit_type = "linear",
    predictor_variable = "torque_mean_Nm",
    predictor_definition = "Mean calibrated torque (Nm) from the middle 1 s reference window",
    response_definition = dplyr::case_when(
      response_variable == "rms_mid_1s" ~ "RMS computed on the middle 1 s reference window",
      response_variable == "mdf_mid_1s_hz" ~ "Median frequency (Hz) computed on the middle 1 s reference window",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::filter(
    is.finite(torque_mean_Nm),
    is.finite(response_value)
  ) |>
  dplyr::arrange(channel, response_variable, reference_level_id)

reference_calibration_table <- reference_calibration_input_table |>
  dplyr::group_by(
    channel,
    response_variable,
    fit_type,
    predictor_variable,
    predictor_definition,
    response_definition
  ) |>
  dplyr::summarise(
    n_reference_levels = dplyr::n(),
    slope = unname(coef(lm(response_value ~ torque_mean_Nm))[2]),
    intercept = unname(coef(lm(response_value ~ torque_mean_Nm))[1]),
    r_squared = summary(lm(response_value ~ torque_mean_Nm))$r.squared,
    .groups = "drop"
  ) |>
  dplyr::arrange(response_variable, channel)

print(reference_calibration_table)

reference_calibration_table_nm <- reference_calibration_table
```

```{r build_reference_calibration_table_norm}
# Build a separate normalized reference fit for RMS only: rms_norm ~ torque_norm.
reference_calibration_input_table_norm <- reference_level_feature_table_norm |>
  dplyr::select(
    channel,
    reference_level_id,
    reference_level,
    torque_norm,
    rms_norm
  ) |>
  dplyr::mutate(
    response_variable = "rms_norm",
    fit_type = "linear",
    predictor_variable = "torque_norm",
    predictor_definition = "Mean normalized torque from the middle 1 s reference window (MVC_REF = 1)",
    response_definition = "Normalized RMS from the middle 1 s reference window (MVC_REF = 1)"
  ) |>
  dplyr::filter(
    is.finite(torque_norm),
    is.finite(rms_norm)
  ) |>
  dplyr::arrange(channel, reference_level_id)

reference_calibration_table_norm <- reference_calibration_input_table_norm |>
  dplyr::group_by(
    channel,
    response_variable,
    fit_type,
    predictor_variable,
    predictor_definition,
    response_definition
  ) |>
  dplyr::summarise(
    n_reference_levels = dplyr::n(),
    slope = unname(coef(lm(rms_norm ~ torque_norm))[2]),
    intercept = unname(coef(lm(rms_norm ~ torque_norm))[1]),
    r_squared = summary(lm(rms_norm ~ torque_norm))$r.squared,
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)

print(reference_calibration_table_norm)
```

```{r augment_ratio_tables_with_reference_expected}
# Add reference-derived expected ratio to both fatigue ratio tables.
# expected_ratio = (intercept + slope * torque_norm) / torque_norm
# This is what the ratio WOULD be for a fresh subject at each observed torque level,
# derived from the pre-fatigue calibration (rms_norm ~ torque_norm from SVC/MVC reference).
reference_norm_coefs <- reference_calibration_table_norm |>
  dplyr::select(channel, slope_ref = slope, intercept_ref = intercept)

ex_dyn_emg_torque_ratio_table <- ex_dyn_emg_torque_ratio_table |>
  dplyr::left_join(reference_norm_coefs, by = "channel") |>
  dplyr::mutate(
    expected_ratio = (intercept_ref + slope_ref * torque_norm) / torque_norm
  ) |>
  dplyr::select(-slope_ref, -intercept_ref)

ex_sta_emg_torque_ratio_table <- ex_sta_emg_torque_ratio_table |>
  dplyr::left_join(reference_norm_coefs, by = "channel") |>
  dplyr::mutate(
    expected_ratio = (intercept_ref + slope_ref * torque_norm) / torque_norm
  ) |>
  dplyr::select(-slope_ref, -intercept_ref)
```

```{r svc_ref_vc_plots_RMS_MF}
svc_df <- reference_level_feature_table |>
  dplyr::mutate(
    reference_level_id = as.integer(reference_level_id),
    intensity = factor(reference_level,
                       levels = c("30%", "50%", "70%", "100%")),
    torque_Nm = as.numeric(torque_mean_Nm)
  ) |>
  dplyr::filter(
    is.finite(reference_level_id),
    is.finite(torque_Nm),
    is.finite(rms_mid_1s),
    is.finite(mdf_mid_1s_hz)
  )

p_mdf <- ggplot(svc_df, aes(x = intensity, y = mdf_mid_1s_hz, group = 1)) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "Reference set - MDF (middle 1s)",
    x = "Reference level (% MVC)",
    y = "MDF (Hz)"
  )

p_mdf

p_rms <- ggplot(svc_df, aes(x = intensity, y = rms_mid_1s, group = 1)) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "Reference set - RMS across intensities (middle 1s)",
    x = "Reference level (% MVC)",
    y = "RMS (a.u.)"
  )

p_rms
```

```{r percent_emg_vs_percent_force_svc_ref}
# %EMG and %Force ()
svc_pct_df <- reference_level_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    reference_level_id = as.integer(reference_level_id),
    torque_pct = 100 * torque_mean_Nm / torque_mvc_ref_baseline,
    emg_pct    = 100 * rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    intensity  = factor(reference_level, levels = c("30%", "50%", "70%", "100%"))
  ) |>
  dplyr::filter(is.finite(reference_level_id))

# %EMG vs %Force 
percent_emg_vs_percent_force_plot <- ggplot(svc_pct_df, aes(x = torque_pct, y = emg_pct)) +
  geom_point(aes(shape = intensity), size = 2) +
  geom_line(aes(group = 1)) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "Reference set - %EMG vs %Force (normalized to MVC_REF)",
    x = "% Force (Torque / MVC_REF torque)",
    y = "% EMG (RMS / MVC_REF RMS)"
  )

percent_emg_vs_percent_force_plot
```

```{r force_vs_emg_%}
svc_pct_summary_table <- svc_pct_df |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    `Slope (%EMG / %Force)` = coef(lm(emg_pct ~ torque_pct))[2],
    `Pearson r (emg_pct, torque_pct)` = cor(emg_pct, torque_pct, use = "complete.obs"),
    `R² (emg_pct ~ torque_pct)` = summary(lm(emg_pct ~ torque_pct))$r.squared,
    .groups = "drop"
  )

svc_pct_summary_table
```

```{r build_ex_dyn_reference_rms_comparison_table}
# Compare observed EX_DYN RMS against the validated per-channel reference fit in torque_mean_Nm space.
ex_dyn_reference_rms_fit_table <- reference_calibration_table |>
  dplyr::filter(response_variable == "rms_mid_1s") |>
  dplyr::select(
    channel,
    response_variable,
    predictor_variable,
    fit_type,
    slope,
    intercept,
    r_squared
  )

ex_dyn_reference_rms_comparison_table <- ex_dyn_mid1s_feature_table |>
  dplyr::select(
    SEQ,
    VC_count,
    channel,
    torque_mean_Nm,
    rms_mid_1s
  ) |>
  dplyr::left_join(
    ex_dyn_reference_rms_fit_table,
    by = "channel"
  ) |>
  dplyr::mutate(
    observed_rms_mid_1s = rms_mid_1s,
    expected_rms_mid_1s = intercept + slope * torque_mean_Nm,
    residual_rms_mid_1s = observed_rms_mid_1s - expected_rms_mid_1s,
    relative_residual_rms = dplyr::if_else(
      is.finite(expected_rms_mid_1s) & expected_rms_mid_1s != 0,
      residual_rms_mid_1s / expected_rms_mid_1s,
      NA_real_
    )
  ) |>
  dplyr::select(
    SEQ,
    VC_count,
    channel,
    torque_mean_Nm,
    response_variable,
    predictor_variable,
    fit_type,
    observed_rms_mid_1s,
    expected_rms_mid_1s,
    residual_rms_mid_1s,
    relative_residual_rms,
    r_squared
  ) |>
  dplyr::arrange(channel, VC_count)

print(ex_dyn_reference_rms_comparison_table)

ex_dyn_reference_rms_fit_table_nm <- ex_dyn_reference_rms_fit_table
ex_dyn_reference_rms_comparison_table_nm <- ex_dyn_reference_rms_comparison_table
```

```{r build_ex_dyn_reference_rms_comparison_table_norm}
# Project the fixed normalized reference fit onto EX_DYN fatigue points without refitting.
ex_dyn_reference_rms_fit_table_norm <- reference_calibration_table_norm |>
  dplyr::select(
    channel,
    response_variable,
    predictor_variable,
    fit_type,
    slope,
    intercept,
    r_squared
  )

ex_dyn_reference_rms_comparison_table_norm <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::select(
    SEQ,
    VC_count,
    channel,
    torque_norm,
    rms_norm
  ) |>
  dplyr::left_join(
    ex_dyn_reference_rms_fit_table_norm,
    by = "channel"
  ) |>
  dplyr::mutate(
    observed_rms_norm = rms_norm,
    expected_rms_norm = intercept + slope * torque_norm,
    residual_rms_norm = observed_rms_norm - expected_rms_norm,
    relative_residual_rms_norm = dplyr::if_else(
      is.finite(expected_rms_norm) & expected_rms_norm != 0,
      residual_rms_norm / expected_rms_norm,
      NA_real_
    )
  ) |>
  dplyr::select(
    SEQ,
    VC_count,
    channel,
    torque_norm,
    response_variable,
    predictor_variable,
    fit_type,
    observed_rms_norm,
    expected_rms_norm,
    residual_rms_norm,
    relative_residual_rms_norm,
    r_squared
  ) |>
  dplyr::arrange(channel, VC_count)

print(ex_dyn_reference_rms_comparison_table_norm)

ex_dyn_reference_projection_table_norm <- ex_dyn_reference_rms_comparison_table_norm
```

```{r plot_ex_dyn_reference_rms_overlay}
reference_rms_points_plot_df <- reference_level_feature_table |>
  dplyr::select(channel, reference_level, torque_mean_Nm, rms_mid_1s)

ex_dyn_reference_rms_overlay_plot <- ggplot() +
  geom_abline(
    data = ex_dyn_reference_rms_fit_table,
    aes(intercept = intercept, slope = slope),
    color = "#4d4d4d",
    linewidth = 0.9,
    alpha = 0.9
  ) +
  geom_point(
    data = reference_rms_points_plot_df,
    aes(x = torque_mean_Nm, y = rms_mid_1s),
    color = "#a6a6a6",
    size = 2.1,
    alpha = 0.9
  ) +
  geom_point(
    data = ex_dyn_reference_rms_comparison_table,
    aes(x = torque_mean_Nm, y = observed_rms_mid_1s, color = VC_count),
    size = 1.7,
    alpha = 0.8
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  scale_color_gradient(low = "#1b6ca8", high = "#c0392b", name = "Repetition") +
  labs(
    title = "EX_DYN RMS vs subject's reference fit",
    subtitle = "Grey: reference fit & levels.  Colour: repetition order (early \u2192 late).",
    x = "Torque mean (Nm)",
    y = "RMS (a.u.)"
  ) +
  theme(legend.position = "top")

print(ex_dyn_reference_rms_overlay_plot)

ex_dyn_reference_rms_overlay_plot_nm <- ex_dyn_reference_rms_overlay_plot
```

```{r plot_ex_dyn_reference_rms_overlay_norm}
reference_rms_points_plot_df_norm <- reference_level_feature_table_norm |>
  dplyr::select(channel, reference_level, torque_norm, rms_norm)

ref_rms_ci_df <- reference_level_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit    <- lm(rms_norm ~ torque_norm, data = .x)
    x_grid <- tibble::tibble(torque_norm = seq(min(.x$torque_norm), max(.x$torque_norm), length.out = 100))
    pred   <- predict(fit, newdata = x_grid, interval = "confidence", level = 0.95)
    dplyr::bind_cols(x_grid, tibble::as_tibble(pred))
  }) |>
  dplyr::ungroup()

ex_dyn_reference_rms_overlay_plot_norm <- ggplot() +
  geom_ribbon(
    data = ref_rms_ci_df,
    aes(x = torque_norm, ymin = lwr, ymax = upr),
    fill = "#444444", alpha = 0.12
  ) +
  geom_line(
    data = ref_rms_ci_df,
    aes(x = torque_norm, y = fit),
    colour = "#444444", linewidth = 1.0, alpha = 0.9
  ) +
  geom_point(
    data = reference_rms_points_plot_df_norm,
    aes(x = torque_norm, y = rms_norm),
    color = "#aaaaaa",
    size = 2.1,
    alpha = 0.9
  ) +
  geom_point(
    data = ex_dyn_reference_rms_comparison_table_norm,
    aes(x = torque_norm, y = observed_rms_norm, color = VC_count),
    size = 1.7,
    alpha = 0.8
  ) +
  geom_point(
    data = mvc_recov_dyn_mid1s_feature_table_norm,
    aes(x = torque_norm, y = rms_norm),
    shape = 8, size = 2.8, stroke = 1.1, color = "#e6ab02",
    inherit.aes = FALSE
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  scale_color_gradient(low = "#1b6ca8", high = "#c0392b", name = "Repetition") +
  labs(
    title = "EX_DYN normalized RMS vs normalized ref fit",
    subtitle = "Grey: reference levels, fit, and 95% CI.  Colour: rep order.  \u2605 = MVC_RECOV_DYN.",
    x = "Normalized torque (MVC_REF = 1)",
    y = "Normalized RMS (MVC_REF = 1)"
  ) +
  theme(legend.position = "top")

print(ex_dyn_reference_rms_overlay_plot_norm)
```

```{r ex_dyn_reference_residual_plot_norm}
ex_dyn_reference_residual_plot_norm <- ggplot(
  ex_dyn_reference_projection_table_norm,
  aes(x = VC_count, y = residual_rms_norm)
) +
  geom_hline(yintercept = 0, linewidth = 0.7, color = "#4d4d4d", linetype = "dashed") +
  geom_point(alpha = 0.8, size = 1.5, color = "#1b6ca8") +
  geom_line(linewidth = 0.7, color = "#1b6ca8") +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "EX_DYN - Normalized RMS residual from fixed normalized reference across repetitions",
    subtitle = "Residual = observed rms_norm - expected rms_norm from the fixed reference fit (rms_norm ~ torque_norm)",
    x = "Repetition",
    y = "Residual normalized RMS (MVC_REF = 1)"
  )

print(ex_dyn_reference_residual_plot_norm)
```

```{r ex_dyn_reference_observed_expected_over_repetition_plot_norm}
ex_dyn_reference_observed_expected_over_repetition_plot_norm_data <- ex_dyn_reference_projection_table_norm |>
  dplyr::select(channel, VC_count, observed_rms_norm, expected_rms_norm) |>
  tidyr::pivot_longer(
    cols = c(observed_rms_norm, expected_rms_norm),
    names_to = "series",
    values_to = "rms_norm_value"
  ) |>
  dplyr::mutate(
    series = dplyr::recode(
      series,
      observed_rms_norm = "Observed normalized RMS",
      expected_rms_norm = "Expected normalized RMS from fixed reference"
    ),
    series = factor(
      series,
      levels = c(
        "Observed normalized RMS",
        "Expected normalized RMS from fixed reference"
      )
    )
  )

ex_dyn_reference_observed_expected_over_repetition_plot_norm <- ggplot(
  ex_dyn_reference_observed_expected_over_repetition_plot_norm_data,
  aes(x = VC_count, y = rms_norm_value, color = series)
) +
  geom_point(alpha = 0.8, size = 1.4) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ channel, scales = "free_y") +
  scale_color_manual(
    values = c(
      "Observed normalized RMS" = "#1b6ca8",
      "Expected normalized RMS from fixed reference" = "#4d4d4d"
    )
  ) +
  labs(
    title = "EX_DYN - Observed and expected normalized RMS across repetitions",
    subtitle = "Expected values are projected from the fixed normalized reference fit; the fatigue data are not re-fit",
    x = "Repetition",
    y = "Normalized RMS (MVC_REF = 1)",
    color = NULL
  ) +
  theme(legend.position = "top")

print(ex_dyn_reference_observed_expected_over_repetition_plot_norm)
```

```{r build_ex_dyn_reference_mdf_comparison_table}
# Compare observed EX_DYN MDF against the validated per-channel reference fit in torque_mean_Nm space.
# Mirrors the RMS residual analysis: isolates true spectral fatigue from the torque-loss artifact.
ex_dyn_reference_mdf_fit_table <- reference_calibration_table |>
  dplyr::filter(response_variable == "mdf_mid_1s_hz") |>
  dplyr::select(
    channel,
    response_variable,
    predictor_variable,
    fit_type,
    slope,
    intercept,
    r_squared
  )

ex_dyn_reference_mdf_comparison_table <- ex_dyn_mid1s_feature_table |>
  dplyr::select(
    SEQ,
    VC_count,
    channel,
    torque_mean_Nm,
    mdf_mid_1s_hz
  ) |>
  dplyr::left_join(
    ex_dyn_reference_mdf_fit_table,
    by = "channel"
  ) |>
  dplyr::mutate(
    observed_mdf_mid_1s_hz = mdf_mid_1s_hz,
    expected_mdf_mid_1s_hz = intercept + slope * torque_mean_Nm,
    residual_mdf_mid_1s_hz = observed_mdf_mid_1s_hz - expected_mdf_mid_1s_hz,
    relative_residual_mdf = dplyr::if_else(
      is.finite(expected_mdf_mid_1s_hz) & expected_mdf_mid_1s_hz != 0,
      residual_mdf_mid_1s_hz / expected_mdf_mid_1s_hz,
      NA_real_
    )
  ) |>
  dplyr::select(
    SEQ,
    VC_count,
    channel,
    torque_mean_Nm,
    response_variable,
    predictor_variable,
    fit_type,
    observed_mdf_mid_1s_hz,
    expected_mdf_mid_1s_hz,
    residual_mdf_mid_1s_hz,
    relative_residual_mdf,
    r_squared
  ) |>
  dplyr::arrange(channel, VC_count)

print(ex_dyn_reference_mdf_comparison_table)
```

```{r plot_ex_dyn_reference_mdf_overlay}
reference_mdf_points_plot_df <- reference_level_feature_table |>
  dplyr::select(channel, reference_level, torque_mean_Nm, mdf_mid_1s_hz)

ex_dyn_reference_mdf_overlay_plot <- ggplot() +
  geom_abline(
    data = ex_dyn_reference_mdf_fit_table,
    aes(intercept = intercept, slope = slope),
    color = "#4d4d4d",
    linewidth = 0.9,
    alpha = 0.9
  ) +
  geom_point(
    data = reference_mdf_points_plot_df,
    aes(x = torque_mean_Nm, y = mdf_mid_1s_hz),
    color = "#a6a6a6",
    size = 2.1,
    alpha = 0.9
  ) +
  geom_point(
    data = ex_dyn_reference_mdf_comparison_table,
    aes(x = torque_mean_Nm, y = observed_mdf_mid_1s_hz, color = VC_count),
    size = 1.7,
    alpha = 0.8
  ) +
  facet_wrap(~ channel, scales = "free_y") +
  scale_color_gradient(low = "#1b6ca8", high = "#c0392b", name = "Repetition") +
  labs(
    title = "EX_DYN MDF vs subject's reference fit",
    subtitle = "Grey: reference fit & levels.  Colour: repetition order (early \u2192 late).",
    x = "Torque mean (Nm)",
    y = "MDF (Hz)"
  ) +
  theme(legend.position = "top")

print(ex_dyn_reference_mdf_overlay_plot)
```

```{r ex_dyn_reference_mdf_residual_plot}
ex_dyn_reference_mdf_residual_plot <- ggplot(
  ex_dyn_reference_mdf_comparison_table,
  aes(x = VC_count, y = residual_mdf_mid_1s_hz)
) +
  geom_hline(yintercept = 0, linewidth = 0.7, color = "#4d4d4d", linetype = "dashed") +
  geom_point(alpha = 0.8, size = 1.5, color = "#d17a22") +
  geom_line(linewidth = 0.7, color = "#d17a22") +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "EX_DYN - MDF residual from fixed reference across repetitions",
    subtitle = "Residual = observed MDF - expected MDF from the fixed reference fit (mdf_mid_1s_hz ~ torque_mean_Nm)",
    x = "Repetition",
    y = "Residual MDF (Hz)"
  )

print(ex_dyn_reference_mdf_residual_plot)
```

```{r ex_dyn_reference_mdf_observed_expected_over_repetition_plot}
ex_dyn_reference_mdf_observed_expected_data <- ex_dyn_reference_mdf_comparison_table |>
  dplyr::select(channel, VC_count, observed_mdf_mid_1s_hz, expected_mdf_mid_1s_hz) |>
  tidyr::pivot_longer(
    cols = c(observed_mdf_mid_1s_hz, expected_mdf_mid_1s_hz),
    names_to = "series",
    values_to = "mdf_hz_value"
  ) |>
  dplyr::mutate(
    series = dplyr::recode(
      series,
      observed_mdf_mid_1s_hz = "Observed MDF",
      expected_mdf_mid_1s_hz = "Expected MDF from fixed reference"
    ),
    series = factor(
      series,
      levels = c("Observed MDF", "Expected MDF from fixed reference")
    )
  )

ex_dyn_reference_mdf_observed_expected_plot <- ggplot(
  ex_dyn_reference_mdf_observed_expected_data,
  aes(x = VC_count, y = mdf_hz_value, color = series)
) +
  geom_point(alpha = 0.8, size = 1.4) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ channel, scales = "free_y") +
  scale_color_manual(
    values = c(
      "Observed MDF" = "#d17a22",
      "Expected MDF from fixed reference" = "#4d4d4d"
    )
  ) +
  labs(
    title = "EX_DYN - Observed and expected MDF across repetitions",
    subtitle = "Expected values are projected from the fixed reference fit; the fatigue data are not re-fit",
    x = "Repetition",
    y = "MDF (Hz)",
    color = NULL
  ) +
  theme(legend.position = "top")

print(ex_dyn_reference_mdf_observed_expected_plot)
```

```{r mvc_torque_plot_summary}
mvc_torque_plot_df <- mvc_level_summary |>
  dplyr::distinct(SEQ, torque_mean_Nm) |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

mvc_torque_summary_plot <- ggplot(mvc_torque_plot_df, aes(x = SEQ, y = torque_mean_Nm)) +
  geom_point(size = 3) +
  geom_line(aes(group = 1)) +
  labs(
    title = "MVC — Torque baseline vs recovery",
    x = NULL,
    y = "Torque (Nm)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

mvc_torque_summary_plot
```

```{r mvc_recov_plot_summary}
mvc_rms_plot_df <- mvc_level_summary |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

mvc_rms_summary_plot <- ggplot(mvc_rms_plot_df, aes(x = SEQ, y = rms_mid_1s_mean)) +
  geom_point(size = 2) +
  geom_line(aes(group = channel)) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "MVC — RMS baseline vs recovery (middle 1s)",
    x = NULL,
    y = "RMS (a.u.)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

mvc_rms_summary_plot
```

```{r mvc_mdf_plot_summary}
mvc_mdf_plot_df <- mvc_level_summary |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

mvc_mdf_summary_plot <- ggplot(mvc_mdf_plot_df, aes(x = SEQ, y = mdf_mid_1s_hz_mean)) +
  geom_point(size = 2) +
  geom_line(aes(group = channel)) +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "MVC — MDF baseline vs recovery (middle 1s)",
    x = NULL,
    y = "MDF (Hz)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

mvc_mdf_summary_plot
```

```{r mvc_dyn_recovery_plot}
mvc_dyn_recovery_df <- mvc_level_summary |>
  dplyr::select(SEQ, channel, n_reps, torque_mean_Nm, rms_mid_1s_mean, mdf_mid_1s_hz_mean) |>
  dplyr::filter(SEQ %in% c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")) |>
  dplyr::mutate(
    SEQ = factor(SEQ, levels = c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

# build table
plot_long <- mvc_dyn_recovery_df |>
  dplyr::group_by(SEQ) |>
  dplyr::mutate(torque_mean_Nm = mean(torque_mean_Nm)) |>
  dplyr::ungroup() |>
  tidyr::pivot_longer(
    cols = c(torque_mean_Nm, rms_mid_1s_mean, mdf_mid_1s_hz_mean),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_mean_Nm = "Torque (Nm)",
      rms_mid_1s_mean = "RMS (a.u.)",
      mdf_mid_1s_hz_mean = "MDF (Hz)"
    ),
    metric = factor(metric, levels = c("Torque (Nm)", "RMS (a.u.)", "MDF (Hz)"))
  ) |>
  dplyr::filter(is.finite(value))

# Plot
mvc_recovery_combined_plot <- ggplot(plot_long, aes(x = SEQ, y = value)) +
  geom_point(size = 2) +
  geom_line(aes(group = channel)) +
  facet_grid(metric ~ channel, scales = "free_y") +
  scale_x_discrete(
    labels = c(
      MVC_REF = "MVC",
      MVC_RECOV_DYN = "RECOV dyn",
      MVC_RECOV_STA = "RECOV sta"
    )
  ) +
  labs(
    title = "REF -> R_DYN -> R_STA",
    x = NULL,
    y = NULL
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

mvc_recovery_combined_plot
```




## EX_DYN Trends

```{r ex_dyn_torque_vs_rep}
ex_dyn_torque_vs_rep_plot <- ggplot(ex_dyn_mid1s_feature_table_norm,
       aes(x = VC_count, y = torque_norm)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - Normalized torque across repetitions",
    x = "Repetition",
    y = "Normalized torque"
  )

ex_dyn_torque_vs_rep_plot
```


```{r ex_dyn_rms_vs_torque}
# CI ribbon and fit line from the pre-fatigue reference calibration (ref_rms_ci_df, already computed above).
ex_dyn_rms_vs_torque_plot <- ggplot() +
  geom_ribbon(
    data = ref_rms_ci_df,
    aes(x = torque_norm, ymin = lwr, ymax = upr),
    fill = "#888888", alpha = 0.12
  ) +
  geom_line(
    data = ref_rms_ci_df,
    aes(x = torque_norm, y = fit),
    colour = "#888888", linewidth = 1.0
  ) +
  geom_point(
    data = ex_dyn_mid1s_feature_table_norm,
    aes(x = torque_norm, y = rms_norm, color = VC_count),
    alpha = 0.7, size = 1.4
  ) +
  scale_color_gradient(low = "#1b6ca8", high = "#c0392b", name = "Repetition") +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - EMG activation vs normalized torque",
    subtitle = "Grey line + ribbon = reference calibration fit (95% CI); colour = repetition order",
    x = "Normalized torque (MVC_REF = 1)",
    y = "Normalized RMS"
  )

ex_dyn_rms_vs_torque_plot
```  


```{r ex_dyn_mdf_vs_rep}
ex_dyn_mdf_rep_ci_df <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit    <- lm(mdf_mid_1s_hz ~ VC_count, data = .x)
    x_grid <- tibble::tibble(VC_count = seq(min(.x$VC_count), max(.x$VC_count)))
    pred   <- predict(fit, newdata = x_grid, interval = "confidence", level = 0.95)
    dplyr::bind_cols(x_grid, tibble::as_tibble(pred))
  }) |>
  dplyr::ungroup()

ex_dyn_mdf_vs_rep_plot <- ggplot() +
  geom_ribbon(
    data = ex_dyn_mdf_rep_ci_df,
    aes(x = VC_count, ymin = lwr, ymax = upr),
    fill = "#d17a22", alpha = 0.15
  ) +
  geom_line(
    data = ex_dyn_mdf_rep_ci_df,
    aes(x = VC_count, y = fit),
    colour = "#d17a22", linewidth = 1.0
  ) +
  geom_point(
    data = ex_dyn_mid1s_feature_table_norm,
    aes(x = VC_count, y = mdf_mid_1s_hz),
    alpha = 0.7, size = 1.4
  ) +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - Median frequency across repetitions",
    subtitle = "Line + ribbon = linear fit on observed data (95% CI)",
    x = "Repetition",
    y = "Median frequency (Hz)"
  )

ex_dyn_mdf_vs_rep_plot
```

```{r ex_dyn_mdf_vs_torque}
ex_dyn_mdf_torque_ci_df <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit    <- lm(mdf_mid_1s_hz ~ torque_norm, data = .x)
    x_grid <- tibble::tibble(torque_norm = seq(min(.x$torque_norm), max(.x$torque_norm), length.out = 100))
    pred   <- predict(fit, newdata = x_grid, interval = "confidence", level = 0.95)
    dplyr::bind_cols(x_grid, tibble::as_tibble(pred))
  }) |>
  dplyr::ungroup()

ex_dyn_mdf_vs_torque_plot <- ggplot() +
  geom_ribbon(
    data = ex_dyn_mdf_torque_ci_df,
    aes(x = torque_norm, ymin = lwr, ymax = upr),
    fill = "#d17a22", alpha = 0.15
  ) +
  geom_line(
    data = ex_dyn_mdf_torque_ci_df,
    aes(x = torque_norm, y = fit),
    colour = "#d17a22", linewidth = 1.0
  ) +
  geom_point(
    data = ex_dyn_mid1s_feature_table_norm,
    aes(x = torque_norm, y = mdf_mid_1s_hz),
    alpha = 0.7, size = 1.4
  ) +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - Median frequency vs normalized torque",
    subtitle = "Line + ribbon = linear fit on observed data (95% CI)",
    x = "Normalized torque (MVC_REF = 1)",
    y = "Median frequency (Hz)"
  )

ex_dyn_mdf_vs_torque_plot
```  

```{r ex_dyn_dual_axis_rms_mdf_vs_torque}
ex_dyn_dual_axis_params <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    rms_min = min(rms_norm, na.rm = TRUE),
    rms_max = max(rms_norm, na.rm = TRUE),
    mdf_min = min(mdf_mid_1s_hz, na.rm = TRUE),
    mdf_max = max(mdf_mid_1s_hz, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    rms_span = rms_max - rms_min,
    mdf_span = mdf_max - mdf_min,
    scale_factor = dplyr::if_else(rms_span > 0 & mdf_span > 0, rms_span / mdf_span, 1)
  )

ex_dyn_dual_axis_plot_data <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::left_join(ex_dyn_dual_axis_params, by = "channel") |>
  dplyr::mutate(
    mdf_scaled = (mdf_mid_1s_hz - mdf_min) * scale_factor + rms_min
  )

ex_dyn_dual_axis_plots <- lapply(unique(ex_dyn_dual_axis_plot_data$channel), function(ch) {
  plot_df <- ex_dyn_dual_axis_plot_data |>
    dplyr::filter(channel == ch)

  rms_min <- plot_df$rms_min[1]
  rms_max <- plot_df$rms_max[1]
  mdf_min <- plot_df$mdf_min[1]
  scale_factor <- plot_df$scale_factor[1]

  ggplot(plot_df, aes(x = torque_norm)) +
    geom_point(aes(y = rms_norm, color = "RMS (normalized)"), alpha = 0.8, size = 1.8) +
    geom_line(aes(y = rms_norm, color = "RMS (normalized)"), linewidth = 0.7) +
    geom_point(aes(y = mdf_scaled, color = "MDF (Hz)"), alpha = 0.8, size = 1.8, shape = 17) +
    geom_line(aes(y = mdf_scaled, color = "MDF (Hz)"), linewidth = 0.7) +
    scale_y_continuous(
      name = "RMS (normalized)",
      sec.axis = sec_axis(
        transform = ~ (. - rms_min) / scale_factor + mdf_min,
        name = "MDF (Hz)"
      )
    ) +
    scale_color_manual(
      values = c("RMS (normalized)" = "#1b6ca8", "MDF (Hz)" = "#d17a22"),
      breaks = c("RMS (normalized)", "MDF (Hz)")
    ) +
    labs(
      title = paste("EX_DYN - RMS and MDF vs normalized torque", ch),
      subtitle = "(Right axis = rescaled)",
      x = "Normalized torque",
      color = NULL
    ) +
    coord_cartesian(ylim = c(rms_min, rms_max)) +
    theme(legend.position = "top")
})

names(ex_dyn_dual_axis_plots) <- unique(ex_dyn_dual_axis_plot_data$channel)

for (plot_obj in ex_dyn_dual_axis_plots) {
  print(plot_obj)
}
```

```{r ex_dyn_rms_vs_repetition}
ex_dyn_rms_rep_ci_df <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit    <- lm(rms_norm ~ VC_count, data = .x)
    x_grid <- tibble::tibble(VC_count = seq(min(.x$VC_count), max(.x$VC_count)))
    pred   <- predict(fit, newdata = x_grid, interval = "confidence", level = 0.95)
    dplyr::bind_cols(x_grid, tibble::as_tibble(pred))
  }) |>
  dplyr::ungroup()

ex_dyn_rms_vs_repetition_plot <- ggplot() +
  geom_ribbon(
    data = ex_dyn_rms_rep_ci_df,
    aes(x = VC_count, ymin = lwr, ymax = upr),
    fill = "#1b6ca8", alpha = 0.15
  ) +
  geom_line(
    data = ex_dyn_rms_rep_ci_df,
    aes(x = VC_count, y = fit),
    colour = "#1b6ca8", linewidth = 1.0
  ) +
  geom_point(
    data = ex_dyn_mid1s_feature_table_norm,
    aes(x = VC_count, y = rms_norm),
    alpha = 0.7, size = 1.4
  ) +
  facet_wrap(~ channel) +
  labs(
    title = "EX_DYN - Normalized RMS across repetitions",
    subtitle = "Line + ribbon = linear fit on observed data (95% CI)",
    x = "Repetition",
    y = "Normalized RMS"
  )

ex_dyn_rms_vs_repetition_plot
```

```{r ex_dyn_emg_torque_ratio_vs_rep}
ex_dyn_emg_torque_ratio_trend_plot <- ggplot(
  ex_dyn_emg_torque_ratio_table,
  aes(x = VC_count, y = emg_torque_ratio)
) +
  geom_point(alpha = 0.7, size = 1.2, color = "#1b6ca8") +
  geom_line(linewidth = 0.7, color = "#1b6ca8") +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "EX_DYN - EMG/torque ratio over repetitions",
    x = "Repetition",
    y = "EMG/torque ratio (RMS_norm / torque_norm)"
  )

print(ex_dyn_emg_torque_ratio_trend_plot)
```

```{r ex_dyn_combined_fatigue_plot}
# --- Reference lines: MVC_REF baseline per metric × channel ---
mvc_ref_mdf_baseline <- mvc_ref_mid1s_feature_table |>
  dplyr::group_by(channel) |>
  dplyr::summarise(ref_val = mean(mdf_mid_1s_hz, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(metric = factor("MDF (Hz)",
                                levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)")))

ex_dyn_ref_lines_df <- dplyr::bind_rows(
  tidyr::expand_grid(
    channel = c("emg6", "emg8", "emg10"),
    metric  = factor(c("Normalized torque", "RMS_norm (a.u.)"),
                     levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)"))
  ) |> dplyr::mutate(ref_val = 1),
  mvc_ref_mdf_baseline
)

ex_dyn_plot_long <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::select(channel, VC_count, torque_norm, rms_norm, mdf_mid_1s_hz) |>
  tidyr::pivot_longer(
    cols = c(torque_norm, rms_norm, mdf_mid_1s_hz),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_norm = "Normalized torque",
      rms_norm = "RMS_norm (a.u.)",
      mdf_mid_1s_hz = "MDF (Hz)"
    ),
    metric = factor(metric, levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)"))
  )

ex_dyn_recov_star_long <- mvc_recov_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    torque_norm   = mean(torque_norm,   na.rm = TRUE),
    rms_norm      = mean(rms_norm,      na.rm = TRUE),
    mdf_mid_1s_hz = mean(mdf_mid_1s_hz, na.rm = TRUE),
    .groups = "drop"
  ) |>
  tidyr::pivot_longer(
    cols = c(torque_norm, rms_norm, mdf_mid_1s_hz),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_norm   = "Normalized torque",
      rms_norm      = "RMS_norm (a.u.)",
      mdf_mid_1s_hz = "MDF (Hz)"
    ),
    metric = factor(metric, levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)")),
    x_star = max(ex_dyn_mid1s_feature_table_norm$VC_count) + 3L
  )

ex_dyn_combined_fatigue_plot <- ggplot(ex_dyn_plot_long, aes(x = VC_count, y = value)) +
  geom_hline(
    data = ex_dyn_ref_lines_df,
    aes(yintercept = ref_val),
    linetype = "dashed", color = "#888888", linewidth = 0.55
  ) +
  geom_point(alpha = 0.65, size = 1.0, color = "#4d4d4d") +
  facet_grid(metric ~ channel, scales = "free_y") +
  labs(
    title = "EX_DYN - Fatigue progression (middle 1s): normalized torque, RMS_norm, MDF",
    x = "Repetition",
    y = NULL
  ) +
  theme(strip.text.y = element_text(size = rel(0.78)))
print(ex_dyn_combined_fatigue_plot)
```

## EX_STA Trends

```{r ex_sta_emg_torque_ratio_vs_time}
ex_sta_emg_torque_ratio_trend_plot <- ggplot(
  ex_sta_emg_torque_ratio_table,
  aes(x = window_mid_s, y = emg_torque_ratio)
) +
  geom_point(alpha = 0.7, size = 1.2, color = "#1b6ca8") +
  geom_line(linewidth = 0.7, color = "#1b6ca8") +
  facet_wrap(~ channel, scales = "free_y") +
  labs(
    title = "EX_STA - EMG/torque ratio over time",
    x = "Window midpoint time (s)",
    y = "EMG/torque ratio (RMS_norm / torque_norm)"
  )

print(ex_sta_emg_torque_ratio_trend_plot)
```

```{r print_emg_torque_ratio_slope_summary}
print(emg_torque_ratio_slope_summary_table)
```

```{r ex_sta_triple_panel_fatigue}
# --- Reference lines + CI ribbons: SVC_REF 70% baseline per metric × channel ---
# RMS  : CI from normalized calibration fit (lm rms_norm ~ torque_norm, all 4 ref levels)
# MDF  : CI from Nm-space calibration fit   (lm mdf_mid_1s_hz ~ torque_mean_Nm, all 4 ref levels)
# Torque: fixed measured value from SVC rep 3 — no calibration CI available, line only

ex_sta_rms_ref_ci <- reference_level_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit        <- lm(rms_norm ~ torque_norm, data = .x)
    torque_70  <- dplyr::filter(.x, reference_level == "70%") |> dplyr::pull(torque_norm)
    pred       <- predict(fit, newdata = tibble::tibble(torque_norm = torque_70),
                          interval = "confidence", level = 0.95)
    tibble::tibble(ref_val = pred[1, "fit"], ref_lwr = pred[1, "lwr"], ref_upr = pred[1, "upr"])
  }) |>
  dplyr::ungroup() |>
  dplyr::mutate(metric = factor("RMS_norm (a.u.)",
                                levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)")))

ex_sta_mdf_ref_ci <- reference_level_feature_table |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit           <- lm(mdf_mid_1s_hz ~ torque_mean_Nm, data = .x)
    torque_70_nm  <- dplyr::filter(.x, reference_level == "70%") |> dplyr::pull(torque_mean_Nm)
    pred          <- predict(fit, newdata = tibble::tibble(torque_mean_Nm = torque_70_nm),
                             interval = "confidence", level = 0.95)
    tibble::tibble(ref_val = pred[1, "fit"], ref_lwr = pred[1, "lwr"], ref_upr = pred[1, "upr"])
  }) |>
  dplyr::ungroup() |>
  dplyr::mutate(metric = factor("MDF (Hz)",
                                levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)")))

ex_sta_ref_ribbon_df <- dplyr::bind_rows(
  reference_level_feature_table_norm |>
    dplyr::filter(reference_level == "70%") |>
    dplyr::group_by(channel) |>
    dplyr::summarise(ref_val = mean(torque_norm, na.rm = TRUE), .groups = "drop") |>
    dplyr::mutate(
      metric  = factor("Normalized torque",
                       levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)")),
      ref_lwr = NA_real_, ref_upr = NA_real_
    ),
  ex_sta_rms_ref_ci,
  ex_sta_mdf_ref_ci
)

ex_sta_plot_long <- ex_sta_mid1s_feature_table_norm |>
  dplyr::left_join(
    ex_sta_window_table |>
      dplyr::select(window_id, window_mid_s),
    by = "window_id"
  ) |>
  dplyr::select(channel, window_id, window_mid_s, torque_norm, rms_norm, mdf_mid_1s_hz) |>
  tidyr::pivot_longer(
    cols = c(torque_norm, rms_norm, mdf_mid_1s_hz),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_norm = "Normalized torque",
      rms_norm = "RMS_norm (a.u.)",
      mdf_mid_1s_hz = "MDF (Hz)"
    ),
    metric = factor(metric, levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)"))
  )

ex_sta_recov_star_long <- mvc_recov_sta_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    torque_norm   = mean(torque_norm,   na.rm = TRUE),
    rms_norm      = mean(rms_norm,      na.rm = TRUE),
    mdf_mid_1s_hz = mean(mdf_mid_1s_hz, na.rm = TRUE),
    .groups = "drop"
  ) |>
  tidyr::pivot_longer(
    cols = c(torque_norm, rms_norm, mdf_mid_1s_hz),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_norm   = "Normalized torque",
      rms_norm      = "RMS_norm (a.u.)",
      mdf_mid_1s_hz = "MDF (Hz)"
    ),
    metric = factor(metric, levels = c("Normalized torque", "RMS_norm (a.u.)", "MDF (Hz)")),
    x_star = max(ex_sta_plot_long$window_mid_s) + 5
  )

ex_sta_triple_panel_fatigue <- ggplot(ex_sta_plot_long, aes(x = window_mid_s, y = value)) +
  geom_rect(
    data        = dplyr::filter(ex_sta_ref_ribbon_df, !is.na(ref_lwr)),
    aes(ymin = ref_lwr, ymax = ref_upr, xmin = -Inf, xmax = Inf),
    fill        = "#888888", alpha = 0.10, inherit.aes = FALSE
  ) +
  geom_hline(
    data = ex_sta_ref_ribbon_df,
    aes(yintercept = ref_val),
    linetype = "dashed", color = "#888888", linewidth = 0.55
  ) +
  geom_point(alpha = 0.65, size = 1.0, color = "#4d4d4d") +
  facet_grid(metric ~ channel, scales = "free_y") +
  labs(
    title    = "EX_STA - Fatigue progression (1 s windows)",
    subtitle = "Dashed line = SVC_REF 70% baseline; ribbon = 95% CI from calibration fit (RMS and MDF only)",
    x        = "Time (s)",
    y        = NULL
  ) +
  theme(strip.text.y = element_text(size = rel(0.78)))
print(ex_sta_triple_panel_fatigue)
```


## Trend Summary Table

EX_DYN


```{r ex_dyn_summary_slopes_with_units}

ex_dyn_summary_table <-
  ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::summarise(

    `Normalized torque slope (ratio / rep)` =
      coef(lm(torque_norm ~ VC_count))[2],

    `RMS_norm slope (Δ RMS_norm / rep)` =
      coef(lm(rms_norm ~ VC_count))[2],

    `MDF slope (Hz / rep)` =
      coef(lm(mdf_mid_1s_hz ~ VC_count))[2],

    `RMS_norm vs normalized torque slope (Δ RMS_norm / ratio)` =
      coef(lm(rms_norm ~ torque_norm))[2],

    `MDF vs normalized torque slope (Hz / ratio)` =
      coef(lm(mdf_mid_1s_hz ~ torque_norm))[2],

    .groups = "drop"
  )

ex_dyn_summary_table

```



# Results

```{r emg_plot_gallery, fig.width=11, fig.height=7}
# Reprint the main EMG figures together at the end without recomputing them.
SHOW_MAIN_PLOT_GALLERY <- TRUE

if (isTRUE(SHOW_MAIN_PLOT_GALLERY)) {
  print(p_mdf)
  print(p_rms)
  print(percent_emg_vs_percent_force_plot)
  print(mvc_torque_summary_plot)
  print(mvc_rms_summary_plot)
  print(mvc_mdf_summary_plot)
  print(mvc_recovery_combined_plot)
  print(ex_dyn_torque_vs_rep_plot)
  print(ex_dyn_rms_vs_torque_plot)
  print(ex_dyn_mdf_vs_rep_plot)
  print(ex_dyn_mdf_vs_torque_plot)

  for (plot_obj in ex_dyn_dual_axis_plots) {
    print(plot_obj)
  }

  print(ex_dyn_rms_vs_repetition_plot)
  print(ex_dyn_reference_rms_overlay_plot)
  print(ex_dyn_reference_rms_overlay_plot_norm)
  print(ex_dyn_reference_residual_plot_norm)
  print(ex_dyn_reference_observed_expected_over_repetition_plot_norm)
  print(ex_dyn_reference_mdf_overlay_plot)
  print(ex_dyn_reference_mdf_residual_plot)
  print(ex_dyn_reference_mdf_observed_expected_plot)
  print(ex_dyn_emg_torque_ratio_trend_plot)
  print(ex_dyn_combined_fatigue_plot)
  print(ex_sta_emg_torque_ratio_trend_plot)
  print(ex_sta_triple_panel_fatigue)
}
```

```{r export_emg_plots, message=FALSE}
# Optional export: key summary plots as a single PDF.
# Toggle EXPORT_EMG_PLOTS and the per-plot flags below to control what is saved.
# Naming mirrors the Python QC export convention (RUN_ID-based filename).
EXPORT_EMG_PLOTS        <- TRUE
EXPORT_RMS_OVERLAY_NORM <- TRUE
EXPORT_DYN_COMBINED     <- TRUE
EXPORT_DYN_RATIO        <- TRUE
EXPORT_STA_COMBINED     <- TRUE
EXPORT_STA_RATIO        <- TRUE

plot_export_flags <- c(
  rms_overlay_norm = EXPORT_RMS_OVERLAY_NORM,
  dyn_combined     = EXPORT_DYN_COMBINED,
  dyn_ratio        = EXPORT_DYN_RATIO,
  sta_combined     = EXPORT_STA_COMBINED,
  sta_ratio        = EXPORT_STA_RATIO
)

plot_export_map <- list(
  rms_overlay_norm = ex_dyn_reference_rms_overlay_plot_norm,
  dyn_combined     = ex_dyn_combined_fatigue_plot,
  dyn_ratio        = ex_dyn_emg_torque_ratio_trend_plot,
  sta_combined     = ex_sta_triple_panel_fatigue,
  sta_ratio        = ex_sta_emg_torque_ratio_trend_plot
)

if (isTRUE(EXPORT_EMG_PLOTS)) {
  selected_plot_names <- names(plot_export_flags)[plot_export_flags]
  if (length(selected_plot_names) == 0L) {
    stop("EXPORT_EMG_PLOTS=TRUE but no plot selector is set to TRUE.")
  }

  qc_export_root <- file.path("results", "QC_EXPORT")
  dir.create(qc_export_root, recursive = TRUE, showWarnings = FALSE)

  qc_emg_export_path <- file.path(
    qc_export_root,
    paste0("qc_emg_export_", RUN_ID, ".pdf")
  )

  grDevices::pdf(file = qc_emg_export_path, width = 11, height = 8.5, onefile = TRUE)
  for (plot_name in selected_plot_names) {
    print(plot_export_map[[plot_name]])
  }
  grDevices::dev.off()

  message("\n[QC EMG EXPORT] PDF -> ", normalizePath(qc_emg_export_path, winslash = "/", mustWork = FALSE))
}
```

```{r export_analysis_tables, message=FALSE}
# Optional export: per-run derived tables for later cohort analysis.
# Keep this separate from the Python DATA_EXPORT parquet files.
EXPORT_ANALYSIS_TABLES              <- TRUE
EXPORT_RUN_METADATA                 <- TRUE
EXPORT_REFERENCE_LEVEL_FEATURES     <- TRUE
EXPORT_EX_DYN_FEATURES              <- TRUE
EXPORT_EX_STA_FEATURES              <- TRUE
EXPORT_EX_DYN_RATIO_TABLE           <- TRUE
EXPORT_EX_STA_RATIO_TABLE           <- TRUE
EXPORT_RATIO_SLOPE_SUMMARY          <- TRUE
EXPORT_REFERENCE_CALIBRATION_NORM   <- TRUE

analysis_export_flags <- c(
  run_metadata                   = EXPORT_RUN_METADATA,
  reference_level_features_norm  = EXPORT_REFERENCE_LEVEL_FEATURES,
  ex_dyn_features_norm           = EXPORT_EX_DYN_FEATURES,
  ex_sta_features_norm           = EXPORT_EX_STA_FEATURES,
  ex_dyn_ratio_table             = EXPORT_EX_DYN_RATIO_TABLE,
  ex_sta_ratio_table             = EXPORT_EX_STA_RATIO_TABLE,
  ratio_slope_summary            = EXPORT_RATIO_SLOPE_SUMMARY,
  reference_calibration_norm     = EXPORT_REFERENCE_CALIBRATION_NORM
)

run_metadata_table <- participant_table |>
  dplyr::transmute(
    RUN_ID,
    participant_id,
    sex,
    height_cm,
    mass_kg,
    dominant_side,
    practice_level,
    run_number,
    mvc_ref
  )

reference_level_feature_table_norm_export <- reference_level_feature_table_norm |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

ex_dyn_mid1s_feature_table_norm_export <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

ex_sta_mid1s_feature_table_norm_export <- ex_sta_mid1s_feature_table_norm |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

ex_dyn_emg_torque_ratio_table_export <- ex_dyn_emg_torque_ratio_table |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

ex_sta_emg_torque_ratio_table_export <- ex_sta_emg_torque_ratio_table |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

emg_torque_ratio_slope_summary_table_export <- emg_torque_ratio_slope_summary_table |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

reference_calibration_table_norm_export <- reference_calibration_table_norm |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

analysis_export_map <- list(
  run_metadata                  = run_metadata_table,
  reference_level_features_norm = reference_level_feature_table_norm_export,
  ex_dyn_features_norm          = ex_dyn_mid1s_feature_table_norm_export,
  ex_sta_features_norm          = ex_sta_mid1s_feature_table_norm_export,
  ex_dyn_ratio_table            = ex_dyn_emg_torque_ratio_table_export,
  ex_sta_ratio_table            = ex_sta_emg_torque_ratio_table_export,
  ratio_slope_summary           = emg_torque_ratio_slope_summary_table_export,
  reference_calibration_norm    = reference_calibration_table_norm_export
)

analysis_export_file_map <- c(
  run_metadata                  = "run_metadata.parquet",
  reference_level_features_norm = "reference_level_feature_table_norm.parquet",
  ex_dyn_features_norm          = "ex_dyn_mid1s_feature_table_norm.parquet",
  ex_sta_features_norm          = "ex_sta_mid1s_feature_table_norm.parquet",
  ex_dyn_ratio_table            = "ex_dyn_emg_torque_ratio_table.parquet",
  ex_sta_ratio_table            = "ex_sta_emg_torque_ratio_table.parquet",
  ratio_slope_summary           = "emg_torque_ratio_slope_summary_table.parquet",
  reference_calibration_norm    = "reference_calibration_table_norm.parquet"
)

if (isTRUE(EXPORT_ANALYSIS_TABLES)) {
  selected_table_names <- names(analysis_export_flags)[analysis_export_flags]
  if (length(selected_table_names) == 0L) {
    stop("EXPORT_ANALYSIS_TABLES=TRUE but no analysis table selector is set to TRUE.")
  }

  analysis_export_root <- file.path("results", "R_ANALYSIS", RUN_ID)
  dir.create(analysis_export_root, recursive = TRUE, showWarnings = FALSE)

  for (table_name in selected_table_names) {
    arrow::write_parquet(
      analysis_export_map[[table_name]],
      sink = file.path(analysis_export_root, analysis_export_file_map[[table_name]])
    )
  }

  message("\n[R ANALYSIS EXPORT] Tables -> ", normalizePath(analysis_export_root, winslash = "/", mustWork = FALSE))
}
```

