---
title: "AMSTROM BIA Cohort Analysis"
author: "Morgan Virolan"
date: "2026-05-16"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: cosmo
---

# Overview

Cohort BIA/EIM analysis: pool run-level exports, harmonize EX_STA duration, plot RelDev/Cole trajectories, and optionally write cohort tables.

---

# 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(arrow)
library(dplyr)
library(tidyr)
library(ggplot2)

options(dplyr.summarise.inform = FALSE)
source(here::here("sources", "R", "config", "amstrom_style.R"))
source(here::here("sources", "R", "config", "cohort_config.R"))

freq_levels <- c("9.76 kHz", "48.8 kHz", "97.6 kHz")
PRINT_SUMMARY_TABLES <- FALSE

# Colour per frequency (reference/checkpoint plot)
bia_freq_pal <- c(
  "9.76 kHz" = "#b5a0cc",
  "48.8 kHz" = "#7b5ea7",
  "97.6 kHz" = "#4a2e7f"
)
```

---

# 2. Discover cohort input runs

Runs are pooled from the shared `cohort_final` block so all cohort notebooks use the same sample.

```{r discover_paths}
analysis_root        <- here::here("results", "R_ANALYSIS")
export_root          <- here::here("results", "DATA_EXPORT")
cohort_output_root   <- file.path(analysis_root, "cohort")
subset_registry_path <- file.path(analysis_root, "subset_registry",
                                   "multimodal_subset_registry.parquet")
```

```{r discover_run_ids}
if (!file.exists(subset_registry_path)) {
  stop(
    "Subset registry not found: ",
    normalizePath(subset_registry_path, winslash = "/", mustWork = FALSE),
    "\nRun multimodal_subset_checker.R first."
  )
}

subset_registry <- arrow::read_parquet(subset_registry_path)

run_ids <- subset_registry |>
  dplyr::filter(.data$analysis_block == "cohort_final", .data$included == 1L) |>
  dplyr::pull(.data$RUN_ID) |>
  sort()

if (length(run_ids) == 0L) {
  stop("No runs in cohort_final. Re-run multimodal_subset_checker.R.")
}

cat("Cohort N (cohort_final):", length(run_ids), "\n")
run_ids
```

---

# 3. Pool run-level BIA tables

Read each run's BIA exports and stack into pooled tables. Only the three analysis frequencies
(9.76 / 48.8 / 97.6 kHz) are kept, and 9.76 kHz Xc/PhA are dropped here and at §4 (inductive
artifact -> see the §4 comment).

```{r pool_bia_tables}
# read_bia_table: read one per-run BIA export, erroring if a file is missing.
read_bia_table <- function(run_id, file_name) {
  path <- file.path(analysis_root, run_id, file_name)
  if (!file.exists(path)) {
    stop("Missing BIA file: ", normalizePath(path, winslash = "/", mustWork = FALSE))
  }
  return(arrow::read_parquet(path))
}

cohort_bia_reference_feature_table <- dplyr::bind_rows(
  lapply(run_ids, function(r) read_bia_table(r, "bia_reference_feature_table.parquet"))
) |>
  dplyr::filter(.data$freq_hz %in% c(9760L, 48800L, 97600L)) |>
  dplyr::mutate(freq_label = factor(.data$freq_label, levels = freq_levels)) |>
  dplyr::arrange(.data$RUN_ID, .data$summary_variable, .data$freq_hz, .data$state_label)

cohort_bia_ex_dyn_feature_table <- dplyr::bind_rows(
  lapply(run_ids, function(r) read_bia_table(r, "bia_ex_dyn_feature_table.parquet"))
) |>
  dplyr::filter(.data$freq_hz %in% c(9760L, 48800L, 97600L)) |>
  dplyr::mutate(freq_label = factor(.data$freq_label, levels = freq_levels)) |>
  dplyr::arrange(.data$RUN_ID, .data$summary_variable, .data$freq_hz, .data$VC_count) |>
  # 9.76 kHz Xc and PhA excluded — same inductive-parasitic artifact as EX_STA (see §4).
  dplyr::filter(!(.data$freq_hz == 9760L & .data$summary_variable %in% c("Xc", "PhA")))

cohort_bia_ex_sta_feature_table_raw <- dplyr::bind_rows(
  lapply(run_ids, function(r) read_bia_table(r, "bia_ex_sta_feature_table.parquet"))
) |>
  dplyr::filter(.data$freq_hz %in% c(9760L, 48800L, 97600L)) |>
  dplyr::mutate(freq_label = factor(.data$freq_label, levels = freq_levels)) |>
  dplyr::arrange(.data$RUN_ID, .data$summary_variable, .data$freq_hz, .data$window_index)

cohort_bia_run_summary_table <- dplyr::bind_rows(
  lapply(run_ids, function(r) read_bia_table(r, "bia_run_summary_table.parquet"))
) |>
  dplyr::arrange(.data$analysis_block, .data$summary_variable, .data$freq_hz, .data$RUN_ID)
```

---

# 4. EX_STA common-duration harmonization

EX_STA is truncated to `GLOBAL_STA_N_MIN` so all modalities share the same window count.

```{r harmonize_ex_sta_common_duration}
cohort_bia_ex_sta_run_profile <- cohort_bia_ex_sta_feature_table_raw |>
  dplyr::group_by(.data$RUN_ID) |>
  dplyr::summarise(last_window_index = max(.data$window_index, na.rm = TRUE),  # per-run EX_STA length
                   .groups = "drop")

sta_n_min <- GLOBAL_STA_N_MIN

cohort_bia_ex_sta_feature_table <- cohort_bia_ex_sta_feature_table_raw |>
  dplyr::filter(.data$window_index <= sta_n_min) |>
  # 9.76 kHz Xc and PhA excluded: same inductive-parasitic artifact that caused 4880 Hz
  # to be dropped in Cole fitting. At 9.76 kHz, tissue Xc is near the measurement noise
  # floor for most subjects (Cole arc far below fc ~75 kHz). Inductive artifact contaminates
  # the near-zero Xc, producing RelDev denominators close to zero and swings up to ±6724%.
  # R and absZ at 9.76 kHz are unaffected (resistance >> artifact). See article methods.
  dplyr::filter(!(.data$freq_hz == 9760L & .data$summary_variable %in% c("Xc", "PhA")))

cohort_bia_ex_sta_run_profile
cat("Truncating EX_STA to", sta_n_min, "windows (GLOBAL_STA_N_MIN from cohort_config.R).\n")
cat("9.76 kHz Xc and PhA excluded (inductive artifact — see comment above).\n")
```

---

# 5. Optional cohort export

Writes the pooled tables to `results/R_ANALYSIS/cohort/` when `EXPORT_COHORT_PARQUET` is TRUE.
Off by default -- the analysis below runs from the in-memory pools regardless.

```{r cohort_export_control}
EXPORT_COHORT_PARQUET <- FALSE

if (isTRUE(EXPORT_COHORT_PARQUET)) {
  dir.create(cohort_output_root, recursive = TRUE, showWarnings = FALSE)
  arrow::write_parquet(cohort_bia_reference_feature_table,
    sink = file.path(cohort_output_root, "cohort_bia_reference_feature_table.parquet"))
  arrow::write_parquet(cohort_bia_ex_dyn_feature_table,
    sink = file.path(cohort_output_root, "cohort_bia_ex_dyn_feature_table.parquet"))
  arrow::write_parquet(cohort_bia_ex_sta_feature_table,
    sink = file.path(cohort_output_root, "cohort_bia_ex_sta_feature_table.parquet"))
  arrow::write_parquet(cohort_bia_run_summary_table,
    sink = file.path(cohort_output_root, "cohort_bia_run_summary_table.parquet"))
}
```

---

# 6. Cole model — spectral structure

Cole arcs are reconstructed only over the measured range (9.76–122 kHz); `R_inf` and high `fc_hz` fits remain extrapolation-sensitive.

## 6.1 Pool and harmonize

Pool the three per-run Cole tables (reference / EX_DYN / EX_STA); EX_STA is truncated to
`GLOBAL_STA_N_MIN` to match the other modalities.

```{r pool_cole_tables}
cole_params <- c("R_inf", "delta_R", "R0", "fc_hz", "alpha")

cohort_cole_reference_table <- dplyr::bind_rows(
  lapply(run_ids, function(r) read_bia_table(r, "bia_cole_reference_table.parquet"))
) |>
  dplyr::arrange(.data$RUN_ID, .data$source_sequence)

cohort_cole_ex_dyn_table <- dplyr::bind_rows(
  lapply(run_ids, function(r) read_bia_table(r, "bia_cole_ex_dyn_table.parquet"))
) |>
  dplyr::arrange(.data$RUN_ID, .data$VC_count)

cohort_cole_ex_sta_table <- dplyr::bind_rows(
  lapply(run_ids, function(r) read_bia_table(r, "bia_cole_ex_sta_table.parquet"))
) |>
  dplyr::filter(.data$window_index <= GLOBAL_STA_N_MIN) |>
  dplyr::arrange(.data$RUN_ID, .data$window_index)
```

## 6.2 Nyquist arc plot

```{r cole_nyquist, fig.width=10, fig.height=4}
# Reconstruct Cole arc from parameters over a frequency sweep
reconstruct_arc <- function(R_inf, delta_R, fc_hz, alpha, n_pts = 400) {
  # Use actual measurement range only — fc is often above 122 kHz so arcs are partial
  freqs <- 10^seq(log10(9760), log10(122000), length.out = n_pts)
  z     <- R_inf + delta_R / (1 + (1i * freqs / fc_hz)^alpha)
  return(data.frame(cole_R = Re(z), cole_negXc = -Im(z)))
}

# --- EX_DYN: mean arc at 4 equally-spaced reps ---
mean_cole_ex_dyn <- cohort_cole_ex_dyn_table |>
  dplyr::filter(!is.na(.data$R_inf)) |>
  dplyr::group_by(.data$VC_count) |>
  dplyr::summarise(
    R_inf   = mean(.data$R_inf,   na.rm = TRUE),  # cohort mean R_inf [extrapolated]
    delta_R = mean(.data$delta_R, na.rm = TRUE),  # cohort mean delta_R
    fc_hz   = mean(.data$fc_hz,   na.rm = TRUE),  # cohort mean fc [extrapolated]
    alpha   = mean(.data$alpha,   na.rm = TRUE),  # cohort mean alpha
    .groups = "drop"
  )

all_vcs  <- sort(unique(mean_cole_ex_dyn$VC_count))
sel_vcs  <- all_vcs[round(seq(1, length(all_vcs), length.out = 4))]

ex_dyn_arcs <- dplyr::bind_rows(lapply(sel_vcs, function(vc) {
  p <- mean_cole_ex_dyn[mean_cole_ex_dyn$VC_count == vc, ]
  reconstruct_arc(p$R_inf, p$delta_R, p$fc_hz, p$alpha) |>
    dplyr::mutate(label = paste0("Rep ", vc))
})) |>
  dplyr::mutate(label = factor(.data$label, levels = paste0("Rep ", sel_vcs)))

# --- Reference sequences ---
mean_cole_ref <- cohort_cole_reference_table |>
  dplyr::filter(!is.na(.data$R_inf)) |>
  dplyr::group_by(.data$source_sequence) |>
  dplyr::summarise(
    R_inf   = mean(.data$R_inf,   na.rm = TRUE),  # cohort mean R_inf [extrapolated]
    delta_R = mean(.data$delta_R, na.rm = TRUE),  # cohort mean delta_R
    fc_hz   = mean(.data$fc_hz,   na.rm = TRUE),  # cohort mean fc [extrapolated]
    alpha   = mean(.data$alpha,   na.rm = TRUE),  # cohort mean alpha
    .groups = "drop"
  )

ref_seq_levels <- c("SVC_REF", "MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")
ref_seq_present <- intersect(ref_seq_levels, mean_cole_ref$source_sequence)

ref_arcs <- dplyr::bind_rows(lapply(ref_seq_present, function(seq_name) {
  p <- mean_cole_ref[mean_cole_ref$source_sequence == seq_name, ]
  reconstruct_arc(p$R_inf, p$delta_R, p$fc_hz, p$alpha) |>
    dplyr::mutate(label = seq_name)
})) |>
  dplyr::mutate(label = factor(.data$label, levels = ref_seq_present))

# --- EX_STA ---
mean_cole_ex_sta <- cohort_cole_ex_sta_table |>
  dplyr::filter(!is.na(.data$R_inf)) |>
  dplyr::group_by(.data$window_index) |>
  dplyr::summarise(
    R_inf   = mean(.data$R_inf,   na.rm = TRUE),  # cohort mean R_inf [extrapolated]
    delta_R = mean(.data$delta_R, na.rm = TRUE),  # cohort mean delta_R
    fc_hz   = mean(.data$fc_hz,   na.rm = TRUE),  # cohort mean fc [extrapolated]
    alpha   = mean(.data$alpha,   na.rm = TRUE),  # cohort mean alpha
    .groups = "drop"
  )

all_wins  <- sort(unique(mean_cole_ex_sta$window_index))
sel_wins  <- all_wins[round(seq(1, length(all_wins), length.out = 4))]

ex_sta_arcs <- dplyr::bind_rows(lapply(sel_wins, function(wi) {
  p <- mean_cole_ex_sta[mean_cole_ex_sta$window_index == wi, ]
  reconstruct_arc(p$R_inf, p$delta_R, p$fc_hz, p$alpha) |>
    dplyr::mutate(label = paste0("Win ", wi))
})) |>
  dplyr::mutate(label = factor(.data$label, levels = paste0("Win ", sel_wins)))

# --- Plots ---
nyquist_theme <- theme_amstrom_hgrid +
  theme(legend.position = "right")

p_ref_nyq <- ggplot(ref_arcs, aes(x = .data$cole_R, y = .data$cole_negXc,
                                    colour = .data$label)) +
  geom_path(linewidth = 0.9) +
  coord_fixed() +
  scale_colour_viridis_d(option = "plasma", direction = -1, name = NULL) +
  labs(title = "Cole Nyquist - reference sequences",
       subtitle = "Arcs over 9.76-122 kHz. fc median ~75 kHz (most fits within window); ~10% of reps fc > 122 kHz.",
       x = "R (Ohm)", y = "-Xc (Ohm)") +
  nyquist_theme

p_dyn_nyq <- ggplot(ex_dyn_arcs, aes(x = .data$cole_R, y = .data$cole_negXc,
                                       colour = .data$label)) +
  geom_path(linewidth = 0.9) +
  coord_fixed() +
  scale_colour_viridis_d(option = "plasma", direction = -1, name = NULL) +
  labs(title = "Cole Nyquist - EX_DYN (dark = early, bright = late)",
       subtitle = "Arcs over 9.76-122 kHz. fc median ~75 kHz (most fits within window); ~10% of reps fc > 122 kHz.",
       x = "R (Ohm)", y = "-Xc (Ohm)") +
  nyquist_theme

p_sta_nyq <- ggplot(ex_sta_arcs, aes(x = .data$cole_R, y = .data$cole_negXc,
                                       colour = .data$label)) +
  geom_path(linewidth = 0.9) +
  coord_fixed() +
  scale_colour_viridis_d(option = "plasma", direction = -1, name = NULL) +
  labs(title = "Cole Nyquist - EX_STA (dark = early, bright = late)",
       subtitle = "Arcs over 9.76-122 kHz. fc median ~75 kHz (most fits within window); ~10% of reps fc > 122 kHz.",
       x = "R (Ohm)", y = "-Xc (Ohm)") +
  nyquist_theme

print(p_ref_nyq)
print(p_dyn_nyq)
print(p_sta_nyq)
```

## 6.3 Cole parameter trajectories

```{r cole_parameter_trajectories, fig.width=10, fig.height=10}
# Cohort mean +/- SD per timepoint for each Cole parameter (the ribbon shows the SD band).
# R_inf and fc carry an [extrapolated] tag in the labels.
cole_param_labels <- c(
  R_inf   = "R_inf (Ohm) [extrapolated]",
  delta_R = "delta_R (Ohm)",
  R0      = "R0 (Ohm)",
  fc_hz   = "fc (Hz) [extrapolated]",
  alpha   = "alpha"
)

# --- EX_DYN ---
cole_ex_dyn_long <- cohort_cole_ex_dyn_table |>
  dplyr::filter(!is.na(.data$R_inf)) |>
  tidyr::pivot_longer(cols = all_of(cole_params), names_to = "parameter", values_to = "value") |>
  dplyr::group_by(.data$VC_count, .data$parameter) |>
  dplyr::summarise(
    mean_value = mean(.data$value, na.rm = TRUE),  # cohort mean of this Cole parameter
    sd_value   = sd(.data$value, na.rm = TRUE),    # between-run SD (drawn as the ribbon)
    .groups    = "drop"
  ) |>
  dplyr::mutate(parameter = factor(.data$parameter, levels = cole_params))

# --- EX_STA ---
cole_ex_sta_long <- cohort_cole_ex_sta_table |>
  dplyr::filter(!is.na(.data$R_inf)) |>
  tidyr::pivot_longer(cols = all_of(cole_params), names_to = "parameter", values_to = "value") |>
  dplyr::group_by(.data$window_index, .data$parameter) |>
  dplyr::summarise(
    mean_value = mean(.data$value, na.rm = TRUE),  # cohort mean of this Cole parameter
    sd_value   = sd(.data$value, na.rm = TRUE),    # between-run SD (drawn as the ribbon)
    .groups    = "drop"
  ) |>
  dplyr::mutate(parameter = factor(.data$parameter, levels = cole_params))

p_cole_dyn <- ggplot(cole_ex_dyn_long,
  aes(x = .data$VC_count, y = .data$mean_value)) +
  geom_ribbon(aes(ymin = .data$mean_value - .data$sd_value,
                  ymax = .data$mean_value + .data$sd_value),
              fill = col_bia, alpha = 0.15) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_bia) +
  geom_point(shape = 15, size = 2.2, colour = col_bia) +
  facet_wrap(~ parameter, ncol = 2, scales = "free_y",
             labeller = ggplot2::labeller(parameter = cole_param_labels)) +
  labs(title = "Cole parameters - EX_DYN cohort mean",
       subtitle = "R_inf and fc are extrapolated (fc typically 70-570 kHz, above 122 kHz ceiling)",
       x = "Repetition", y = NULL, caption = "Ribbon = mean +/- SD") +
  theme_amstrom_hgrid

p_cole_sta <- ggplot(cole_ex_sta_long,
  aes(x = .data$window_index, y = .data$mean_value)) +
  geom_ribbon(aes(ymin = .data$mean_value - .data$sd_value,
                  ymax = .data$mean_value + .data$sd_value),
              fill = col_bia, alpha = 0.15) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_bia) +
  geom_point(shape = 15, size = 2.2, colour = col_bia) +
  facet_wrap(~ parameter, ncol = 2, scales = "free_y",
             labeller = ggplot2::labeller(parameter = cole_param_labels)) +
  labs(title = "Cole parameters - EX_STA cohort mean",
       subtitle = "R_inf and fc are extrapolated (fc typically 70-570 kHz, above 122 kHz ceiling)",
       x = "Window index", y = NULL, caption = "Ribbon = mean +/- SD") +
  theme_amstrom_hgrid

print(p_cole_dyn)
print(p_cole_sta)
```

---

# 7. BIA RelDev trajectories

Cohort mean RelDev from the sequence-matched baseline.

## 7.1 All variables — EX_DYN and EX_STA

```{r bia_reldev_trajectories, fig.width=9, fig.height=7}
# --- EX_DYN RelDev ---
ex_dyn_reldev_mean <- cohort_bia_ex_dyn_feature_table |>
  dplyr::group_by(.data$VC_count, .data$summary_variable, .data$freq_label) |>
  dplyr::summarise(mean_reldev = mean(.data$reldev_pct, na.rm = TRUE), .groups = "drop")  # cohort mean RelDev from MVC_REF

p_ex_dyn_reldev <- ggplot(
  ex_dyn_reldev_mean,
  aes(x = .data$VC_count, y = .data$mean_reldev,
      colour = .data$freq_label, group = .data$freq_label)
) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  geom_line(linewidth = 0.35) +
  geom_point(shape = 15, size = 2.2) +
  facet_wrap(~ summary_variable, ncol = 2, scales = "free_y") +
  scale_colour_manual(values = bia_freq_pal, name = NULL) +
  labs(
    title   = "BIA EX_DYN - RelDev from MVC_REF",
    x       = "Repetition",
    y       = "RelDev (%)",
    caption = "Cohort mean. Dashed line = 0% (MVC_REF level). Curves start below 0 due to SVC_REF pre-loading."
  ) +
  theme_amstrom_hgrid

# --- EX_STA RelDev ---
ex_sta_reldev_mean <- cohort_bia_ex_sta_feature_table |>
  dplyr::group_by(.data$window_index, .data$summary_variable, .data$freq_label) |>
  dplyr::summarise(mean_reldev = mean(.data$reldev_seq_pct, na.rm = TRUE), .groups = "drop")  # cohort mean within-sequence RelDev

p_ex_sta_reldev <- ggplot(
  ex_sta_reldev_mean,
  aes(x = .data$window_index, y = .data$mean_reldev,
      colour = .data$freq_label, group = .data$freq_label)
) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  geom_line(linewidth = 0.35) +
  geom_point(shape = 15, size = 2.2) +
  facet_wrap(~ summary_variable, ncol = 2, scales = "free_y") +
  scale_colour_manual(values = bia_freq_pal, name = NULL) +
  labs(
    title   = "BIA EX_STA - RelDev from sequence onset (first 3 windows)",
    x       = "Window index",
    y       = "RelDev (%)",
    caption = "Cohort mean. Dashed line = 0% (sequence onset)."
  ) +
  theme_amstrom_hgrid

print(p_ex_dyn_reldev)
print(p_ex_sta_reldev)
```

## 7.2 Phase angle — standalone

```{r bia_pha_trajectory, fig.width=8, fig.height=8}
# Phase angle on its own (the most discriminating BIA fatigue marker): EX_DYN and EX_STA, per freq.
pha_dyn_mean <- ex_dyn_reldev_mean |>
  dplyr::filter(.data$summary_variable == "PhA")

pha_sta_mean <- ex_sta_reldev_mean |>
  dplyr::filter(.data$summary_variable == "PhA")

p_pha_dyn <- ggplot(pha_dyn_mean,
  aes(x = .data$VC_count, y = .data$mean_reldev)
) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  geom_line(linewidth = 0.35, colour = col_bia) +
  geom_point(shape = 15, size = 2.2, colour = col_bia) +
  facet_wrap(~ freq_label, ncol = 1, scales = "free_y") +
  labs(title = "Phase angle - EX_DYN",
       x = "Repetition", y = "PhA RelDev (%)") +
  theme_amstrom_hgrid

p_pha_sta <- ggplot(pha_sta_mean,
  aes(x = .data$window_index, y = .data$mean_reldev)
) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  geom_line(linewidth = 0.35, colour = col_bia) +
  geom_point(shape = 15, size = 2.2, colour = col_bia) +
  facet_wrap(~ freq_label, ncol = 1, scales = "free_y") +
  labs(title = "Phase angle - EX_STA",
       x = "Window index", y = "PhA RelDev (%)",
       caption = "Baseline = SVC_70% mean. EX_STA onset carries pre-loaded fatigue from EX_DYN.") +
  theme_amstrom_hgrid

print(p_pha_dyn)
print(p_pha_sta)
```

## 7.3 Diagnostic: 9.76 kHz EX_STA extreme swings

```{r diag_9khz_ex_sta, fig.width=10, fig.height=7}
sta_9k <- cohort_bia_ex_sta_feature_table |>
  dplyr::filter(.data$freq_hz == 9760)

# --- Table: worst offenders per variable ---
outlier_table <- sta_9k |>
  dplyr::group_by(.data$RUN_ID, .data$summary_variable) |>
  dplyr::summarise(
    max_abs_reldev = round(max(abs(.data$reldev_seq_pct), na.rm = TRUE), 1),  # worst swing magnitude
    min_reldev     = round(min(.data$reldev_seq_pct, na.rm = TRUE), 1),       # most negative
    max_reldev     = round(max(.data$reldev_seq_pct, na.rm = TRUE), 1),       # most positive
    .groups = "drop"
  ) |>
  dplyr::arrange(dplyr::desc(.data$max_abs_reldev))

cat("Top 15 run x variable combinations by max |RelDev| at 9.76 kHz EX_STA:\n")
print(head(outlier_table, 15))

# --- Spaghetti plot ---
sta_9k_mean <- sta_9k |>
  dplyr::group_by(.data$window_index, .data$summary_variable) |>
  dplyr::summarise(mean_reldev = mean(.data$reldev_seq_pct, na.rm = TRUE),
                   .groups = "drop")

ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72",
             linewidth = 0.35) +
  geom_line(
    data = sta_9k,
    aes(x = .data$window_index, y = .data$reldev_seq_pct, group = .data$RUN_ID),
    colour = "grey75", linewidth = 0.35, alpha = 0.7
  ) +
  geom_line(
    data = sta_9k_mean,
    aes(x = .data$window_index, y = .data$mean_reldev),
    colour = col_bia, linewidth = 1.1
  ) +
  facet_wrap(~ summary_variable, ncol = 2, scales = "free_y") +
  labs(
    title   = "9.76 kHz EX_STA - individual runs (grey) vs. cohort mean (purple)",
    x       = "Window index",
    y       = "RelDev (%)",
    caption = "If only a few grey lines are extreme, the mean is being pulled by outliers."
  ) +
  theme_amstrom_hgrid
```

---

# 8. BIA reference and checkpoint overview

Cohort mean ± SD of absolute BIA values across contraction levels and recovery checkpoints,
faceted by variable and coloured by frequency.

```{r bia_reference_overview, fig.width=10, fig.height=8}
# Cohort mean/SD of absolute BIA values per state, faceted by variable, coloured by frequency.
reference_plot_table <- cohort_bia_reference_feature_table |>
  dplyr::group_by(.data$state_label, .data$summary_variable,
                  .data$freq_hz, .data$freq_label) |>
  dplyr::summarise(
    mean_value = mean(.data$summary_value, na.rm = TRUE),  # cohort mean absolute value
    sd_value   = sd(.data$summary_value, na.rm = TRUE),    # between-run SD
    n_runs     = dplyr::n_distinct(.data$RUN_ID),          # runs at this state
    .groups    = "drop"
  ) |>
  dplyr::mutate(
    state_label = factor(.data$state_label,
      levels = c("SVC_30%", "SVC_50%", "SVC_70%", "MVC_REF",
                 "MVC_RECOV_DYN", "MVC_RECOV_STA"))
  )

p_reference <- ggplot(
  reference_plot_table,
  aes(x = .data$state_label, y = .data$mean_value,
      colour = .data$freq_label, group = .data$freq_label)
) +
  geom_errorbar(
    aes(ymin = .data$mean_value - .data$sd_value,
        ymax = .data$mean_value + .data$sd_value),
    width = 0, linewidth = 0.4, alpha = 0.28, position = position_dodge(0.5)
  ) +
  geom_line(linetype = "dotted", linewidth = 0.35, position = position_dodge(0.5)) +
  geom_point(shape = 15, size = 2.2, position = position_dodge(0.5)) +
  facet_wrap(~ summary_variable, ncol = 2, scales = "free_y") +
  scale_colour_manual(values = bia_freq_pal, name = NULL) +
  labs(
    title   = "BIA reference and checkpoint — cohort overview",
    subtitle = "Absolute values across contraction levels and recovery checkpoints",
    x       = NULL, y = NULL,
    caption = "Error bars = mean +/- SD. SVC = sustained voluntary contraction; MVC = maximal voluntary contraction."
  ) +
  theme_amstrom_hgrid +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

print(p_reference)
```

---

# 9. Summary tables

Set `PRINT_SUMMARY_TABLES <- TRUE` in §1 setup to render these.

## Reference absolute values (key sequences)

```{r summary_ref, eval = PRINT_SUMMARY_TABLES}
reference_plot_table |>
  dplyr::filter(.data$state_label %in% c("SVC_70%", "MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")) |>
  dplyr::mutate(dplyr::across(c(mean_value, sd_value), ~round(.x, 3))) |>
  dplyr::select(state_label, summary_variable, freq_label, mean_value, sd_value, n_runs) |>
  dplyr::arrange(.data$summary_variable, .data$freq_label, .data$state_label) |>
  knitr::kable(
    col.names = c("State", "Variable", "Freq", "Mean", "SD", "N"),
    caption   = "BIA absolute values at key reference sequences (cohort mean +/- SD)."
  )
```

## EX_DYN RelDev - first vs last rep

```{r summary_ex_dyn, eval = PRINT_SUMMARY_TABLES}
ex_dyn_reldev_mean |>
  dplyr::group_by(.data$summary_variable, .data$freq_label) |>
  dplyr::filter(.data$VC_count == min(.data$VC_count) | .data$VC_count == max(.data$VC_count)) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    position     = dplyr::if_else(.data$VC_count == min(.data$VC_count), "First rep", "Last rep"),
    mean_reldev  = round(.data$mean_reldev, 2)
  ) |>
  dplyr::select(summary_variable, freq_label, position, VC_count, mean_reldev) |>
  dplyr::arrange(.data$summary_variable, .data$freq_label, .data$VC_count) |>
  knitr::kable(
    col.names = c("Variable", "Freq", "Position", "Rep", "Mean RelDev (%)"),
    caption   = "EX_DYN RelDev from MVC_REF - first vs last rep. Expected direction: decrease (Freeborn & Fu 2019)."
  )
```

## EX_STA RelDev - first vs last window

```{r summary_ex_sta, eval = PRINT_SUMMARY_TABLES}
ex_sta_reldev_mean |>
  dplyr::group_by(.data$summary_variable, .data$freq_label) |>
  dplyr::filter(.data$window_index == min(.data$window_index) | .data$window_index == max(.data$window_index)) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    position    = dplyr::if_else(.data$window_index == min(.data$window_index), "First window", "Last window"),
    mean_reldev = round(.data$mean_reldev, 2)
  ) |>
  dplyr::select(summary_variable, freq_label, position, window_index, mean_reldev) |>
  dplyr::arrange(.data$summary_variable, .data$freq_label, .data$window_index) |>
  knitr::kable(
    col.names = c("Variable", "Freq", "Position", "Window", "Mean RelDev (%)"),
    caption   = "EX_STA RelDev from SVC_70% - first vs last window. PhA is the most discriminating marker (Dalton-Alves 2025)."
  )
```

## Cole parameters at reference sequences

```{r summary_cole, eval = PRINT_SUMMARY_TABLES}
cohort_cole_reference_table |>
  dplyr::filter(!is.na(.data$R_inf)) |>
  dplyr::group_by(.data$source_sequence) |>
  dplyr::summarise(
    n       = dplyr::n_distinct(.data$RUN_ID),
    R_inf   = round(mean(.data$R_inf,   na.rm = TRUE), 2),
    delta_R = round(mean(.data$delta_R, na.rm = TRUE), 2),
    R0      = round(mean(.data$R0,      na.rm = TRUE), 2),
    fc_khz  = round(mean(.data$fc_hz,   na.rm = TRUE) / 1000, 1),
    alpha   = round(mean(.data$alpha,   na.rm = TRUE), 3),
    .groups = "drop"
  ) |>
  knitr::kable(
    col.names = c("Sequence", "N", "R_inf (Ohm)", "delta_R (Ohm)", "R0 (Ohm)", "fc (kHz)", "alpha"),
    caption   = "Cole model parameters at reference sequences (cohort mean). R_inf and fc are extrapolated for ~10% of fits (fc > 122 kHz)."
  )
