---
title: "Individual BIA Analysis Pipeline"
author: "Morgan Virolan"
date: "2026-05-15"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: cosmo

params:
  RUN_ID: "011BeSa_20251023"
---

# Overview

Single-run BIA/EIM analysis: load aligned `bia4` data, build reference/EX_DYN/EX_STA features, fit Cole parameters, and write cohort-ready tables.

# 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(arrow)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(glue)

options(dplyr.summarise.inform = FALSE)
source(here::here("sources", "R", "config", "amstrom_style.R"))
```

```{r controls}
RUN_ID <- params$RUN_ID

PIPELINE_MODE <- "full_run"
SHOW_MAIN_PLOT_GALLERY <- identical(PIPELINE_MODE, "full_run")
EXPORT_BIA_PLOTS <- identical(PIPELINE_MODE, "full_run")
EXPORT_ANALYSIS_TABLES <- identical(PIPELINE_MODE, "full_run")

BIA_PRIMARY_CONFIG <- "bia4"
BIA_TARGET_FREQS_HZ <- c(9760, 48800, 97600)
BIS_ALL_FREQS_HZ    <- c(4880, 9760, 14640, 19520, 24400, 29280, 34160, 39040,
                          48800, 58560, 68320, 78080, 87840, 97600, 107360, 122000)

EX_STA_WINDOW_LENGTH_S <- 1.0
EX_STA_WINDOW_STEP_S <- 1.0

critical_seqs <- c("MVC_REF", "SVC_REF", "EX_DYN", "EX_STA", "MVC_RECOV_DYN", "MVC_RECOV_STA")
```

# 2. Load exported data

```{r load_data}
# Load the aligned bia4 table + master grid + participant info for this run.
export_root <- here::here("results", "DATA_EXPORT")
run_directory <- file.path(export_root, RUN_ID)

master_file_path <- file.path(run_directory, "master_index_grid.parquet")
bia_file_path <- file.path(run_directory, "preprocessed_bia4.parquet")
participant_file_path <- file.path(run_directory, "participant.parquet")

master_table <- arrow::read_parquet(
  master_file_path,
  col_select = c("time_index", "time_ref_s", "SEQ_index", "SEQ", "VC", "VC_count")
)

bia_table <- arrow::read_parquet(bia_file_path)

participant_table <- arrow::read_parquet(participant_file_path) |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, .before = 1)

participant_id <- participant_table$participant_id[[1]]

bia_analysis_table <- bia_table |>
  dplyr::left_join(
    master_table |>
      dplyr::select(time_index, time_ref_s),
    by = "time_index"
  )

cat("RUN_ID:", RUN_ID, "\n")
cat("participant_id:", participant_id, "\n")
cat("nrow(master_table):", nrow(master_table), "\n")
cat("nrow(bia_table):   ", nrow(bia_table), "\n")
cat("nrow(bia_analysis_table):", nrow(bia_analysis_table), "\n")
```

# 3. Sequence validity gate

```{r check_bia_sequence_validity}
# Validity gate: stop if any critical sequence is missing or flagged invalid.
bia_validity <- bia_table |>
  dplyr::distinct(.data$SEQ, .data$sequence_valid)

print(bia_validity)

bia_validity_crit <- bia_validity |>
  dplyr::filter(.data$SEQ %in% critical_seqs)

missing_critical_seqs <- setdiff(critical_seqs, bia_validity_crit$SEQ)
if (length(missing_critical_seqs) > 0L) {
  stop(glue(
    "BIA sequence validity check FAILED for {RUN_ID}: missing critical sequence(s): ",
    paste(missing_critical_seqs, collapse = ", ")
  ))
}

invalid_bia_seqs <- bia_validity_crit |>
  dplyr::filter(.data$sequence_valid == 0)

if (nrow(invalid_bia_seqs) > 0L) {
  stop(glue(
    "BIA sequence validity check FAILED for {RUN_ID}:\n",
    paste(capture.output(print(invalid_bia_seqs)), collapse = "\n")
  ))
}

cat("BIA sequence validity OK for all critical sequences.\n")
```

# 4. Build the long BIA analysis table

```{r build_bia_long_table}
# BIA is much sparser than EMG. 
# keep all BIA samples that already fall inside the contraction-labeled rows,
# then summarise within VC_count or within 1 s static windows.

build_bia_long <- function(df, config, freq_hz_vec) {
  return(dplyr::bind_rows(lapply(freq_hz_vec, function(freq_hz) {
    freq_tag <- gsub("\\.", "p", format(freq_hz, scientific = FALSE, trim = TRUE))
    r_col <- glue("{config}_R_ohm__f_{freq_tag}Hz")
    x_col <- glue("{config}_Xc_ohm__f_{freq_tag}Hz")
    pha_col <- glue("{config}_PhA_deg__f_{freq_tag}Hz")

    df |>
      dplyr::mutate(
        freq_hz = freq_hz,
        R = .data[[r_col]],
        Xc = .data[[x_col]],
        PhA = .data[[pha_col]],
        absZ = sqrt(.data$R^2 + .data$Xc^2)
      ) |>
      dplyr::select(
        time_index, time_ref_s,
        SEQ_index, SEQ, VC, VC_count,
        sequence_valid,
        freq_hz,
        R, Xc, PhA, absZ
      )
  })) |>
    tidyr::pivot_longer(
      cols = c("R", "Xc", "PhA", "absZ"),
      names_to = "summary_variable",
      values_to = "summary_value"
    ) |>
    dplyr::mutate(
      summary_unit = dplyr::case_when(
        .data$summary_variable %in% c("R", "Xc", "absZ") ~ "ohm",
        .data$summary_variable == "PhA" ~ "deg",
        TRUE ~ NA_character_
      ),
      freq_label = factor(
        .data$freq_hz,
        levels = freq_hz_vec,
        labels = vapply(freq_hz_vec, label_bia_frequency, character(1))
      )
    ))
}

bia_long_table <- build_bia_long(
  df = bia_analysis_table,
  config = BIA_PRIMARY_CONFIG,
  freq_hz_vec = BIS_ALL_FREQS_HZ
)
```

# 5. Reference and checkpoint tables

```{r build_reference_checkpoint_feature_table}
# Reference + checkpoint table: one row per sequence x VC_count x frequency x variable.
# BIA is sparse (~2-3 Hz) so we just average the contraction-labeled samples within each
# group rather than windowing. The reference_level labels (30/50/70/100%) come from the
# SVC_REF rep order, with MVC_REF standing in as the 100% anchor.
reference_checkpoint_feature_table <- bia_long_table |>
  dplyr::filter(.data$SEQ %in% c("SVC_REF", "MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")) |>
  dplyr::filter(.data$VC == 1) |>
  dplyr::group_by(
    .data$SEQ, .data$VC_count,
    .data$freq_hz, .data$freq_label,
    .data$summary_variable, .data$summary_unit
  ) |>
  dplyr::summarise(
    summary_value = mean(.data$summary_value, na.rm = TRUE),  # mean BIA value over the contraction
    time_mid_s = mean(.data$time_ref_s, na.rm = TRUE),        # contraction midpoint time
    n_samples = sum(is.finite(.data$summary_value)),          # BIA samples averaged
    .groups = "drop"
  ) |>
  dplyr::mutate(
    RUN_ID = .env$RUN_ID,
    participant_id = .env$participant_id,
    analysis_family = "bia",
    analysis_block = "bia_reference_checkpoint",
    source_sequence = .data$SEQ,
    reference_level = dplyr::case_when(
      .data$SEQ == "SVC_REF" & .data$VC_count == 1L ~ "30%",
      .data$SEQ == "SVC_REF" & .data$VC_count == 2L ~ "50%",
      .data$SEQ == "SVC_REF" & .data$VC_count == 3L ~ "70%",
      .data$SEQ == "MVC_REF" ~ "100%",
      TRUE ~ NA_character_
    ),
    state_label = dplyr::case_when(
      .data$SEQ == "SVC_REF" ~ paste0("SVC_", .data$reference_level),
      TRUE ~ .data$SEQ
    ),
    evidence_status = "provisional",
    .before = 1
  ) |>
  dplyr::select(
    RUN_ID, participant_id,
    analysis_family, analysis_block,
    source_sequence, state_label,
    reference_level,
    VC_count, freq_hz, freq_label,
    summary_variable, summary_unit,
    summary_value,
    time_mid_s, n_samples,
    evidence_status
  ) |>
  dplyr::arrange(.data$summary_variable, .data$freq_hz, .data$source_sequence, .data$VC_count)
```

# 6. EX_DYN trajectory table

```{r build_ex_dyn_feature_table}
# EX_DYN trajectory: same per-contraction averaging, one row per rep (capped at 60).
bia_ex_dyn_feature_table <- bia_long_table |>
  dplyr::filter(.data$SEQ == "EX_DYN", .data$VC == 1) |>
  dplyr::group_by(
    .data$SEQ, .data$VC_count,
    .data$freq_hz, .data$freq_label,
    .data$summary_variable, .data$summary_unit
  ) |>
  dplyr::summarise(
    summary_value = mean(.data$summary_value, na.rm = TRUE),  # mean BIA value over the contraction
    time_mid_s = mean(.data$time_ref_s, na.rm = TRUE),        # contraction midpoint time
    n_samples = sum(is.finite(.data$summary_value)),          # BIA samples averaged
    .groups = "drop"
  ) |>
  dplyr::mutate(
    RUN_ID = .env$RUN_ID,
    participant_id = .env$participant_id,
    analysis_family = "bia",
    analysis_block = "bia_ex_dyn",
    source_sequence = .data$SEQ,
    evidence_status = "provisional",
    .before = 1
  ) |>
  dplyr::filter(.data$VC_count <= 60L) |>
  dplyr::arrange(.data$summary_variable, .data$freq_hz, .data$VC_count)
```

# 7. EX_STA window table

```{r build_ex_sta_feature_table}
# EX_STA has no discrete reps, so bin the static hold into fixed 1 s windows and average the
# BIA samples landing in each. Same time-binning the EMG and NIRS pipelines use -> the three
# modalities share window_id, which is what makes the multimodal joins line up later.
ex_sta_samples <- bia_long_table |>
  dplyr::filter(.data$SEQ == "EX_STA", .data$VC == 1) |>
  dplyr::arrange(.data$time_ref_s)

if (nrow(ex_sta_samples) == 0L) {
  stop("EX_STA contains no contraction-labeled BIA samples after filtering.")
}

t_start <- min(ex_sta_samples$time_ref_s)
t_end <- max(ex_sta_samples$time_ref_s)
window_start_s <- seq(t_start, t_end - EX_STA_WINDOW_LENGTH_S, by = EX_STA_WINDOW_STEP_S)

ex_sta_window_table <- tibble::tibble(
  window_id = seq_along(window_start_s),
  window_start_s = window_start_s,
  window_end_s = window_start_s + EX_STA_WINDOW_LENGTH_S,
  window_mid_s = window_start_s + EX_STA_WINDOW_LENGTH_S / 2
)

bia_ex_sta_feature_table <- ex_sta_samples |>
  dplyr::mutate(
    window_id = floor((.data$time_ref_s - t_start) / EX_STA_WINDOW_STEP_S) + 1L
  ) |>
  dplyr::filter(
    .data$window_id >= 1L,
    .data$window_id <= nrow(ex_sta_window_table)
  ) |>
  dplyr::left_join(ex_sta_window_table, by = "window_id") |>
  dplyr::group_by(
    .data$window_id, .data$window_mid_s,
    .data$freq_hz, .data$freq_label,
    .data$summary_variable, .data$summary_unit
  ) |>
  dplyr::summarise(
    summary_value = mean(.data$summary_value, na.rm = TRUE),  # mean BIA value in this 1s window
    n_samples = sum(is.finite(.data$summary_value)),          # BIA samples averaged
    .groups = "drop"
  ) |>
  dplyr::arrange(.data$window_id, .data$freq_hz, .data$summary_variable)
```

```{r filter_ex_sta_onset_windows}
# Common onset keeps the EX_STA denominator aligned with EMG and NIRS exports.
EX_STA_COMMON_ONSET_WINDOW_ID <- 2L

ex_sta_first_retained_window_id  <- EX_STA_COMMON_ONSET_WINDOW_ID
ex_sta_first_retained_window_mid_s <- bia_ex_sta_feature_table |>
  dplyr::distinct(.data$window_id, .data$window_mid_s) |>
  dplyr::filter(.data$window_id == ex_sta_first_retained_window_id) |>
  dplyr::pull(.data$window_mid_s)

window_index_lookup <- bia_ex_sta_feature_table |>
  dplyr::distinct(.data$window_id, .data$window_mid_s) |>
  dplyr::filter(.data$window_id >= ex_sta_first_retained_window_id) |>
  dplyr::arrange(.data$window_id) |>
  dplyr::mutate(
    window_index             = dplyr::row_number(),
    window_time_from_start_s = .data$window_mid_s - ex_sta_first_retained_window_mid_s
  ) |>
  dplyr::select(window_id, window_index, window_time_from_start_s)

bia_ex_sta_feature_table <- bia_ex_sta_feature_table |>
  dplyr::filter(.data$window_id >= ex_sta_first_retained_window_id) |>
  dplyr::left_join(window_index_lookup, by = "window_id") |>
  dplyr::mutate(
    RUN_ID = .env$RUN_ID,
    participant_id = .env$participant_id,
    analysis_family = "bia",
    analysis_block = "bia_ex_sta",
    source_sequence = "EX_STA",
    evidence_status = "provisional",
    .before = 1
  ) |>
  dplyr::select(
    RUN_ID, participant_id,
    analysis_family, analysis_block,
    source_sequence,
    window_id, window_index,
    window_mid_s, window_time_from_start_s,
    freq_hz, freq_label,
    summary_variable, summary_unit,
    summary_value,
    n_samples,
    evidence_status
  ) |>
  dplyr::arrange(.data$summary_variable, .data$freq_hz, .data$window_index)
```

# 7b. Cole model fitting

```{r fit_cole_model, message=FALSE}
# Cole model: Z(f) = R_inf + delta_R / (1 + (j·f/fc)^alpha)
# Xc stored as negative (capacitive); model imaginary part is also negative — used as-is.
# Drop 4.88–19.52 kHz range entirely: inductive parasitics corrupt Xc at these frequencies
# (CV=32% at 9.76 kHz vs 2.2% at 48.8 kHz; positive-Xc hook artifacts at 9.76/14.64/19.52 kHz).
# Cole fitting requires both R and Xc clean — cannot keep R-only at contaminated frequency points.
# ML independently keeps R/|Z| at all 16 frequencies (independent columns, no arc constraint).
all_sweep_freqs_hz <- sort(as.integer(
  sub("bia4_R_ohm__f_(\\d+)Hz", "\\1",
      grep("^bia4_R_ohm__f_\\d+Hz$", names(bia_analysis_table), value = TRUE))
))
cole_fit_freqs_hz <- all_sweep_freqs_hz[all_sweep_freqs_hz >= 24400L]

cole_predict_arc <- function(freq_hz, R_inf, delta_R, fc_hz, alpha) {
  u        <- (freq_hz / fc_hz)^alpha
  cos_a    <- cos(alpha * pi / 2)
  sin_a    <- sin(alpha * pi / 2)
  denom_sq <- (1 + u * cos_a)^2 + (u * sin_a)^2
  return(list(
    R  = R_inf + delta_R * (1 + u * cos_a) / denom_sq,
    Xc = -delta_R * u * sin_a / denom_sq
  ))
}

cole_sse_fn <- function(params, freq_hz, R_obs, Xc_obs) {
  pred <- cole_predict_arc(freq_hz, params[1L], params[2L], params[3L], params[4L])
  return(sum((R_obs - pred$R)^2 + (Xc_obs - pred$Xc)^2))
}

fit_cole_arc <- function(R_obs, Xc_obs, freqs, n_starts = 10L) {
  valid <- is.finite(R_obs) & is.finite(Xc_obs)
  if (sum(valid) < 4L) {
    return(tibble::tibble(R_inf = NA_real_, delta_R = NA_real_, R0 = NA_real_,
                          fc_hz = NA_real_, alpha = NA_real_, fit_sse = NA_real_,
                          fit_n_freq = sum(valid)))
  }
  R_obs <- R_obs[valid]; Xc_obs <- Xc_obs[valid]; freqs <- freqs[valid]

  R_min        <- min(R_obs)
  delta_R_init <- max(max(R_obs) - R_min, 1.0)
  best_par <- NULL; best_sse <- Inf

  # Multi-start to dodge local minima: start 1 is a sensible fixed guess, the rest random
  # restarts. Keep whichever run converged to the lowest SSE.
  for (i in seq_len(n_starts)) {
    if (i == 1L) {
      par_init <- c(R_min * 0.9, delta_R_init * 1.1, 70000, 0.75)
    } else {
      par_init <- c(
        stats::runif(1L, 0.6 * R_min, R_min),
        stats::runif(1L, 0.5 * delta_R_init, 2.0 * delta_R_init),
        10^stats::runif(1L, log10(9760), log10(122000)),
        stats::runif(1L, 0.50, 0.97)
      )
    }
    res <- tryCatch(
      stats::optim(
        par    = par_init,
        fn     = cole_sse_fn,
        method = "L-BFGS-B",
        lower  = c(0.01, 0.01, 1e3, 0.10),
        upper  = c(1e4,  1e4,  1e7, 1.00),
        freq_hz = freqs, R_obs = R_obs, Xc_obs = Xc_obs
      ),
      error = function(e) NULL
    )
    if (!is.null(res) && res$convergence %in% c(0L, 1L) && res$value < best_sse) {
      best_sse <- res$value
      best_par <- res$par
    }
  }

  if (is.null(best_par)) {
    return(tibble::tibble(R_inf = NA_real_, delta_R = NA_real_, R0 = NA_real_,
                          fc_hz = NA_real_, alpha = NA_real_, fit_sse = NA_real_,
                          fit_n_freq = as.integer(sum(valid))))
  }
  return(tibble::tibble(
    R_inf      = best_par[1L],
    delta_R    = best_par[2L],
    R0         = best_par[1L] + best_par[2L],
    fc_hz      = best_par[3L],
    alpha      = best_par[4L],
    fit_sse    = best_sse,
    fit_n_freq = as.integer(sum(valid))
  ))
}

set.seed(42L)

# --- Reference / checkpoint sequences ---
cole_input_ref <- bia_analysis_table |>
  dplyr::filter(
    .data$SEQ %in% c("SVC_REF", "MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA"),
    .data$VC == 1L
  ) |>
  dplyr::group_by(.data$SEQ, .data$VC_count) |>
  dplyr::summarise(
    dplyr::across(
      dplyr::matches("^bia4_(R|Xc)_ohm__f_\\d+Hz$"),
      ~ mean(.x, na.rm = TRUE)
    ),
    n_samples = dplyr::n(),
    .groups = "drop"
  )

cole_reference_table <- dplyr::bind_rows(lapply(seq_len(nrow(cole_input_ref)), function(i) {
  row    <- cole_input_ref[i, ]
  R_obs  <- as.numeric(row[paste0("bia4_R_ohm__f_",  cole_fit_freqs_hz, "Hz")])
  Xc_obs <- as.numeric(row[paste0("bia4_Xc_ohm__f_", cole_fit_freqs_hz, "Hz")])
  dplyr::bind_cols(
    tibble::tibble(
      RUN_ID          = RUN_ID,
      participant_id  = participant_id,
      source_sequence = row$SEQ[[1L]],
      VC_count        = row$VC_count[[1L]],
      n_samples       = row$n_samples[[1L]]
    ),
    fit_cole_arc(R_obs, Xc_obs, cole_fit_freqs_hz)
  )
}))

# --- EX_DYN ---
cole_input_exdyn <- bia_analysis_table |>
  dplyr::filter(.data$SEQ == "EX_DYN", .data$VC == 1L, .data$VC_count <= 60L) |>
  dplyr::group_by(.data$VC_count) |>
  dplyr::summarise(
    dplyr::across(
      dplyr::matches("^bia4_(R|Xc)_ohm__f_\\d+Hz$"),
      ~ mean(.x, na.rm = TRUE)
    ),
    n_samples = dplyr::n(),
    .groups = "drop"
  )

cole_ex_dyn_table <- dplyr::bind_rows(lapply(seq_len(nrow(cole_input_exdyn)), function(i) {
  row    <- cole_input_exdyn[i, ]
  R_obs  <- as.numeric(row[paste0("bia4_R_ohm__f_",  cole_fit_freqs_hz, "Hz")])
  Xc_obs <- as.numeric(row[paste0("bia4_Xc_ohm__f_", cole_fit_freqs_hz, "Hz")])
  dplyr::bind_cols(
    tibble::tibble(
      RUN_ID          = RUN_ID,
      participant_id  = participant_id,
      source_sequence = "EX_DYN",
      VC_count        = row$VC_count[[1L]],
      n_samples       = row$n_samples[[1L]]
    ),
    fit_cole_arc(R_obs, Xc_obs, cole_fit_freqs_hz)
  )
}))

# --- EX_STA (onset-retained windows only) ---
# window_id assigned using same time-binning formula as build_ex_sta_feature_table.
# window_index_lookup and ex_sta_first_retained_window_id are from filter_ex_sta_onset_windows.
cole_input_exsta_raw <- bia_analysis_table |>
  dplyr::filter(.data$SEQ == "EX_STA", .data$VC == 1L) |>
  dplyr::mutate(
    window_id = floor((.data$time_ref_s - t_start) / EX_STA_WINDOW_STEP_S) + 1L
  ) |>
  dplyr::filter(
    .data$window_id >= 1L,
    .data$window_id <= nrow(ex_sta_window_table),
    .data$window_id >= ex_sta_first_retained_window_id
  ) |>
  dplyr::group_by(.data$window_id) |>
  dplyr::summarise(
    dplyr::across(
      dplyr::matches("^bia4_(R|Xc)_ohm__f_\\d+Hz$"),
      ~ mean(.x, na.rm = TRUE)
    ),
    n_samples = dplyr::n(),
    .groups = "drop"
  ) |>
  dplyr::inner_join(window_index_lookup, by = "window_id")

cole_ex_sta_table <- dplyr::bind_rows(lapply(seq_len(nrow(cole_input_exsta_raw)), function(i) {
  row    <- cole_input_exsta_raw[i, ]
  R_obs  <- as.numeric(row[paste0("bia4_R_ohm__f_",  cole_fit_freqs_hz, "Hz")])
  Xc_obs <- as.numeric(row[paste0("bia4_Xc_ohm__f_", cole_fit_freqs_hz, "Hz")])
  dplyr::bind_cols(
    tibble::tibble(
      RUN_ID                   = RUN_ID,
      participant_id           = participant_id,
      source_sequence          = "EX_STA",
      window_id                = row$window_id[[1L]],
      window_index             = row$window_index[[1L]],
      window_time_from_start_s = row$window_time_from_start_s[[1L]],
      n_samples                = row$n_samples[[1L]]
    ),
    fit_cole_arc(R_obs, Xc_obs, cole_fit_freqs_hz)
  )
}))

cole_reference_table
cole_ex_dyn_table |> dplyr::select(source_sequence, VC_count, R_inf, delta_R, R0, fc_hz, alpha, fit_sse)
cole_ex_sta_table |> dplyr::select(window_index, R_inf, delta_R, R0, fc_hz, alpha, fit_sse)
```

# 8. Run-level summary table

```{r build_bia_run_summary_table}
# Run-level scalars in three flavours: linear slope over time, last-minus-first delta, and
# checkpoint-minus-baseline (recovery vs MVC_REF). One value per frequency x variable.
make_linear_slope_summary <- function(df, x_col, analysis_block, source_sequence) {
  return(split(df, interaction(df$freq_hz, df$summary_variable, drop = TRUE)) |>
    lapply(function(group_df) {
      group_df <- group_df |>
        dplyr::arrange(.data[[x_col]]) |>
        dplyr::filter(is.finite(.data[[x_col]]), is.finite(.data$summary_value))

      if (nrow(group_df) < 3L || dplyr::n_distinct(group_df[[x_col]]) < 3L) {
        return(NULL)
      }

      fit_df <- group_df |>
        dplyr::mutate(x_value = .data[[x_col]])

      fit <- stats::lm(summary_value ~ x_value, data = fit_df)
      tibble::tibble(
        analysis_block = analysis_block,
        freq_hz = group_df$freq_hz[[1]],
        summary_variable = group_df$summary_variable[[1]],
        summary_unit = group_df$summary_unit[[1]],
        summary_type = paste0(x_col, "_linear_slope"),
        source_sequence = source_sequence,
        summary_value = unname(stats::coef(fit)[[2]]),
        n_points = nrow(group_df)
      )
    }) |>
    dplyr::bind_rows())
}

make_last_minus_first_summary <- function(df, x_col, analysis_block, source_sequence) {
  return(split(df, interaction(df$freq_hz, df$summary_variable, drop = TRUE)) |>
    lapply(function(group_df) {
      group_df <- group_df |>
        dplyr::arrange(.data[[x_col]]) |>
        dplyr::filter(is.finite(.data[[x_col]]), is.finite(.data$summary_value))

      if (nrow(group_df) < 2L) {
        return(NULL)
      }

      tibble::tibble(
        analysis_block = analysis_block,
        freq_hz = group_df$freq_hz[[1]],
        summary_variable = group_df$summary_variable[[1]],
        summary_unit = group_df$summary_unit[[1]],
        summary_type = "last_minus_first",
        source_sequence = source_sequence,
        summary_value = dplyr::last(group_df$summary_value) - dplyr::first(group_df$summary_value),
        n_points = nrow(group_df)
      )
    }) |>
    dplyr::bind_rows())
}

make_checkpoint_delta_summary <- function(df, checkpoint_seq) {
  baseline_df <- df |>
    dplyr::filter(.data$source_sequence == "MVC_REF") |>
    dplyr::group_by(.data$freq_hz, .data$summary_variable, .data$summary_unit) |>
    dplyr::summarise(baseline_value = mean(.data$summary_value, na.rm = TRUE), .groups = "drop")

  checkpoint_df <- df |>
    dplyr::filter(.data$source_sequence == checkpoint_seq) |>
    dplyr::group_by(.data$freq_hz, .data$summary_variable, .data$summary_unit) |>
    dplyr::summarise(checkpoint_value = mean(.data$summary_value, na.rm = TRUE), .groups = "drop")

  return(dplyr::inner_join(
    baseline_df,
    checkpoint_df,
    by = c("freq_hz", "summary_variable", "summary_unit")
  ) |>
    dplyr::mutate(
      analysis_block = "bia_reference_checkpoint_summary",
      summary_type = paste0(tolower(checkpoint_seq), "_minus_mvc_ref"),
      source_sequence = checkpoint_seq,
      summary_value = .data$checkpoint_value - .data$baseline_value,
      n_points = NA_integer_
    ) |>
    dplyr::select(
      analysis_block,
      freq_hz,
      summary_variable,
      summary_unit,
      summary_type,
      source_sequence,
      summary_value,
      n_points
    ))
}

bia_run_summary_table <- dplyr::bind_rows(
  make_linear_slope_summary(bia_ex_dyn_feature_table, "VC_count", "bia_ex_dyn_run_summary", "EX_DYN"),
  make_last_minus_first_summary(bia_ex_dyn_feature_table, "VC_count", "bia_ex_dyn_run_summary", "EX_DYN"),
  make_linear_slope_summary(bia_ex_sta_feature_table, "window_index", "bia_ex_sta_run_summary", "EX_STA"),
  make_last_minus_first_summary(bia_ex_sta_feature_table, "window_index", "bia_ex_sta_run_summary", "EX_STA"),
  make_checkpoint_delta_summary(reference_checkpoint_feature_table, "MVC_RECOV_DYN"),
  make_checkpoint_delta_summary(reference_checkpoint_feature_table, "MVC_RECOV_STA")
) |>
  dplyr::mutate(
    RUN_ID = .env$RUN_ID,
    participant_id = .env$participant_id,
    analysis_family = "bia",
    evidence_status = "provisional",
    source_sequence = .data$source_sequence,
    .before = 1
  ) |>
  dplyr::select(
    RUN_ID, participant_id,
    analysis_family, analysis_block,
    summary_variable, summary_value,
    summary_unit, summary_type,
    source_sequence, evidence_status,
    freq_hz, n_points
  ) |>
  dplyr::arrange(.data$analysis_block, .data$summary_variable, .data$freq_hz, .data$summary_type)
```

# 9. Diagnostic plots

```{r compute_reldev}
# RelDev = (value - ref_mean) / ref_mean * 100.
# EX_DYN and reference checkpoints: MVC_REF (100% pre-fatigue) — same intensity as EX_DYN.
# EX_STA: SVC_70% (actual measured 70% pre-fatigue) — same intensity as EX_STA.
# Using MVC_REF for EX_STA would confound the activation-level difference with fatigue !
mvc_ref_baseline <- reference_checkpoint_feature_table |>
  dplyr::filter(.data$source_sequence == "MVC_REF") |>
  dplyr::group_by(.data$freq_hz, .data$summary_variable) |>
  dplyr::summarise(ref_mean = mean(.data$summary_value, na.rm = TRUE), .groups = "drop")

svc_70_baseline <- reference_checkpoint_feature_table |>
  dplyr::filter(.data$reference_level == "70%") |>
  dplyr::group_by(.data$freq_hz, .data$summary_variable) |>
  dplyr::summarise(ref_mean = mean(.data$summary_value, na.rm = TRUE), .groups = "drop")

join_reldev <- function(df, baseline) {
  return(df |>
    dplyr::left_join(baseline, by = c("freq_hz", "summary_variable")) |>
    dplyr::mutate(reldev_pct = (.data$summary_value - .data$ref_mean) / .data$ref_mean * 100) |>  # % deviation from the matched reference
    dplyr::select(-ref_mean))
}

reference_checkpoint_feature_table <- join_reldev(reference_checkpoint_feature_table, mvc_ref_baseline)
bia_ex_dyn_feature_table           <- join_reldev(bia_ex_dyn_feature_table, mvc_ref_baseline)
bia_ex_sta_feature_table           <- join_reldev(bia_ex_sta_feature_table, svc_70_baseline)

# reldev_seq_pct: within-sequence relative change, baselined on the first 3 contractions/windows.
# This is the first-to-last RelDev convention -- change measured within the exercise sequence itself.
add_reldev_seq <- function(df, time_var) {
  seq_ref <- df |>
    dplyr::filter(.data[[time_var]] <= 3L) |>
    dplyr::group_by(.data$freq_hz, .data$summary_variable) |>
    dplyr::summarise(seq_ref_mean = mean(.data$summary_value, na.rm = TRUE), .groups = "drop")
  return(df |>
    dplyr::left_join(seq_ref, by = c("freq_hz", "summary_variable")) |>
    dplyr::mutate(reldev_seq_pct = (.data$summary_value - .data$seq_ref_mean) / .data$seq_ref_mean * 100) |>  # % change vs the sequence's first 3 points
    dplyr::select(-seq_ref_mean))
}

bia_ex_dyn_feature_table <- add_reldev_seq(bia_ex_dyn_feature_table, "VC_count")
bia_ex_sta_feature_table <- add_reldev_seq(bia_ex_sta_feature_table, "window_index")
```

```{r build_bia_plots}
reference_state_levels <- c("SVC_30%", "SVC_50%", "SVC_70%", "MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")

bia_ex_dyn_plot_tbl <- dplyr::filter(bia_ex_dyn_feature_table, .data$freq_hz %in% BIA_TARGET_FREQS_HZ)
bia_ex_sta_plot_tbl <- dplyr::filter(bia_ex_sta_feature_table, .data$freq_hz %in% BIA_TARGET_FREQS_HZ)

# SVC_REF rows already have distinct state_labels per intensity (30/50/70%)
# MVC-type sequences have multiple VC_counts at the same intensity (so we average across them)
bia_reference_plot_table <- dplyr::bind_rows(
  reference_checkpoint_feature_table |>
    dplyr::filter(.data$source_sequence == "SVC_REF"),
  reference_checkpoint_feature_table |>
    dplyr::filter(.data$source_sequence != "SVC_REF") |>
    dplyr::group_by(
      .data$source_sequence, .data$state_label,
      .data$freq_hz, .data$freq_label,
      .data$summary_variable, .data$summary_unit
    ) |>
    dplyr::summarise(
      summary_value = mean(.data$summary_value, na.rm = TRUE),
      n_samples     = sum(.data$n_samples, na.rm = TRUE),
      .groups = "drop"
    )
) |>
  dplyr::filter(.data$freq_hz %in% BIA_TARGET_FREQS_HZ) |>
  dplyr::mutate(
    state_label = factor(.data$state_label, levels = reference_state_levels)
  )

bia_reference_plot <- ggplot(
  bia_reference_plot_table,
  aes(x = .data$state_label, y = .data$summary_value, colour = .data$freq_label, group = .data$freq_label)
) +
  geom_line(linetype = "dotted", linewidth = 0.35) +
  geom_point(shape = 15, size = 2.2) +
  facet_wrap(~ summary_variable, scales = "free_y", ncol = 2) +
  labs(
    title = paste0("BIA reference and checkpoint overview - ", RUN_ID),
    x = NULL,
    y = NULL,
    colour = "Frequency"
  ) +
  theme_amstrom_hgrid +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

bia_ex_dyn_plot <- ggplot(
  bia_ex_dyn_plot_tbl,
  aes(x = .data$VC_count, y = .data$summary_value, colour = .data$freq_label, group = .data$freq_label)
) +
  geom_line(linetype = "dotted", linewidth = 0.35) +
  geom_point(shape = 15, size = 2.2) +
  facet_wrap(~ summary_variable, scales = "free_y", ncol = 2) +
  labs(
    title = paste0("BIA EX_DYN trajectories - ", RUN_ID),
    x = "Repetition (VC_count)",
    y = NULL,
    colour = "Frequency"
  ) +
  theme_amstrom_hgrid

bia_ex_sta_plot <- ggplot(
  bia_ex_sta_plot_tbl,
  aes(x = .data$window_index, y = .data$summary_value, colour = .data$freq_label, group = .data$freq_label)
) +
  geom_line(linetype = "dotted", linewidth = 0.35) +
  geom_point(shape = 15, size = 2.2) +
  facet_wrap(~ summary_variable, scales = "free_y", ncol = 2) +
  labs(
    title = paste0("BIA EX_STA trajectories - ", RUN_ID),
    x = "Window index",
    y = NULL,
    colour = "Frequency"
  ) +
  theme_amstrom_hgrid

plot_export_map <- list(
  reference_checkpoint = bia_reference_plot,
  ex_dyn = bia_ex_dyn_plot,
  ex_sta = bia_ex_sta_plot
)
```

```{r build_reldev_plots}
bia_ex_dyn_plot_tbl <- dplyr::filter(bia_ex_dyn_feature_table, .data$freq_hz %in% BIA_TARGET_FREQS_HZ)
bia_ex_sta_plot_tbl <- dplyr::filter(bia_ex_sta_feature_table, .data$freq_hz %in% BIA_TARGET_FREQS_HZ)

bia_reference_reldev_plot_table <- dplyr::bind_rows(
  reference_checkpoint_feature_table |>
    dplyr::filter(.data$source_sequence == "SVC_REF"),
  reference_checkpoint_feature_table |>
    dplyr::filter(.data$source_sequence != "SVC_REF") |>
    dplyr::group_by(
      .data$source_sequence, .data$state_label,
      .data$freq_hz, .data$freq_label,
      .data$summary_variable, .data$summary_unit
    ) |>
    dplyr::summarise(reldev_pct = mean(.data$reldev_pct, na.rm = TRUE), .groups = "drop")
) |>
  dplyr::filter(.data$freq_hz %in% BIA_TARGET_FREQS_HZ) |>
  dplyr::mutate(state_label = factor(.data$state_label, levels = reference_state_levels))

bia_reference_reldev_plot <- ggplot(
  bia_reference_reldev_plot_table,
  aes(x = .data$state_label, y = .data$reldev_pct, colour = .data$freq_label, group = .data$freq_label)
) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  geom_line(linetype = "dotted", linewidth = 0.35) +
  geom_point(shape = 15, size = 2.2) +
  facet_wrap(~ summary_variable, scales = "free_y", ncol = 2) +
  labs(
    title = paste0("BIA reference RelDev from MVC_REF - ", RUN_ID),
    x = NULL,
    y = "RelDev (%)",
    colour = "Frequency"
  ) +
  theme_amstrom_hgrid +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

bia_ex_dyn_reldev_plot <- ggplot(
  bia_ex_dyn_plot_tbl,
  aes(x = .data$VC_count, y = .data$reldev_pct, colour = .data$freq_label, group = .data$freq_label)
) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  geom_line(linetype = "dotted", linewidth = 0.35) +
  geom_point(shape = 15, size = 2.2) +
  facet_wrap(~ summary_variable, scales = "free_y", ncol = 2) +
  labs(
    title = paste0("BIA EX_DYN RelDev from MVC_REF - ", RUN_ID),
    x = "Repetition",
    y = "RelDev (%)",
    colour = "Frequency"
  ) +
  theme_amstrom_hgrid

bia_ex_sta_reldev_plot <- ggplot(
  bia_ex_sta_plot_tbl,
  aes(x = .data$window_index, y = .data$reldev_pct, colour = .data$freq_label, group = .data$freq_label)
) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  geom_line(linetype = "dotted", linewidth = 0.35) +
  geom_point(shape = 15, size = 2.2) +
  facet_wrap(~ summary_variable, scales = "free_y", ncol = 2) +
  labs(
    title = paste0("BIA EX_STA RelDev from MVC_REF - ", RUN_ID),
    x = "Window index",
    y = "RelDev (%)",
    colour = "Frequency"
  ) +
  theme_amstrom_hgrid

plot_export_map[["reference_reldev"]] <- bia_reference_reldev_plot
plot_export_map[["ex_dyn_reldev"]]    <- bia_ex_dyn_reldev_plot
plot_export_map[["ex_sta_reldev"]]    <- bia_ex_sta_reldev_plot
```

```{r bia_nyquist_plot}
# Nyquist arcs (R vs -Xc across the full sweep), one panel per sequence. Each panel holds a
# few snapshots coloured fresh -> late: reference sequences get one arc per VC_count, EX_DYN
# picks 5 reps spread across the set, EX_STA picks 5 equally spaced timepoints.
nyquist_panel_seqs  <- c("MVC_REF", "EX_DYN", "EX_STA", "SVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")
reference_seqs_nyq  <- c("MVC_REF", "SVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")
ex_dyn_target_reps  <- c(1L, 16L, 31L, 46L, 60L)
ex_sta_time_labels  <- c("t0", "t25", "t50", "t75", "t100")

sweep_freqs_hz <- sort(as.integer(
  sub("bia4_R_ohm__f_(\\d+)Hz", "\\1",
      grep("^bia4_R_ohm__f_\\d+Hz$", names(bia_analysis_table), value = TRUE))
))

nyquist_seq_index_map <- bia_analysis_table |>
  dplyr::distinct(.data$SEQ, .data$SEQ_index) |>
  dplyr::filter(.data$SEQ %in% nyquist_panel_seqs) |>
  tibble::deframe()

extract_nyquist_arc <- function(row_df, freqs) {
  return(tibble::tibble(
    freq_hz  = freqs,
    R        = vapply(freqs, function(f) as.numeric(row_df[[paste0("bia4_R_ohm__f_", f, "Hz")]]), numeric(1)),
    minus_Xc = vapply(freqs, function(f) -as.numeric(row_df[[paste0("bia4_Xc_ohm__f_", f, "Hz")]]), numeric(1))
  ))
}

pick_mid_row <- function(df) return(df[max(1L, ceiling(nrow(df) / 2L)), ])

nyquist_data <- dplyr::bind_rows(lapply(nyquist_panel_seqs, function(seq_name) {
  if (!seq_name %in% names(nyquist_seq_index_map)) return(NULL)
  seq_idx <- nyquist_seq_index_map[[seq_name]]
  df_seq  <- bia_analysis_table[
    as.integer(bia_analysis_table$SEQ_index) == seq_idx &
    as.integer(bia_analysis_table$VC_count)  >  0L, ]
  if (nrow(df_seq) == 0L) return(NULL)

  curves <- if (seq_name %in% reference_seqs_nyq) {
    vc_ids <- sort(unique(as.integer(df_seq$VC_count)))
    lapply(seq_along(vc_ids), function(i) {
      df_block <- df_seq[as.integer(df_seq$VC_count) == vc_ids[i], ]
      extract_nyquist_arc(pick_mid_row(df_block), sweep_freqs_hz) |>
        dplyr::mutate(
          snapshot_label = as.character(vc_ids[i]),
          curve_position = (i - 1L) / max(length(vc_ids) - 1L, 1L)
        )
    })
  } else if (seq_name == "EX_DYN") {
    vc_ids_present <- sort(unique(as.integer(df_seq$VC_count)))
    rep_ids <- ex_dyn_target_reps[ex_dyn_target_reps %in% vc_ids_present]
    if (length(rep_ids) == 0L) {
      rep_ids <- vc_ids_present[round(seq(1, length(vc_ids_present), length.out = 5))]
    }
    lapply(seq_along(rep_ids), function(i) {
      df_block <- df_seq[as.integer(df_seq$VC_count) == rep_ids[i], ]
      extract_nyquist_arc(pick_mid_row(df_block), sweep_freqs_hz) |>
        dplyr::mutate(
          snapshot_label = paste0("#", rep_ids[i]),
          curve_position = (i - 1L) / max(length(rep_ids) - 1L, 1L)
        )
    })
  } else {
    n_rows   <- nrow(df_seq)
    row_idxs <- round(seq(1, n_rows, length.out = 5))
    lapply(seq_along(row_idxs), function(i) {
      extract_nyquist_arc(df_seq[row_idxs[i], ], sweep_freqs_hz) |>
        dplyr::mutate(
          snapshot_label = ex_sta_time_labels[i],
          curve_position = (i - 1L) / 4L
        )
    })
  }

  dplyr::bind_rows(curves) |> dplyr::mutate(panel_seq = seq_name)
})) |>
  dplyr::mutate(panel_seq = factor(.data$panel_seq, levels = nyquist_panel_seqs))

bia_nyquist_plot <- ggplot(
  nyquist_data,
  aes(x = .data$R, y = .data$minus_Xc, group = .data$snapshot_label, colour = .data$curve_position)
) +
  geom_path(linewidth = 1.2) +
  facet_wrap(~ panel_seq, scales = "free", ncol = 3) +
  scale_colour_viridis_c(
    name = NULL,
    option = "plasma",
    breaks = c(0, 1),
    labels = c("fresh", "late")
  ) +
  labs(
    title    = paste0("BIA 4PT Nyquist - ", RUN_ID),
    subtitle = "Each arc: R vs -Xc across full sweep (4.9-122 kHz). Colour: temporal position in sequence.",
    x        = "Resistance R (Ohm)",
    y        = "-Reactance -Xc (Ohm)"
  ) +
  theme_amstrom_hgrid +
  theme(aspect.ratio = 1)

plot_export_map[["nyquist"]] <- bia_nyquist_plot
```

```{r show_main_plots, fig.width=11, fig.height=8}
if (SHOW_MAIN_PLOT_GALLERY) {
  for (plot_name in names(plot_export_map)) {
    print(plot_export_map[[plot_name]])
  }
}
```

```{r export_bia_plots, message=FALSE}
if (isTRUE(EXPORT_BIA_PLOTS)) {
  qc_bia_export_path <- here::here("results", "QC_EXPORT", "bia_individual_analysis",
    paste0(RUN_ID, "_bia_qc.pdf")
  )
  dir.create(dirname(qc_bia_export_path), recursive = TRUE, showWarnings = FALSE)

  grDevices::pdf(file = qc_bia_export_path, width = 11, height = 8, onefile = TRUE)
  for (plot_name in names(plot_export_map)) {
    print(plot_export_map[[plot_name]])
  }
  grDevices::dev.off()

  message("\n[BIA QC EXPORT] PDF -> ", normalizePath(qc_bia_export_path, winslash = "/", mustWork = FALSE))
}
```

# 10. Export analysis tables

```{r export_analysis_tables, message=FALSE}
EXPORT_BIA_REFERENCE_FEATURES <- TRUE
EXPORT_BIA_EX_DYN_FEATURES <- TRUE
EXPORT_BIA_EX_STA_FEATURES <- TRUE
EXPORT_BIA_RUN_SUMMARY <- TRUE
EXPORT_BIA_COLE_FEATURES <- TRUE

analysis_export_flags <- c(
  bia_reference_features = EXPORT_BIA_REFERENCE_FEATURES,
  bia_ex_dyn_features = EXPORT_BIA_EX_DYN_FEATURES,
  bia_ex_sta_features = EXPORT_BIA_EX_STA_FEATURES,
  bia_run_summary = EXPORT_BIA_RUN_SUMMARY,
  bia_cole_reference = EXPORT_BIA_COLE_FEATURES,
  bia_cole_ex_dyn = EXPORT_BIA_COLE_FEATURES,
  bia_cole_ex_sta = EXPORT_BIA_COLE_FEATURES
)

analysis_export_map <- list(
  bia_reference_features = reference_checkpoint_feature_table,
  bia_ex_dyn_features = bia_ex_dyn_feature_table,
  bia_ex_sta_features = bia_ex_sta_feature_table,
  bia_run_summary = bia_run_summary_table,
  bia_cole_reference = cole_reference_table,
  bia_cole_ex_dyn = cole_ex_dyn_table,
  bia_cole_ex_sta = cole_ex_sta_table
)

analysis_export_file_map <- c(
  bia_reference_features = "bia_reference_feature_table.parquet",
  bia_ex_dyn_features = "bia_ex_dyn_feature_table.parquet",
  bia_ex_sta_features = "bia_ex_sta_feature_table.parquet",
  bia_run_summary = "bia_run_summary_table.parquet",
  bia_cole_reference = "bia_cole_reference_table.parquet",
  bia_cole_ex_dyn = "bia_cole_ex_dyn_table.parquet",
  bia_cole_ex_sta = "bia_cole_ex_sta_table.parquet"
)

if (isTRUE(EXPORT_ANALYSIS_TABLES)) {
  selected_table_names <- names(analysis_export_flags)[analysis_export_flags]
  if (length(selected_table_names) == 0L) {
    stop("EXPORT_ANALYSIS_TABLES=TRUE but no BIA analysis table selector is set to TRUE.")
  }

  analysis_export_root <- here::here("results", "R_ANALYSIS", RUN_ID)
  dir.create(analysis_export_root, recursive = TRUE, showWarnings = FALSE)

  for (table_name in selected_table_names) {
    arrow::write_parquet(
      analysis_export_map[[table_name]],
      sink = file.path(analysis_export_root, analysis_export_file_map[[table_name]])
    )
  }

  message("\n[R ANALYSIS EXPORT] BIA tables -> ", normalizePath(analysis_export_root, winslash = "/", mustWork = FALSE))
}
```
