---
title: "Individual EMG Analysis Pipeline"
author: "Morgan Virolan"
date: "2026-04-17"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: cosmo

params:
  RUN_ID: "011BeSa_20251023"

---

# Overview
Single-run sEMG analysis: load exported data, repair rare amplitude spikes, extract 1 s features, and write cohort-ready tables.

Set `RUN_ID` and the workflow controls below, then run the notebook from top to bottom.

# Data loading
```{r setup, message=FALSE, warning=FALSE}
library(arrow)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(signal)   # butter(), filtfilt()
library(glue)


options(dplyr.summarise.inform = FALSE)
```

```{r plot_style}
col_rms          <- "#3d6b99"
col_mdf          <- "#a0682a"
col_ref_line     <- "#666666"
col_ref_fill     <- "#888888"
col_ref_pts      <- "#aaaaaa"
col_neutral_data <- "#4d4d4d"

# Keys must match the metric factor labels used in pivot_longer calls below.
col_by_metric <- c(
  "nTorque"     = col_neutral_data,
  "nRMS (a.u.)" = col_rms,
  "nMDF"        = col_mdf
)

theme_amstrom <- theme_bw(base_size = 11) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border     = element_rect(colour = "grey55", fill = NA, linewidth = 0.5),
    strip.background = element_rect(fill = "grey93", colour = NA),
    strip.text       = element_text(face = "bold", size = rel(0.88)),
    legend.position  = "top",
    legend.key.size  = unit(0.8, "lines"),
    legend.text      = element_text(size = rel(0.88)),
    axis.ticks       = element_line(colour = "grey55", linewidth = 0.4)
  )

theme_amstrom_hgrid <- theme_amstrom +
  theme(panel.grid.major.y = element_line(colour = "grey91", linewidth = 0.35))
```

```{r load_data}

# Set RUN_ID to the participant folder name (format: XXXNnPp_YYYYMMDD)
RUN_ID <- params$RUN_ID

PIPELINE_MODE <- "full_run"
APPLY_ARTIFACT_HANDLING <- TRUE
# In tuning mode, cap each table to this many seconds before artifact handling.
# Set to NULL to disable (full data). Speeds up spike_pct iteration dramatically.
TUNING_PREVIEW_SECONDS <- 120
# Set to NULL to show the full current sequence in artifact QC plots.
# In tuning mode this means the full preview sequence after any TUNING_PREVIEW_SECONDS cap.
ARTIFACT_QC_MAX_ROWS <- NULL
SHOW_TUNING_QC_PLOTS <- identical(PIPELINE_MODE, "tuning")
SHOW_MAIN_PLOT_GALLERY <- identical(PIPELINE_MODE, "full_run")
EXPORT_EMG_PLOTS <- identical(PIPELINE_MODE, "full_run")
EXPORT_ANALYSIS_TABLES <- identical(PIPELINE_MODE, "full_run")

# Display labels for plots only. Keep internal/export channel IDs unchanged.
channel_display_labels <- c(
  emg6 = "RF",
  emg8 = "VM",
  emg10 = "VL"
)

export_root <- here::here("results", "DATA_EXPORT")
run_directory <- file.path(export_root, RUN_ID)

master_file_path      <- file.path(run_directory, "master_index_grid.parquet")
emg_file_path         <- file.path(run_directory, "preprocessed_emg.parquet")
torque_file_path      <- file.path(run_directory, "preprocessed_torque.parquet")
participant_file_path <- file.path(run_directory, "participant.parquet")

master_table <- read_parquet(
  master_file_path,
  col_select = c("time_index", "time_ref_s", "SEQ_index", "SEQ", "VC", "VC_count")
)
emg_table <- read_parquet(
  emg_file_path,
  col_select = c("time_index", "emg6", "emg8", "emg10")
)
torque_table <- read_parquet(
  torque_file_path,
  col_select = c("time_index", "torque_raw")
)
participant_table <- read_parquet(participant_file_path) |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, .before = 1)

participant_id <- participant_table$participant_id[[1]]

if (nrow(master_table) != nrow(emg_table)) {
  stop("master_table and emg_table do not have the same number of rows.")
}

if (!identical(master_table$time_index, emg_table$time_index)) {
  stop("master_table and emg_table are not aligned on time_index.")
}

# Torque is sparser than the EMG/master clock, so it is joined after the alignment checks.
analysis_table <- dplyr::bind_cols(
  master_table,
  emg_table |> dplyr::select(-time_index)
) |>
  left_join(torque_table, by = "time_index")

cat("RUN_ID:", RUN_ID, "\n")
cat("participant_id:", participant_id, "\n")
cat("nrow(master_table):", nrow(master_table), "\n")
cat("nrow(emg_table):   ", nrow(emg_table), "\n")
cat("nrow(torque_table):", nrow(torque_table), "\n")
cat("nrow(analysis_table):", nrow(analysis_table), "\n")

# check: torque should contain NA values after joining to the denser EMG grid (normal)
analysis_table %>%
  summarise(
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est_hz = n() / duration_s,
    torque_na_frac = mean(is.na(torque_raw))
  )
```

```{r check_emg_sequence_validity}
emg_validity <- read_parquet(
  emg_file_path,
  col_select = c("SEQ", "sequence_valid_emg6", "sequence_valid_emg8", "sequence_valid_emg10")
) |>
  distinct(SEQ, sequence_valid_emg6, sequence_valid_emg8, sequence_valid_emg10)

print(emg_validity)

critical_seqs     <- c("MVC_REF", "SVC_REF", "EX_DYN", "EX_STA", "MVC_RECOV_DYN", "MVC_RECOV_STA")
emg_validity_crit <- emg_validity |> dplyr::filter(SEQ %in% critical_seqs)

# VL is the required channel for cohort plots; RF/VM failures can continue as partial runs.
invalid_vl_seqs <- emg_validity_crit |>
  dplyr::filter(sequence_valid_emg10 == 0)

invalid_other_seqs <- emg_validity_crit |>
  dplyr::filter(sequence_valid_emg10 != 0) |>
  dplyr::filter(dplyr::if_any(c(sequence_valid_emg6, sequence_valid_emg8), ~ . == 0))

if (nrow(invalid_vl_seqs) > 0) {
  stop(glue(
    "EMG sequence validity check FAILED for {RUN_ID} (VL invalid):\n",
    paste(capture.output(print(invalid_vl_seqs)), collapse = "\n")
  ))
}

if (nrow(invalid_other_seqs) > 0) {
  message(glue(
    "WARNING: bad non-VL channel(s) in {RUN_ID} — continuing (VL OK):\n",
    paste(capture.output(print(invalid_other_seqs)), collapse = "\n")
  ))
} else {
  cat("EMG sequence validity OK for all sequences.\n")
}
```




```{r mvc_svc_load}
# Keep contraction samples for the reference and recovery sequences.
mvc_ref_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_REF",
              .data[["VC"]] == 1) |>
  dplyr::arrange(.data[["time_index"]])

svc_ref_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "SVC_REF",
              .data[["VC"]] == 1) |>
  dplyr::arrange(.data[["time_index"]])

mvc_recov_dyn_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_RECOV_DYN",
              .data[["VC"]] == 1) |>
  dplyr::arrange(.data[["time_index"]])

mvc_recov_sta_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "MVC_RECOV_STA",
              .data[["VC"]] == 1) |>
  dplyr::arrange(.data[["time_index"]])
```


```{r ex_dyn_ex_sta_load}
# Keep contraction samples for the two fatigue sequences.
ex_dyn_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "EX_DYN",
                .data[["VC"]] == 1) |>
  dplyr::arrange(.data[["time_index"]])

ex_sta_table <- analysis_table |>
  dplyr::filter(.data[["SEQ"]] == "EX_STA",
                .data[["VC"]] == 1) |>
  dplyr::arrange(.data[["time_index"]])
```
## Torque calibration
```{r torque_calibration}
# Biodex calibration constants.
torque_offset_volt <- 0.106
torque_gain_Nm_per_volt <- 145.30

calc_torque_nm <- function(df) {
  return(df %>% dplyr::mutate(torque_Nm = torque_gain_Nm_per_volt * (torque_raw - torque_offset_volt)))
}

ex_dyn_table       <- calc_torque_nm(ex_dyn_table)
ex_sta_table       <- calc_torque_nm(ex_sta_table)
mvc_ref_table      <- calc_torque_nm(mvc_ref_table)
svc_ref_table      <- calc_torque_nm(svc_ref_table)
mvc_recov_dyn_table <- calc_torque_nm(mvc_recov_dyn_table)
mvc_recov_sta_table <- calc_torque_nm(mvc_recov_sta_table)
```

```{r cleanup_analysis_table}
# Drop the full joined table after creating the sequence-specific subsets.
rm(analysis_table)
rm(emg_table)
rm(torque_table)
gc()
```

## Signal processing

```{r EMG_filtering}
# 20-450 Hz with zero-phase filtering 
emg_sampling_frequency_hz <- 2148.149

emg_bandpass_low_hz <- 20
emg_bandpass_high_hz <- 450
emg_filter_order <- 2

emg_bandpass_filter <- function(emg_signal, sampling_frequency_hz, 
                                low_hz, high_hz, filter_order) {
  
  # Remove DC offset before filtering (not essential with wireless + hardware filtering.)
  emg_centered <- emg_signal - mean(emg_signal, na.rm = TRUE)
  
  # Express cutoffs relative to the Nyquist frequency.
  normalized_cutoffs <- c(low_hz, high_hz) / (sampling_frequency_hz / 2)
  
  # Build the Butterworth band-pass filter.
  butter_filter <- signal::butter(filter_order, normalized_cutoffs, type = "pass")
  
  # Zero-phase filtering avoids shifting the signal in time.
  return(signal::filtfilt(butter_filter, emg_centered))
}
```

```{r apply_emg_filter}
emg_channels <- c("emg6", "emg8", "emg10")

apply_emg_filter <- function(df) {
  return(df |>
    dplyr::mutate(
      dplyr::across(
        dplyr::all_of(emg_channels),
        ~ emg_bandpass_filter(.x, emg_sampling_frequency_hz,
                              emg_bandpass_low_hz, emg_bandpass_high_hz, emg_filter_order)
      )
    ))
}

ex_dyn_table        <- apply_emg_filter(ex_dyn_table)
ex_sta_table        <- apply_emg_filter(ex_sta_table)
mvc_ref_table       <- apply_emg_filter(mvc_ref_table)
svc_ref_table       <- apply_emg_filter(svc_ref_table)
mvc_recov_dyn_table <- apply_emg_filter(mvc_recov_dyn_table)
mvc_recov_sta_table <- apply_emg_filter(mvc_recov_sta_table)
```

## Artifact handling

Global extreme-percentile repair targets rare hardware-like spikes without anchoring thresholds to the heavy-tailed 100% MVC signal.

```{r amplitude_spike_handling}
# Artifact handling parameters.
# spike_pct:            global extreme-percentile threshold (top 0.01% of |signal|).
# spike_expand_samples: after flagging, expand each cluster by ±N samples to remove
#                       the spike "pedestal". Validated visually across all 30 runs.
# artifact_floor_v:     stop iterating when the threshold would drop below this value.
#                       Prevents the percentile from digging into "real" signal !
#                       Only applied from pass 2 onward — pass 1 always runs.
# spike_max_passes:     default cap on iterations. Most runs stop at pass 1 (floor).
#                       A few runs with extreme spikes benefit from a second pass.
spike_pct            <- 0.9999
spike_expand_samples <- 50L
artifact_floor_v     <- 0.001   # 1 mV

# Per-run pass override. some subject had harsher spikes.
# to hide secondary artifacts; 2 passes needed. All others stop at 1 via the floor.
spike_passes_by_run <- c(
  "151FuCy_20251105" = 2L
)
spike_max_passes <- if (RUN_ID %in% names(spike_passes_by_run))
  spike_passes_by_run[[RUN_ID]] else 5L

apply_amplitude_spike_handling <- function(df, pct = spike_pct,
                                           expand  = spike_expand_samples,
                                           floor_v = artifact_floor_v,
                                           max_passes = spike_max_passes) {
  for (ch in emg_channels) {
    spike_col  <- paste0(ch, "_spike")
    x          <- df[[ch]]
    x_clean    <- x
    total_mask <- rep(FALSE, length(x))

    for (pass in seq_len(max_passes)) {
      threshold  <- quantile(abs(x_clean), pct, na.rm = TRUE)

      # Floor only blocks passes 2+; pass 1 always runs to match previous behaviour.
      if (pass > 1L && threshold < floor_v) break

      spike_mask <- !is.na(x_clean) & abs(x_clean) > threshold

      if (expand > 0L && any(spike_mask)) {
        expanded <- spike_mask
        for (i in which(spike_mask)) {
          i0 <- max(1L, i - expand); i1 <- min(length(spike_mask), i + expand)
          expanded[i0:i1] <- TRUE
        }
        spike_mask <- expanded
      }

      new_flags <- spike_mask & !total_mask
      if (!any(new_flags)) break

      signal_out <- x_clean
      valid_idx  <- which(!spike_mask)
      if (length(valid_idx) > 1) {
        signal_out[spike_mask] <- approx(
          x    = valid_idx,
          y    = x_clean[valid_idx],
          xout = which(spike_mask),
          rule = 2
        )$y
      } else {
        signal_out[spike_mask] <- 0
      }

      total_mask <- total_mask | spike_mask
      x_clean    <- signal_out

      message(sprintf("  %-6s pass %d | threshold = %+.4f V | %d new flagged | %d total (%.4f%%)",
                      ch, pass, threshold, sum(new_flags),
                      sum(total_mask), 100 * mean(total_mask)))
    }

    df[[ch]]        <- x_clean
    df[[spike_col]] <- total_mask
  }
  return(df)
}
```

```{r apply_artifact_handling}
# Store pre-repair copies for QC, then apply extreme-point artifact handling if enabled.
if (APPLY_ARTIFACT_HANDLING) {
  # Tuning preview: cap each table to the first N seconds for fast iteration.
  # Has no effect when TUNING_PREVIEW_SECONDS is NULL or in full_run mode.
  if (PIPELINE_MODE == "tuning" && !is.null(TUNING_PREVIEW_SECONDS)) {
    preview_rows <- round(TUNING_PREVIEW_SECONDS * 2148)
    ex_dyn_table        <- head(ex_dyn_table,        preview_rows)
    ex_sta_table        <- head(ex_sta_table,        preview_rows)
    mvc_ref_table       <- head(mvc_ref_table,       preview_rows)
    svc_ref_table       <- head(svc_ref_table,       preview_rows)
    mvc_recov_dyn_table <- head(mvc_recov_dyn_table, preview_rows)
    mvc_recov_sta_table <- head(mvc_recov_sta_table, preview_rows)
    message(sprintf("Tuning preview: using first %ds of each table (%d rows max)",
                    TUNING_PREVIEW_SECONDS, preview_rows))
  }

  ex_dyn_table_pre        <- ex_dyn_table
  ex_sta_table_pre        <- ex_sta_table
  mvc_ref_table_pre       <- mvc_ref_table
  svc_ref_table_pre       <- svc_ref_table
  mvc_recov_dyn_table_pre <- mvc_recov_dyn_table
  mvc_recov_sta_table_pre <- mvc_recov_sta_table

  message("EX_DYN");        ex_dyn_table        <- apply_amplitude_spike_handling(ex_dyn_table)
  message("EX_STA");        ex_sta_table        <- apply_amplitude_spike_handling(ex_sta_table)
  message("MVC_REF");       mvc_ref_table       <- apply_amplitude_spike_handling(mvc_ref_table)
  message("SVC_REF");       svc_ref_table       <- apply_amplitude_spike_handling(svc_ref_table)
  message("MVC_RECOV_DYN"); mvc_recov_dyn_table <- apply_amplitude_spike_handling(mvc_recov_dyn_table)
  message("MVC_RECOV_STA"); mvc_recov_sta_table <- apply_amplitude_spike_handling(mvc_recov_sta_table)

  # Printed for QC against the artifact PDF; not used downstream.
  spike_summary <- purrr::map_dfr(
    list(ex_dyn = ex_dyn_table, ex_sta = ex_sta_table,
         mvc_ref = mvc_ref_table, svc_ref = svc_ref_table,
         mvc_recov_dyn = mvc_recov_dyn_table, mvc_recov_sta = mvc_recov_sta_table),
    ~ dplyr::summarise(.x,
        dplyr::across(dplyr::ends_with("_spike"), \(x) sum(x, na.rm = TRUE))),
    .id = "table"
  )
  print(spike_summary)
}
```

```{r artifact_qc_plots, eval=APPLY_ARTIFACT_HANDLING}
# Before/after artifact QC: one overview plot per channel for EX_DYN, EX_STA, MVC_REF.
# Red dots show the original amplitude of flagged samples before interpolation.
channel_label_map <- c(emg6 = "RF", emg8 = "VM", emg10 = "VL")

plot_artifact_qc <- function(pre_df, post_df, table_label, max_rows = ARTIFACT_QC_MAX_ROWS) {
  return(purrr::map(emg_channels, function(ch) {
    spike_col <- paste0(ch, "_spike")
    muscle    <- channel_label_map[[ch]]
    n_spikes  <- sum(post_df[[spike_col]], na.rm = TRUE)

    if (is.null(max_rows)) {
      pre_sub  <- pre_df
      post_sub <- post_df
    } else {
      pre_sub  <- head(pre_df,  max_rows)
      post_sub <- head(post_df, max_rows)
    }

    n_rows_shown <- nrow(pre_sub)

    base_df <- data.frame(
      time_ref_s = pre_sub$time_ref_s,
      emg_val    = pre_sub[[ch]]
    )

    spike_idx <- which(post_sub[[spike_col]])
    has_spikes <- length(spike_idx) > 0

    if (has_spikes) {
      spikes_df <- data.frame(
        time_ref_s   = pre_sub$time_ref_s[spike_idx],
        original_val = pre_sub[[ch]][spike_idx],
        repaired_val = post_sub[[ch]][spike_idx]
      )
    }

    p <- ggplot2::ggplot(base_df, ggplot2::aes(x = time_ref_s, y = emg_val)) +
      ggplot2::geom_line(colour = "grey50", linewidth = 0.35, alpha = 0.9)

    if (has_spikes) {
      p <- p +
        # vertical segment: spike tip (red) down to interpolated value (blue)
        ggplot2::geom_segment(
          data = spikes_df,
          ggplot2::aes(x = time_ref_s, xend = time_ref_s,
                       y = original_val, yend = repaired_val),
          colour = "red", linewidth = 1, inherit.aes = FALSE
        ) +
        # original spike amplitude — red filled circle
        ggplot2::geom_point(
          data = spikes_df,
          ggplot2::aes(x = time_ref_s, y = original_val),
          colour = "red", size = 3, shape = 19, inherit.aes = FALSE
        ) +
        # interpolated replacement — blue filled circle
        ggplot2::geom_point(
          data = spikes_df,
          ggplot2::aes(x = time_ref_s, y = repaired_val),
          colour = "steelblue", size = 3, shape = 19, inherit.aes = FALSE
        )
    }

    subtitle <- if (!has_spikes) {
      if (is.null(max_rows)) {
        "No spikes in full table"
      } else {
        paste0("No spikes in first ", n_rows_shown, " rows")
      }
    } else {
      paste0(n_spikes, " spike(s) flagged in full table",
             " | red = original  •  blue = interpolated replacement",
             if (is.null(max_rows)) {
               " | showing full table"
             } else {
               paste0(" | showing first ", n_rows_shown, " rows")
             })
    }

    p <- p +
      ggplot2::labs(
        title    = paste0("Artifact QC — ", table_label, " — ", muscle,
                          "  [spike_pct = ", spike_pct, "]"),
        subtitle = subtitle,
        x = "Time (s)", y = "EMG (V)"
      ) +
      ggplot2::theme_minimal(base_size = 10)

    p
  }))
}

artifact_qc_plots <- list(
  EX_DYN        = plot_artifact_qc(ex_dyn_table_pre,        ex_dyn_table,        "EX_DYN"),
  EX_STA        = plot_artifact_qc(ex_sta_table_pre,        ex_sta_table,        "EX_STA"),
  MVC_REF       = plot_artifact_qc(mvc_ref_table_pre,       mvc_ref_table,       "MVC_REF"),
  SVC_REF       = plot_artifact_qc(svc_ref_table_pre,       svc_ref_table,       "SVC_REF"),
  MVC_RECOV_DYN = plot_artifact_qc(mvc_recov_dyn_table_pre, mvc_recov_dyn_table, "MVC_RECOV_DYN"),
  MVC_RECOV_STA = plot_artifact_qc(mvc_recov_sta_table_pre, mvc_recov_sta_table, "MVC_RECOV_STA")
)

if (SHOW_TUNING_QC_PLOTS) {
  purrr::walk(artifact_qc_plots, ~ purrr::walk(.x, print))
}

artifact_qc_pdf_path <- here::here("results", "QC_EXPORT", "emg_individual_analysis",
  paste0(RUN_ID, "_artifact_qc.pdf")
)
dir.create(dirname(artifact_qc_pdf_path), recursive = TRUE, showWarnings = FALSE)
grDevices::pdf(file = artifact_qc_pdf_path, width = 14, height = 4, onefile = TRUE)
purrr::walk(artifact_qc_plots, ~ purrr::walk(.x, print))
grDevices::dev.off()
message("[ARTIFACT QC] PDF -> ", normalizePath(artifact_qc_pdf_path, winslash = "/", mustWork = FALSE))
```


# Analysis windows

Use 1 s windows throughout: middle 1 s for repeated contractions, non-overlapping 1 s windows for `EX_STA`.

## Middle 1 s window extraction

```{r extract_middle_1s_window_function}
extract_middle_1s_window <- function(rep_df, window_length_s = 1.0) {

  return(rep_df |>
    dplyr::group_by(.data$VC_count) |>
    dplyr::mutate(
      t_mid = (min(.data$time_ref_s) + max(.data$time_ref_s)) / 2,
      mid_start = .data$t_mid - window_length_s / 2,
      mid_end   = .data$t_mid + window_length_s / 2
    ) |>
    dplyr::filter(
      .data$time_ref_s >= mid_start,
      .data$time_ref_s < mid_end
    ) |>
    dplyr::select(-.data$t_mid, -.data$mid_start, -.data$mid_end) |>
    dplyr::ungroup())
}
```

```{r build_1s_table}
# These summaries catch off-by-one windowing errors after protocol/path changes.
mvc_ref_mid1s_table <- extract_middle_1s_window(mvc_ref_table)

mvc_ref_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

svc_ref_mid1s_table <- extract_middle_1s_window(svc_ref_table)

svc_ref_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

mvc_recov_dyn_mid1s_table <- extract_middle_1s_window(mvc_recov_dyn_table)

mvc_recov_dyn_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

mvc_recov_sta_mid1s_table <- extract_middle_1s_window(mvc_recov_sta_table)

mvc_recov_sta_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )

ex_dyn_mid1s_table <- extract_middle_1s_window(ex_dyn_table)

ex_dyn_mid1s_table |>
  dplyr::group_by(VC_count) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    fs_est = n_rows / duration_s,
    .groups = "drop"
  )
```


```{r extract_ex_sta_middle_1s_window}
# EX_STA uses absolute 1 s bins so BIA/NIRS/EMG can share the same time base.

window_length_s <- 1.0
window_step_s   <- 1.0

t_start <- min(ex_sta_table$time_ref_s)
t_end   <- max(ex_sta_table$time_ref_s)

window_start_s <- seq(t_start, t_end - window_length_s, by = window_step_s)

ex_sta_window_table <- tibble::tibble(
  window_id      = seq_along(window_start_s),
  window_start_s = window_start_s,
  window_end_s   = window_start_s + window_length_s,
  window_mid_s   = window_start_s + window_length_s / 2
)

nrow(ex_sta_window_table)
ex_sta_window_table |> dplyr::slice(1:5)

# Attach each EX_STA sample to its window.

ex_sta_mid1s_table <- ex_sta_table |>
  dplyr::mutate(
    window_id = floor((time_ref_s - t_start) / window_step_s) + 1L
  ) |>
  dplyr::filter(
    window_id >= 1L,
    window_id <= nrow(ex_sta_window_table)
  ) |>
  dplyr::left_join(
    ex_sta_window_table,
    by = "window_id"
  ) |>
  dplyr::select(
    time_index, time_ref_s, SEQ_index, SEQ, VC, VC_count,
    window_id, window_start_s, window_end_s, window_mid_s,
    emg6, emg8, emg10, torque_raw, torque_Nm
  )

# Check that each EX_STA window spans about 1 s.
ex_sta_mid1s_table |>
  dplyr::group_by(window_id) |>
  dplyr::summarise(
    n_rows = dplyr::n(),
    duration_s = max(time_ref_s) - min(time_ref_s),
    .groups = "drop"
  ) |>
  dplyr::slice(1:10) |>
  print()
```

# Feature extraction

All EMG features are computed on the 1 s windows defined above.

## EMG amplitude descriptor: RMS

```{r RMS_function}
# RMS amplitude for one window.
emg_rms <- function(emg_signal) {
  return(sqrt(mean(emg_signal^2, na.rm = TRUE)))
}
```


## EMG spectral descriptor: median frequency

```{r MDF_function}
# Median frequency from the single-sided power spectrum.
emg_median_frequency_hz <- function(emg_signal, sampling_frequency_hz) {
  # Re-center inside the feature function so each window is handled independently.
  emg_centered <- emg_signal - mean(emg_signal)
  n_samples <- length(emg_centered)

  # Hann window reduces spectral leakage before the FFT.
  hann_window <- 0.5 - 0.5 * cos(2 * pi * (0:(n_samples - 1)) / (n_samples - 1))
  emg_windowed <- emg_centered * hann_window

  fft_complex <- stats::fft(emg_windowed)
  n_half <- floor(n_samples / 2) + 1
  fft_half <- fft_complex[1:n_half]
  
  power_spectrum <- (Mod(fft_half) ^ 2) / n_samples
  frequency_axis_hz <- (0:(n_half - 1)) * (sampling_frequency_hz / n_samples)
  
  # Drop the 0 Hz bin before computing the median frequency.
  power_spectrum <- power_spectrum[-1]
  frequency_axis_hz <- frequency_axis_hz[-1]

  cumulative_power <- cumsum(power_spectrum)
  median_index <- which(cumulative_power >= 0.5 * sum(power_spectrum))[1]

  return(frequency_axis_hz[median_index])
}
```


# Feature tables

## MVC reference features

```{r compute_feature_mvc_ref_mid1s_ref_table}
# Per-repetition features: one block per contraction (VC_count) -> RMS + median
# frequency for each channel, with that rep's mean torque attached. Torque is
# averaged once per block and repeated across the three channel rows, which keeps
# the long one-row-per-channel shape every downstream table and the cohort scripts
# expect.
channels <- c("emg6", "emg8", "emg10")

mvc_ref_split <- split(mvc_ref_mid1s_table, mvc_ref_mid1s_table$VC_count)

mvc_ref_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_ref_split), function(vc_id) {

  block <- mvc_ref_split[[vc_id]]

  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_REF",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)
```

```{r compute_feature_svc_ref_mid1s_table}
# Same per-rep extraction as MVC_REF above, for the submaximal references (30/50/70%).
svc_ref_split <- split(svc_ref_mid1s_table, svc_ref_mid1s_table$VC_count)

svc_ref_mid1s_feature_table <- dplyr::bind_rows(lapply(names(svc_ref_split), function(vc_id) {

  block <- svc_ref_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "SVC_REF",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)
```

```{r compute_feature_mvc_recov_dyn_mid1s_table}
# Same per-rep extraction, post-fatigue dynamic-recovery MVCs.
mvc_recov_dyn_split <- split(mvc_recov_dyn_mid1s_table, mvc_recov_dyn_mid1s_table$VC_count)

mvc_recov_dyn_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_recov_dyn_split), function(vc_id) {

  block <- mvc_recov_dyn_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_RECOV_DYN",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)
```

```{r compute_feature_mvc_recov_sta_mid1s_table}
# Same per-rep extraction, post-fatigue static-recovery MVCs.
mvc_recov_sta_split <- split(mvc_recov_sta_mid1s_table, mvc_recov_sta_mid1s_table$VC_count)

mvc_recov_sta_mid1s_feature_table <- dplyr::bind_rows(lapply(names(mvc_recov_sta_split), function(vc_id) {

  block <- mvc_recov_sta_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "MVC_RECOV_STA",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, VC_count)
```

```{r combine_feature_mid1s_tables}
# Stack the four reference/recovery feature tables into one long table.
reference_mid1s_feature_table <- dplyr::bind_rows(
  mvc_ref_mid1s_feature_table,
  svc_ref_mid1s_feature_table,
  mvc_recov_dyn_mid1s_feature_table,
  mvc_recov_sta_mid1s_feature_table
) |>
  dplyr::arrange(SEQ, channel, VC_count)
```

```{r mvc_level_summary}
# Collapse the MVC sequences to one mean row per channel. The MVC_REF rows become
# the baseline (torque / RMS / MDF denominators) that every normalization below uses.
mvc_level_summary <- reference_mid1s_feature_table |>
  dplyr::filter(SEQ %in% c("MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")) |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    n_reps = dplyr::n(),
    torque_mean_Nm = mean(torque_mean_Nm, na.rm = TRUE),
    rms_mid_1s_mean = mean(rms_mid_1s, na.rm = TRUE),
    mdf_mid_1s_hz_mean = mean(mdf_mid_1s_hz, na.rm = TRUE),
    .groups = "drop"
  )

mvc_ref_baseline <- mvc_level_summary |>
  dplyr::filter(SEQ == "MVC_REF") |>
  dplyr::select(
    channel,
    torque_ref_Nm = torque_mean_Nm,
    rms_ref = rms_mid_1s_mean,
    mdf_ref_hz = mdf_mid_1s_hz_mean
  )

# Use the MVC reference torque mean as the shared torque baseline.
torque_mvc_ref_baseline <- mean(mvc_ref_baseline$torque_ref_Nm, na.rm = TRUE)
```

## EX_DYN feature table

```{r build_ex_dyn_1mids_feature_table}
# Per-rep features for the dynamic fatigue task (same extraction as the references).
ex_dyn_split <- split(ex_dyn_mid1s_table, ex_dyn_mid1s_table$VC_count)

ex_dyn_mid1s_feature_table <- dplyr::bind_rows(lapply(names(ex_dyn_split), function(vc_id) {

  block <- ex_dyn_split[[vc_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "EX_DYN",
      VC_count = as.integer(vc_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  # Keep the fixed 60-repetition analysis window used by the cohort scripts.
  dplyr::filter(VC_count <= 60L) |>
  dplyr::arrange(channel, VC_count)
```


## EX_STA feature table

```{r compute_feature_ex_sta_mid1s_table}
# EX_STA features per fixed 1 s window (window_id) rather than per contraction 
# the static hold is one continuous effort, not discrete reps.
ex_sta_split <- split(ex_sta_mid1s_table, ex_sta_mid1s_table$window_id)

ex_sta_mid1s_feature_table <- dplyr::bind_rows(lapply(names(ex_sta_split), function(win_id) {

  block <- ex_sta_split[[win_id]]
  torque_mean_Nm <- mean(block$torque_Nm, na.rm = TRUE)

  dplyr::bind_rows(lapply(channels, function(ch) {

    x <- block[[ch]]

    dplyr::tibble(
      SEQ = "EX_STA",
      window_id = as.integer(win_id),
      channel = ch,
      rms_mid_1s = emg_rms(x),
      mdf_mid_1s_hz = emg_median_frequency_hz(x, emg_sampling_frequency_hz),
      torque_mean_Nm = torque_mean_Nm,
      n_samples = sum(is.finite(x))
    )
  }))
})) |>
  dplyr::arrange(channel, window_id)
```

```{r filter_ex_sta_onset_windows}
# Common onset keeps the EX_STA denominator aligned with BIA and NIRS exports.
EX_STA_COMMON_ONSET_WINDOW_ID <- 2L

ex_sta_first_retained_window_id  <- EX_STA_COMMON_ONSET_WINDOW_ID
ex_sta_first_retained_window_mid_s <- ex_sta_window_table |>
  dplyr::filter(window_id == ex_sta_first_retained_window_id) |>
  dplyr::pull(window_mid_s)

ex_sta_onset_filter_log <- tibble::tibble(
  first_retained_window_id    = ex_sta_first_retained_window_id,
  first_retained_window_mid_s = ex_sta_first_retained_window_mid_s
)

ex_sta_mid1s_feature_table <- ex_sta_mid1s_feature_table |>
  dplyr::filter(window_id >= ex_sta_first_retained_window_id)

print(ex_sta_onset_filter_log)
```


# Baseline normalization

## MVC baseline computation

```{r compute_mvc_ref_mid1s_rms_baseline_table}
# Per-channel MVC_REF means -> the RMS and MDF denominators for all normalization below.
mvc_ref_mid1s_rms_baseline_table <- mvc_ref_mid1s_feature_table |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    rms_mvc_ref_mid1s_baseline = mean(rms_mid_1s, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)

mvc_ref_mid1s_mdf_baseline_table <- mvc_ref_mid1s_feature_table |>
  dplyr::group_by(channel) |>
  dplyr::summarise(
    mdf_mvc_ref_mid1s_baseline_hz = mean(mdf_mid_1s_hz, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)
```


## RMS normalization

```{r ex_dyn_mid1s_rms_normalization}
# Normalize EX_DYN to the MVC_REF baseline: nRMS and nMDF as fractions of their MVC mean
# nTorque against the shared torque baseline (so MVC_REF sits at 1.0).
ex_dyn_mid1s_feature_table_norm <- ex_dyn_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::left_join(mvc_ref_mid1s_mdf_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm = rms_mid_1s / rms_mvc_ref_mid1s_baseline,            # nRMS: fraction of the MVC_REF RMS baseline
    mdf_norm = mdf_mid_1s_hz / mdf_mvc_ref_mid1s_baseline_hz,      # nMDF: fraction of the MVC_REF MDF baseline
    torque_norm = torque_mean_Nm / torque_mvc_ref_baseline        # nTorque: fraction of the MVC_REF torque baseline
  )
```


```{r ex_sta_mid1s_rms_normalization}
# Same MVC_REF normalization for EX_STA.
ex_sta_mid1s_feature_table_norm <- ex_sta_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::left_join(mvc_ref_mid1s_mdf_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm = rms_mid_1s / rms_mvc_ref_mid1s_baseline,            # nRMS: fraction of the MVC_REF RMS baseline
    mdf_norm = mdf_mid_1s_hz / mdf_mvc_ref_mid1s_baseline_hz,      # nMDF: fraction of the MVC_REF MDF baseline
    torque_norm = torque_mean_Nm / torque_mvc_ref_baseline        # nTorque: fraction of the MVC_REF torque baseline
  )
```

```{r mvc_recov_mid1s_rms_normalization}
mvc_recov_dyn_mid1s_feature_table_norm <- mvc_recov_dyn_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::left_join(mvc_ref_mid1s_mdf_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm         = rms_mid_1s / rms_mvc_ref_mid1s_baseline,        # nRMS
    mdf_norm         = mdf_mid_1s_hz / mdf_mvc_ref_mid1s_baseline_hz,  # nMDF
    torque_norm      = torque_mean_Nm / torque_mvc_ref_baseline,       # nTorque
    emg_torque_ratio = rms_norm / torque_norm                         # activation per unit torque
  )

mvc_recov_sta_mid1s_feature_table_norm <- mvc_recov_sta_mid1s_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::left_join(mvc_ref_mid1s_mdf_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm         = rms_mid_1s / rms_mvc_ref_mid1s_baseline,        # nRMS
    mdf_norm         = mdf_mid_1s_hz / mdf_mvc_ref_mid1s_baseline_hz,  # nMDF
    torque_norm      = torque_mean_Nm / torque_mvc_ref_baseline,       # nTorque
    emg_torque_ratio = rms_norm / torque_norm                         # activation per unit torque
  )
```

```{r emg_torque_ratio_tables}
# Build the normalized EMG-to-torque ratio.
ex_dyn_emg_torque_ratio_table <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::mutate(
    emg_torque_ratio = rms_norm / torque_norm
  ) |>
  dplyr::select(
    SEQ, VC_count, channel,
    rms_norm, torque_norm, emg_torque_ratio
  )

ex_sta_emg_torque_ratio_table <- ex_sta_mid1s_feature_table_norm |>
  dplyr::left_join(
    ex_sta_window_table |>
      dplyr::select(window_id, window_mid_s),
    by = "window_id"
  ) |>
  dplyr::mutate(
    emg_torque_ratio = rms_norm / torque_norm
  ) |>
  dplyr::select(
    SEQ, window_id, window_mid_s, channel,
    rms_norm, torque_norm, emg_torque_ratio
  )
```

```{r emg_torque_ratio_slope_summary}
# Fit one linear ratio slope per channel.
ex_dyn_emg_torque_ratio_slope_summary_table <- ex_dyn_emg_torque_ratio_table |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    time_index_variable = "VC_count",
    fit_type = "lm(emg_torque_ratio ~ VC_count)",
    slope_estimate = unname(coef(lm(emg_torque_ratio ~ VC_count))[2]),  # ratio change per rep (sign = fatigue direction)
    intercept = unname(coef(lm(emg_torque_ratio ~ VC_count))[1]),
    r_squared = summary(lm(emg_torque_ratio ~ VC_count))$r.squared,
    n_points = dplyr::n(),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)

ex_sta_emg_torque_ratio_slope_summary_table <- ex_sta_emg_torque_ratio_table |>
  dplyr::group_by(SEQ, channel) |>
  dplyr::summarise(
    time_index_variable = "window_mid_s",
    fit_type = "lm(emg_torque_ratio ~ window_mid_s)",
    slope_estimate = unname(coef(lm(emg_torque_ratio ~ window_mid_s))[2]),
    intercept = unname(coef(lm(emg_torque_ratio ~ window_mid_s))[1]),
    r_squared = summary(lm(emg_torque_ratio ~ window_mid_s))$r.squared,
    n_points = dplyr::n(),
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)

emg_torque_ratio_slope_summary_table <- dplyr::bind_rows(
  ex_dyn_emg_torque_ratio_slope_summary_table,
  ex_sta_emg_torque_ratio_slope_summary_table
) |>
  dplyr::select(
    SEQ,
    channel,
    time_index_variable,
    slope_estimate,
    intercept,
    r_squared,
    n_points,
    fit_type
  ) |>
  dplyr::arrange(SEQ, channel)
```


# Reference modeling

## MVC/SVC trends

```{r build_reference_level_feature_table}
# Extend the existing 30/50/70 reference set with a 100% point from MVC_REF.
# The 100% row is an aggregated MVC_REF summary, not a true fourth SVC repetition.
# VC_count = 4L is kept only as a compatibility placeholder; plot/order logic should use reference_level_id.
reference_level_feature_table <- dplyr::bind_rows(
  svc_ref_mid1s_feature_table |>
    dplyr::mutate(
      reference_level_id = VC_count,
      reference_level = factor(
        VC_count,
        levels = c(1, 2, 3, 4),
        labels = c("30%", "50%", "70%", "100%")
      )
    ),
  mvc_ref_mid1s_feature_table |>
    dplyr::group_by(channel) |>
    dplyr::summarise(
      SEQ = "MVC_REF",
      VC_count = 4L,
      rms_mid_1s = mean(rms_mid_1s, na.rm = TRUE),
      mdf_mid_1s_hz = mean(mdf_mid_1s_hz, na.rm = TRUE),
      torque_mean_Nm = mean(torque_mean_Nm, na.rm = TRUE),
      n_samples = mean(n_samples, na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(
      reference_level_id = 4L,
      reference_level = factor(
        reference_level_id,
        levels = c(1, 2, 3, 4),
        labels = c("30%", "50%", "70%", "100%")
      )
    )
) |>
  dplyr::arrange(channel, reference_level_id)
```

```{r build_reference_level_feature_table_norm}
# Keep the normalized reference workflow separate from the validated Nm-space table.
reference_level_feature_table_norm <- reference_level_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::left_join(mvc_ref_mid1s_mdf_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm = rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    mdf_norm = mdf_mid_1s_hz / mdf_mvc_ref_mid1s_baseline_hz,
    torque_norm = torque_mean_Nm / torque_mvc_ref_baseline,
    reference_level = as.character(reference_level)
  ) |>
  dplyr::select(
    channel,
    SEQ,
    VC_count,
    reference_level_id,
    reference_level,
    rms_mid_1s,
    mdf_mid_1s_hz,
    torque_mean_Nm,
    rms_mvc_ref_mid1s_baseline,
    mdf_mvc_ref_mid1s_baseline_hz,
    rms_norm,
    mdf_norm,
    torque_norm,
    n_samples
  )

recov_norm <- reference_mid1s_feature_table |>
  dplyr::filter(.data$SEQ %in% c("MVC_RECOV_DYN", "MVC_RECOV_STA")) |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::left_join(mvc_ref_mid1s_mdf_baseline_table, by = "channel") |>
  dplyr::mutate(
    rms_norm           = rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    mdf_norm           = mdf_mid_1s_hz / mdf_mvc_ref_mid1s_baseline_hz,
    torque_norm        = torque_mean_Nm / torque_mvc_ref_baseline,
    reference_level_id = NA_integer_,
    reference_level    = NA_character_
  ) |>
  dplyr::select(
    channel,
    SEQ,
    VC_count,
    reference_level_id,
    reference_level,
    rms_mid_1s,
    mdf_mid_1s_hz,
    torque_mean_Nm,
    rms_mvc_ref_mid1s_baseline,
    mdf_mvc_ref_mid1s_baseline_hz,
    rms_norm,
    mdf_norm,
    torque_norm,
    n_samples
  )

reference_level_feature_table_norm <- dplyr::bind_rows(
  reference_level_feature_table_norm,
  recov_norm
) |>
  dplyr::arrange(channel, SEQ, VC_count)
```

```{r build_reference_calibration_table_norm}
# Fit the normalized RMS reference model: `rms_norm ~ torque_norm`.
reference_calibration_input_table_norm <- reference_level_feature_table_norm |>
  dplyr::select(
    channel,
    reference_level_id,
    reference_level,
    torque_norm,
    rms_norm
  ) |>
  dplyr::mutate(
    response_variable = "rms_norm",
    fit_type = "linear",
    predictor_variable = "torque_norm",
    predictor_definition = "Mean normalized torque from the middle 1 s reference window (MVC_REF = 1)",
    response_definition = "Normalized RMS from the middle 1 s reference window (MVC_REF = 1)"
  ) |>
  dplyr::filter(
    is.finite(torque_norm),
    is.finite(rms_norm)
  ) |>
  dplyr::arrange(channel, reference_level_id)

reference_calibration_table_norm <- reference_calibration_input_table_norm |>
  dplyr::group_by(
    channel,
    response_variable,
    fit_type,
    predictor_variable,
    predictor_definition,
    response_definition
  ) |>
  dplyr::summarise(
    n_reference_levels = dplyr::n(),                            # points in the fit (30/50/70/100%)
    slope = unname(coef(lm(rms_norm ~ torque_norm))[2]),        # fresh nRMS-vs-nTorque slope
    intercept = unname(coef(lm(rms_norm ~ torque_norm))[1]),
    r_squared = summary(lm(rms_norm ~ torque_norm))$r.squared,  # fit quality
    .groups = "drop"
  ) |>
  dplyr::arrange(channel)
```

```{r augment_ratio_tables_with_reference_expected}
# Add the expected ratio from the pre-fatigue normalized reference fit.
# `expected_ratio = (intercept + slope * torque_norm) / torque_norm`
reference_norm_coefs <- reference_calibration_table_norm |>
  dplyr::select(channel, slope_ref = slope, intercept_ref = intercept)

ex_dyn_emg_torque_ratio_table <- ex_dyn_emg_torque_ratio_table |>
  dplyr::left_join(reference_norm_coefs, by = "channel") |>
  dplyr::mutate(
    expected_ratio = (intercept_ref + slope_ref * torque_norm) / torque_norm
  ) |>
  dplyr::select(-slope_ref, -intercept_ref)

ex_sta_emg_torque_ratio_table <- ex_sta_emg_torque_ratio_table |>
  dplyr::left_join(reference_norm_coefs, by = "channel") |>
  dplyr::mutate(
    expected_ratio = (intercept_ref + slope_ref * torque_norm) / torque_norm
  ) |>
  dplyr::select(-slope_ref, -intercept_ref)
```

```{r percent_emg_vs_percent_force_svc_ref}
# Build a simple normalized %EMG vs %Force view.
svc_pct_df <- reference_level_feature_table |>
  dplyr::left_join(mvc_ref_mid1s_rms_baseline_table, by = "channel") |>
  dplyr::mutate(
    reference_level_id = as.integer(reference_level_id),
    torque_pct = 100 * torque_mean_Nm / torque_mvc_ref_baseline,
    emg_pct    = 100 * rms_mid_1s / rms_mvc_ref_mid1s_baseline,
    intensity  = factor(reference_level, levels = c("30%", "50%", "70%", "100%"))
  ) |>
  dplyr::filter(is.finite(reference_level_id))

# Plot normalized %EMG against %Force.
percent_emg_vs_percent_force_plot <- ggplot(svc_pct_df, aes(x = torque_pct, y = emg_pct)) +
  geom_point(aes(shape = intensity), size = 2) +
  geom_line(aes(group = 1)) +
  facet_wrap(~ channel, scales = "free_y", labeller = ggplot2::as_labeller(channel_display_labels)) +
  labs(
    title = "Reference set - %EMG vs %Force (normalized to MVC_REF)",
    x = "% Force (Torque / MVC_REF torque)",
    y = "% EMG (RMS / MVC_REF RMS)"
  ) +
  theme_amstrom_hgrid
```

```{r build_ex_dyn_reference_rms_comparison_table_norm}
# Project the fixed normalized reference fit onto EX_DYN without refitting.
ex_dyn_reference_rms_fit_table_norm <- reference_calibration_table_norm |>
  dplyr::select(
    channel,
    response_variable,
    predictor_variable,
    fit_type,
    slope,
    intercept,
    r_squared
  )

ex_dyn_reference_rms_comparison_table_norm <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::select(
    SEQ,
    VC_count,
    channel,
    torque_norm,
    rms_norm
  ) |>
  dplyr::left_join(
    ex_dyn_reference_rms_fit_table_norm,
    by = "channel"
  ) |>
  dplyr::mutate(
    observed_rms_norm = rms_norm,
    expected_rms_norm = intercept + slope * torque_norm,
    residual_rms_norm = observed_rms_norm - expected_rms_norm,
    relative_residual_rms_norm = dplyr::if_else(
      is.finite(expected_rms_norm) & expected_rms_norm != 0,
      residual_rms_norm / expected_rms_norm,
      NA_real_
    )
  ) |>
  dplyr::select(
    SEQ,
    VC_count,
    channel,
    torque_norm,
    response_variable,
    predictor_variable,
    fit_type,
    observed_rms_norm,
    expected_rms_norm,
    residual_rms_norm,
    relative_residual_rms_norm,
    r_squared
  ) |>
  dplyr::arrange(channel, VC_count)
```

```{r plot_ex_dyn_reference_rms_overlay_norm}
reference_rms_points_plot_df_norm <- reference_level_feature_table_norm |>
  dplyr::select(channel, reference_level, torque_norm, rms_norm)

ref_rms_ci_df <- reference_level_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit    <- lm(rms_norm ~ torque_norm, data = .x)
    x_grid <- tibble::tibble(torque_norm = seq(min(.x$torque_norm), max(.x$torque_norm), length.out = 100))
    pred   <- predict(fit, newdata = x_grid, interval = "confidence", level = 0.95)
    dplyr::bind_cols(x_grid, tibble::as_tibble(pred))
  }) |>
  dplyr::ungroup()

ex_dyn_reference_rms_overlay_plot_norm <- ggplot() +
  geom_ribbon(
    data = ref_rms_ci_df,
    aes(x = torque_norm, ymin = lwr, ymax = upr),
    fill = col_ref_fill, alpha = 0.12
  ) +
  geom_line(
    data = ref_rms_ci_df,
    aes(x = torque_norm, y = fit),
    colour = col_ref_line, linewidth = 1.0, alpha = 0.9
  ) +
  geom_point(
    data = reference_rms_points_plot_df_norm,
    aes(x = torque_norm, y = rms_norm),
    colour = col_ref_pts,
    size = 2.1,
    alpha = 0.9
  ) +
  geom_point(
    data = ex_dyn_reference_rms_comparison_table_norm,
    aes(x = torque_norm, y = observed_rms_norm, colour = VC_count),
    size = 1.7,
    alpha = 0.8
  ) +
  geom_point(
    data = mvc_recov_dyn_mid1s_feature_table_norm,
    aes(x = torque_norm, y = rms_norm),
    shape = 18, size = 4.8, colour = "#222222", inherit.aes = FALSE
  ) +
  geom_point(
    data = mvc_recov_dyn_mid1s_feature_table_norm,
    aes(x = torque_norm, y = rms_norm),
    shape = 18, size = 3.0, colour = "white", inherit.aes = FALSE
  ) +
  facet_wrap(~ channel, scales = "free_y", labeller = ggplot2::as_labeller(channel_display_labels)) +
  scale_colour_viridis_c(option = "plasma", direction = 1, name = "Repetition") +
  labs(
    title = "EX_DYN normalized RMS vs normalized ref fit",
    subtitle = "Grey: calibration fit \u00b1 95% CI.  Colour: rep order.  \u25c6 = MVC_RECOV_DYN.",
    x = "nTorque (MVC_REF = 1)",
    y = "nRMS (MVC_REF = 1)"
  ) +
  theme_amstrom_hgrid +
  theme(legend.position = "right")
```

## EX_DYN trends

```{r ex_dyn_torque_vs_rep}
ex_dyn_torque_vs_rep_plot <- ggplot(ex_dyn_mid1s_feature_table_norm,
       aes(x = VC_count, y = torque_norm)) +
  geom_point(colour = col_neutral_data, size = 1.2, alpha = 0.8) +
  geom_line(colour = col_neutral_data, linewidth = 0.6) +
  facet_wrap(~ channel, labeller = ggplot2::as_labeller(channel_display_labels)) +
  labs(
    title = "EX_DYN - nTorque across repetitions",
    x = "Repetition",
    y = "nTorque"
  ) +
  theme_amstrom_hgrid
```


```{r ex_dyn_rms_vs_torque}
# Reuse the pre-fatigue normalized RMS reference fit and CI ribbon from earlier.
ex_dyn_rms_vs_torque_plot <- ggplot() +
  geom_ribbon(
    data = ref_rms_ci_df,
    aes(x = torque_norm, ymin = lwr, ymax = upr),
    fill = col_ref_fill, alpha = 0.12
  ) +
  geom_line(
    data = ref_rms_ci_df,
    aes(x = torque_norm, y = fit),
    colour = col_ref_line, linewidth = 1.0
  ) +
  geom_point(
    data = ex_dyn_mid1s_feature_table_norm,
    aes(x = torque_norm, y = rms_norm, colour = VC_count),
    alpha = 0.7, size = 1.4
  ) +
  scale_colour_viridis_c(option = "plasma", direction = 1, name = "Repetition") +
  facet_wrap(~ channel, labeller = ggplot2::as_labeller(channel_display_labels)) +
  labs(
    title = "EX_DYN - EMG activation vs nTorque",
    subtitle = "Grey line + ribbon = reference calibration fit (95% CI); colour = repetition order",
    x = "nTorque (MVC_REF = 1)",
    y = "nRMS"
  ) +
  theme_amstrom_hgrid
```  


```{r ex_dyn_mdf_vs_rep}
ex_dyn_mdf_rep_ci_df <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit    <- lm(mdf_mid_1s_hz ~ VC_count, data = .x)
    x_grid <- tibble::tibble(VC_count = seq(min(.x$VC_count), max(.x$VC_count)))
    pred   <- predict(fit, newdata = x_grid, interval = "confidence", level = 0.95)
    dplyr::bind_cols(x_grid, tibble::as_tibble(pred))
  }) |>
  dplyr::ungroup()

ex_dyn_mdf_vs_rep_plot <- ggplot() +
  geom_ribbon(
    data = ex_dyn_mdf_rep_ci_df,
    aes(x = VC_count, ymin = lwr, ymax = upr),
    fill = col_mdf, alpha = 0.15
  ) +
  geom_line(
    data = ex_dyn_mdf_rep_ci_df,
    aes(x = VC_count, y = fit),
    colour = col_mdf, linewidth = 1.0
  ) +
  geom_point(
    data = ex_dyn_mid1s_feature_table_norm,
    aes(x = VC_count, y = mdf_mid_1s_hz),
    colour = col_mdf, alpha = 0.7, size = 1.4
  ) +
  facet_wrap(~ channel, labeller = ggplot2::as_labeller(channel_display_labels)) +
  labs(
    title = "EX_DYN - Median frequency across repetitions",
    subtitle = "Line + ribbon = linear fit on observed data (95% CI)",
    x = "Repetition",
    y = "Median frequency (Hz)"
  ) +
  theme_amstrom_hgrid
```

```{r ex_dyn_mdf_vs_torque}
ex_dyn_mdf_torque_ci_df <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit    <- lm(mdf_mid_1s_hz ~ torque_norm, data = .x)
    x_grid <- tibble::tibble(torque_norm = seq(min(.x$torque_norm), max(.x$torque_norm), length.out = 100))
    pred   <- predict(fit, newdata = x_grid, interval = "confidence", level = 0.95)
    dplyr::bind_cols(x_grid, tibble::as_tibble(pred))
  }) |>
  dplyr::ungroup()

ex_dyn_mdf_vs_torque_plot <- ggplot() +
  geom_ribbon(
    data = ex_dyn_mdf_torque_ci_df,
    aes(x = torque_norm, ymin = lwr, ymax = upr),
    fill = col_mdf, alpha = 0.15
  ) +
  geom_line(
    data = ex_dyn_mdf_torque_ci_df,
    aes(x = torque_norm, y = fit),
    colour = col_mdf, linewidth = 1.0
  ) +
  geom_point(
    data = ex_dyn_mid1s_feature_table_norm,
    aes(x = torque_norm, y = mdf_mid_1s_hz),
    colour = col_mdf, alpha = 0.7, size = 1.4
  ) +
  facet_wrap(~ channel, labeller = ggplot2::as_labeller(channel_display_labels)) +
  labs(
    title = "EX_DYN - Median frequency vs normalized torque",
    subtitle = "Line + ribbon = linear fit on observed data (95% CI)",
    x = "nTorque (MVC_REF = 1)",
    y = "Median frequency (Hz)"
  ) +
  theme_amstrom_hgrid
```  

```{r ex_dyn_rms_vs_repetition}
ex_dyn_rms_rep_ci_df <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::group_by(channel) |>
  dplyr::group_modify(~ {
    fit    <- lm(rms_norm ~ VC_count, data = .x)
    x_grid <- tibble::tibble(VC_count = seq(min(.x$VC_count), max(.x$VC_count)))
    pred   <- predict(fit, newdata = x_grid, interval = "confidence", level = 0.95)
    dplyr::bind_cols(x_grid, tibble::as_tibble(pred))
  }) |>
  dplyr::ungroup()

ex_dyn_rms_vs_repetition_plot <- ggplot() +
  geom_ribbon(
    data = ex_dyn_rms_rep_ci_df,
    aes(x = VC_count, ymin = lwr, ymax = upr),
    fill = col_rms, alpha = 0.15
  ) +
  geom_line(
    data = ex_dyn_rms_rep_ci_df,
    aes(x = VC_count, y = fit),
    colour = col_rms, linewidth = 1.0
  ) +
  geom_point(
    data = ex_dyn_mid1s_feature_table_norm,
    aes(x = VC_count, y = rms_norm),
    colour = col_rms, alpha = 0.7, size = 1.4
  ) +
  facet_wrap(~ channel, labeller = ggplot2::as_labeller(channel_display_labels)) +
  labs(
    title = "EX_DYN - nRMS across repetitions",
    subtitle = "Line + ribbon = linear fit on observed data (95% CI)",
    x = "Repetition",
    y = "nRMS"
  ) +
  theme_amstrom_hgrid
```

```{r ex_dyn_emg_torque_ratio_vs_rep}
ex_dyn_emg_torque_ratio_trend_plot <- ggplot(
  ex_dyn_emg_torque_ratio_table,
  aes(x = VC_count, y = emg_torque_ratio)
) +
  geom_point(alpha = 0.7, size = 1.2, colour = col_rms) +
  geom_line(linewidth = 0.7, colour = col_rms) +
  facet_wrap(~ channel, scales = "free_y", labeller = ggplot2::as_labeller(channel_display_labels)) +
  labs(
    title = "EX_DYN - EMG/torque ratio over repetitions",
    x = "Repetition",
    y = "EMG/torque ratio (nRMS / nTorque)"
  ) +
  theme_amstrom_hgrid
```

```{r ex_dyn_combined_fatigue_plot}
ex_dyn_plot_long <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::select(channel, VC_count, torque_norm, rms_norm, mdf_norm) |>
  tidyr::pivot_longer(
    cols = c(torque_norm, rms_norm, mdf_norm),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_norm = "nTorque",
      rms_norm = "nRMS (a.u.)",
      mdf_norm = "nMDF"
    ),
    metric = factor(metric, levels = c("nTorque", "nRMS (a.u.)", "nMDF"))
  )

ex_dyn_combined_fatigue_plot <- ggplot(ex_dyn_plot_long, aes(x = VC_count, y = value)) +
  geom_point(aes(colour = metric), alpha = 0.65, size = 1.0) +
  scale_colour_manual(values = col_by_metric, guide = "none") +
  facet_grid(metric ~ channel, scales = "free_y", labeller = ggplot2::labeller(channel = ggplot2::as_labeller(channel_display_labels))) +
  labs(
    title = "EX_DYN - Fatigue progression (middle 1s): nTorque, nRMS, nMDF",
    x = "Repetition",
    y = NULL
  ) +
  theme_amstrom_hgrid +
  theme(strip.text.y = element_text(size = rel(0.78)))
```

## EX_STA trends

```{r ex_sta_emg_torque_ratio_vs_time}
ex_sta_emg_torque_ratio_trend_plot <- ggplot(
  ex_sta_emg_torque_ratio_table,
  aes(x = window_mid_s, y = emg_torque_ratio)
) +
  geom_point(alpha = 0.7, size = 1.2, colour = col_rms) +
  geom_line(linewidth = 0.7, colour = col_rms) +
  facet_wrap(~ channel, scales = "free_y", labeller = ggplot2::as_labeller(channel_display_labels)) +
  labs(
    title = "EX_STA - EMG/torque ratio over time",
    x = "Window midpoint time (s)",
    y = "EMG/torque ratio (nRMS / nTorque)"
  ) +
  theme_amstrom_hgrid
```

```{r ex_sta_triple_panel_fatigue}
ex_sta_plot_long <- ex_sta_mid1s_feature_table_norm |>
  dplyr::left_join(
    ex_sta_window_table |>
      dplyr::select(window_id, window_mid_s),
    by = "window_id"
  ) |>
  dplyr::select(channel, window_id, window_mid_s, torque_norm, rms_norm, mdf_norm) |>
  tidyr::pivot_longer(
    cols = c(torque_norm, rms_norm, mdf_norm),
    names_to = "metric",
    values_to = "value"
  ) |>
  dplyr::mutate(
    metric = dplyr::recode(
      metric,
      torque_norm = "nTorque",
      rms_norm = "nRMS (a.u.)",
      mdf_norm = "nMDF"
    ),
    metric = factor(metric, levels = c("nTorque", "nRMS (a.u.)", "nMDF"))
  )

ex_sta_triple_panel_fatigue <- ggplot(ex_sta_plot_long, aes(x = window_mid_s, y = value)) +
  geom_point(aes(colour = metric), alpha = 0.65, size = 1.0) +
  scale_colour_manual(values = col_by_metric, guide = "none") +
  facet_grid(metric ~ channel, scales = "free_y", labeller = ggplot2::labeller(channel = ggplot2::as_labeller(channel_display_labels))) +
  labs(
    title = "EX_STA - Fatigue progression (1 s windows): nTorque, nRMS, nMDF",
    x     = "Time (s)",
    y     = NULL
  ) +
  theme_amstrom_hgrid +
  theme(strip.text.y = element_text(size = rel(0.78)))
```


# Results

```{r emg_plot_gallery, fig.width=11, fig.height=7}
# Print the main single-run QC plots.
if (isTRUE(SHOW_MAIN_PLOT_GALLERY)) {
  print(percent_emg_vs_percent_force_plot)
  print(ex_dyn_torque_vs_rep_plot)
  print(ex_dyn_rms_vs_torque_plot)
  print(ex_dyn_mdf_vs_rep_plot)
  print(ex_dyn_mdf_vs_torque_plot)
  print(ex_dyn_rms_vs_repetition_plot)
  print(ex_dyn_reference_rms_overlay_plot_norm)
  print(ex_dyn_emg_torque_ratio_trend_plot)
  print(ex_dyn_combined_fatigue_plot)
  print(ex_sta_emg_torque_ratio_trend_plot)
  print(ex_sta_triple_panel_fatigue)
}
```

```{r export_emg_plots, message=FALSE}
# Optionally export the selected QC plots to one PDF.
# Use the flags below to choose which plots to save.
EXPORT_RMS_OVERLAY_NORM <- TRUE
EXPORT_DYN_COMBINED     <- TRUE
EXPORT_DYN_RATIO        <- TRUE
EXPORT_STA_COMBINED     <- TRUE
EXPORT_STA_RATIO        <- TRUE

plot_export_flags <- c(
  rms_overlay_norm = EXPORT_RMS_OVERLAY_NORM,
  dyn_combined     = EXPORT_DYN_COMBINED,
  dyn_ratio        = EXPORT_DYN_RATIO,
  sta_combined     = EXPORT_STA_COMBINED,
  sta_ratio        = EXPORT_STA_RATIO
)

plot_export_map <- list(
  rms_overlay_norm = ex_dyn_reference_rms_overlay_plot_norm,
  dyn_combined     = ex_dyn_combined_fatigue_plot,
  dyn_ratio        = ex_dyn_emg_torque_ratio_trend_plot,
  sta_combined     = ex_sta_triple_panel_fatigue,
  sta_ratio        = ex_sta_emg_torque_ratio_trend_plot
)

if (isTRUE(EXPORT_EMG_PLOTS)) {
  selected_plot_names <- names(plot_export_flags)[plot_export_flags]
  if (length(selected_plot_names) == 0L) {
    stop("EXPORT_EMG_PLOTS=TRUE but no plot selector is set to TRUE.")
  }

  qc_export_root <- here::here("results", "QC_EXPORT", "emg_individual_analysis")
  dir.create(qc_export_root, recursive = TRUE, showWarnings = FALSE)

  qc_emg_export_path <- file.path(
    qc_export_root,
    paste0(RUN_ID, "_emg_qc.pdf")
  )

  grDevices::pdf(file = qc_emg_export_path, width = 11, height = 8.5, onefile = TRUE)
  for (plot_name in selected_plot_names) {
    print(plot_export_map[[plot_name]])
  }
  grDevices::dev.off()

  message("\n[QC EMG EXPORT] PDF -> ", normalizePath(qc_emg_export_path, winslash = "/", mustWork = FALSE))
}
```

```{r export_analysis_tables, message=FALSE}
# Optionally export the per-run R-derived tables for cohort analysis.
# Keep them separate from the Python `DATA_EXPORT` parquet files.
EXPORT_RUN_METADATA                 <- TRUE
EXPORT_REFERENCE_LEVEL_FEATURES     <- TRUE
EXPORT_EX_DYN_FEATURES              <- TRUE
EXPORT_EX_STA_FEATURES              <- TRUE
EXPORT_EX_DYN_RATIO_TABLE           <- TRUE
EXPORT_EX_STA_RATIO_TABLE           <- TRUE
EXPORT_RATIO_SLOPE_SUMMARY          <- TRUE
EXPORT_REFERENCE_CALIBRATION_NORM   <- TRUE

analysis_export_flags <- c(
  run_metadata                   = EXPORT_RUN_METADATA,
  reference_level_features_norm  = EXPORT_REFERENCE_LEVEL_FEATURES,
  ex_dyn_features_norm           = EXPORT_EX_DYN_FEATURES,
  ex_sta_features_norm           = EXPORT_EX_STA_FEATURES,
  ex_dyn_ratio_table             = EXPORT_EX_DYN_RATIO_TABLE,
  ex_sta_ratio_table             = EXPORT_EX_STA_RATIO_TABLE,
  ratio_slope_summary            = EXPORT_RATIO_SLOPE_SUMMARY,
  reference_calibration_norm     = EXPORT_REFERENCE_CALIBRATION_NORM
)

run_metadata_table <- participant_table |>
  dplyr::transmute(
    RUN_ID,
    participant_id,
    sex,
    height_cm,
    mass_kg,
    dominant_side,
    practice_level,
    run_number,
    mvc_ref
  )

reference_level_feature_table_norm_export <- reference_level_feature_table_norm |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

ex_dyn_mid1s_feature_table_norm_export <- ex_dyn_mid1s_feature_table_norm |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

ex_sta_mid1s_feature_table_norm_export <- ex_sta_mid1s_feature_table_norm |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

ex_dyn_emg_torque_ratio_table_export <- ex_dyn_emg_torque_ratio_table |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

ex_sta_emg_torque_ratio_table_export <- ex_sta_emg_torque_ratio_table |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

emg_torque_ratio_slope_summary_table_export <- emg_torque_ratio_slope_summary_table |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

reference_calibration_table_norm_export <- reference_calibration_table_norm |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, participant_id = .env$participant_id, .before = 1)

analysis_export_map <- list(
  run_metadata                  = run_metadata_table,
  reference_level_features_norm = reference_level_feature_table_norm_export,
  ex_dyn_features_norm          = ex_dyn_mid1s_feature_table_norm_export,
  ex_sta_features_norm          = ex_sta_mid1s_feature_table_norm_export,
  ex_dyn_ratio_table            = ex_dyn_emg_torque_ratio_table_export,
  ex_sta_ratio_table            = ex_sta_emg_torque_ratio_table_export,
  ratio_slope_summary           = emg_torque_ratio_slope_summary_table_export,
  reference_calibration_norm    = reference_calibration_table_norm_export
)

analysis_export_file_map <- c(
  run_metadata                  = "run_metadata.parquet",
  reference_level_features_norm = "reference_level_feature_table_norm.parquet",
  ex_dyn_features_norm          = "ex_dyn_mid1s_feature_table_norm.parquet",
  ex_sta_features_norm          = "ex_sta_mid1s_feature_table_norm.parquet",
  ex_dyn_ratio_table            = "ex_dyn_emg_torque_ratio_table.parquet",
  ex_sta_ratio_table            = "ex_sta_emg_torque_ratio_table.parquet",
  ratio_slope_summary           = "emg_torque_ratio_slope_summary_table.parquet",
  reference_calibration_norm    = "reference_calibration_table_norm.parquet"
)

if (isTRUE(EXPORT_ANALYSIS_TABLES)) {
  selected_table_names <- names(analysis_export_flags)[analysis_export_flags]
  if (length(selected_table_names) == 0L) {
    stop("EXPORT_ANALYSIS_TABLES=TRUE but no analysis table selector is set to TRUE.")
  }

  analysis_export_root <- here::here("results", "R_ANALYSIS", RUN_ID)
  dir.create(analysis_export_root, recursive = TRUE, showWarnings = FALSE)

  for (table_name in selected_table_names) {
    arrow::write_parquet(
      analysis_export_map[[table_name]],
      sink = file.path(analysis_export_root, analysis_export_file_map[[table_name]])
    )
  }

  message("\n[R ANALYSIS EXPORT] Tables -> ", normalizePath(analysis_export_root, winslash = "/", mustWork = FALSE))
}
```

