---
title: "Individual NIRS Analysis Pipeline"
author: "Morgan Virolan"
date: "2026-05-16"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: cosmo

params:
  RUN_ID: "209CeJa_20251105"
---

# Overview

Single-run NIRS analysis: load ΔHHb, compute sequence baselines/features, build EX_DYN/EX_STA summaries, and write cohort-ready tables.

# 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(arrow)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(glue)
options(dplyr.summarise.inform = FALSE)
source(here::here("sources", "R", "config", "amstrom_style.R"))
```

```{r controls}
RUN_ID <- params$RUN_ID

PIPELINE_MODE          <- "full_run"
SHOW_NIRS_PLOTS        <- identical(PIPELINE_MODE, "full_run")
EXPORT_NIRS_PLOTS      <- identical(PIPELINE_MODE, "full_run")
EXPORT_ANALYSIS_TABLES <- identical(PIPELINE_MODE, "full_run")

NIRS_SIGNAL_COL <- "nirs_hhb_tx1"   # ΔHHb, Tx1 = 35 mm source-detector distance
NIRS_FS_HZ      <- 10.0             # nominal sampling rate (Oxysoft export)

# Optional low-pass Butterworth filter on the full continuous signal.
# Off by default: per-window averaging already acts as an implicit low-pass filter,
# and isometric contractions on a fixed dynamometer produce minimal motion artifact.
APPLY_NIRS_FILTER  <- FALSE
NIRS_FILTER_CUTOFF <- 0.3           # Hz (1st-order Butterworth LP if enabled)

# EX_STA windowing (parallel to BIA pipeline)
EX_STA_WINDOW_LENGTH_S      <- 1.0
EX_STA_WINDOW_STEP_S        <- 1.0

# EX_DYN onset slope: first N samples per contraction (1 s at 10 Hz)
EXDYN_SLOPE_SAMPLES <- 10L

# Pre-exercise baseline window relative to first VC==1 onset in each sequence
BASELINE_START_S <- -10.0
BASELINE_END_S   <- -5.0

critical_seqs <- c("MVC_REF", "SVC_REF", "EX_DYN", "EX_STA", "MVC_RECOV_DYN", "MVC_RECOV_STA")
```

# 2. Load exported data

```{r load_data}
export_root   <- here::here("results", "DATA_EXPORT")
run_directory <- file.path(export_root, RUN_ID)

master_table <- arrow::read_parquet(
  file.path(run_directory, "master_index_grid.parquet"),
  col_select = c("time_index", "time_ref_s", "SEQ_index", "SEQ", "VC", "VC_count")
)

# Join time_ref_s from the master grid — preprocessed_nirs.parquet carries time_index
# and SEQ/VC/VC_count labels but not the master reference time axis.
nirs_raw <- arrow::read_parquet(
  file.path(run_directory, "preprocessed_nirs.parquet")
) |>
  dplyr::left_join(
    dplyr::select(master_table, time_index, time_ref_s),
    by = "time_index"
  ) |>
  dplyr::arrange(.data$time_ref_s)

participant_table <- arrow::read_parquet(
  file.path(run_directory, "participant.parquet")
) |>
  dplyr::mutate(RUN_ID = .env$RUN_ID, .before = 1)

participant_id <- participant_table$participant_id[[1]]

cat("RUN_ID:", RUN_ID, "\n")
cat("participant_id:", participant_id, "\n")
cat("nrow(master_table):", nrow(master_table), "\n")
cat("nrow(nirs_raw):    ", nrow(nirs_raw), "\n")
cat("signal column:     ", NIRS_SIGNAL_COL, "\n")
```

# 3. Sequence validity gate

```{r check_nirs_sequence_validity}
# Validity gate: every critical sequence must have VC==1 samples, else stop.
seqs_with_contractions <- nirs_raw |>
  dplyr::filter(.data$VC == 1L) |>
  dplyr::distinct(.data$SEQ) |>
  dplyr::pull(.data$SEQ)

missing_critical <- setdiff(critical_seqs, seqs_with_contractions)

if (length(missing_critical) > 0L) {
  stop(glue(
    "NIRS validity check FAILED for {RUN_ID}: ",
    "no VC==1 samples found in critical sequence(s): ",
    paste(missing_critical, collapse = ", ")
  ))
}

cat("NIRS sequence validity OK — all critical sequences have VC==1 samples.\n")
cat("Sequences with contractions:", paste(sort(seqs_with_contractions), collapse = ", "), "\n")
```

# 4. (Optional) Low-pass filter

```{r apply_nirs_filter}
if (isTRUE(APPLY_NIRS_FILTER)) {
  library(signal)
  bf       <- signal::butter(1L, NIRS_FILTER_CUTOFF / (NIRS_FS_HZ / 2), type = "low")
  raw_vals <- nirs_raw[[NIRS_SIGNAL_COL]]

  # Interpolate internal NAs before filtering to prevent them from propagating.
  if (any(is.na(raw_vals))) {
    idx       <- seq_along(raw_vals)
    good      <- is.finite(raw_vals)
    raw_vals  <- approx(idx[good], raw_vals[good], xout = idx, rule = 2L)$y
  }

  nirs_raw[[NIRS_SIGNAL_COL]] <- as.numeric(signal::filtfilt(bf, raw_vals))
  cat(glue("LP filter applied: 1st-order Butterworth at {NIRS_FILTER_CUTOFF} Hz.\n"))
} else {
  cat("No filter applied (APPLY_NIRS_FILTER = FALSE).\n")
}
```

# 5. Baseline computation

```{r compute_nirs_baseline}
# Pre-exercise baseline: mean ΔHHb in [onset - 10 s, onset - 5 s) for each sequence.
# The 5 s gap before onset avoids anticipatory contraction artefacts.
# Baseline is stored for subtraction in each feature table (hhb_from_baseline_umol).

compute_seq_baseline <- function(nirs_df, seq_name) {
  onset_s <- nirs_df |>
    dplyr::filter(.data$SEQ == seq_name, .data$VC == 1L) |>
    dplyr::summarise(onset = min(.data$time_ref_s, na.rm = TRUE)) |>
    dplyr::pull(.data$onset)

  if (length(onset_s) == 0L || !is.finite(onset_s)) {
    return(tibble::tibble(
      SEQ = seq_name, hhb_baseline_umol = NA_real_,
      o2hb_baseline_umol = NA_real_, thb_baseline_umol = NA_real_,
      n_baseline_samples = 0L
    ))
  }

  return(nirs_df |>
    dplyr::filter(
      .data$time_ref_s >= onset_s + BASELINE_START_S,
      .data$time_ref_s <  onset_s + BASELINE_END_S
    ) |>
    dplyr::summarise(
      SEQ                = seq_name,
      hhb_baseline_umol  = mean(.data[[NIRS_SIGNAL_COL]], na.rm = TRUE),  # resting ΔHHb
      o2hb_baseline_umol = mean(.data$nirs_o2hb_tx1, na.rm = TRUE),       # resting O2Hb
      thb_baseline_umol  = mean(.data$nirs_thb_tx1, na.rm = TRUE),        # resting tHb
      n_baseline_samples = sum(is.finite(.data[[NIRS_SIGNAL_COL]]))       # samples in the rest window
    ))
}

nirs_baseline_table <- dplyr::bind_rows(
  lapply(critical_seqs, compute_seq_baseline, nirs_df = nirs_raw)
)

# Fresh (global) baseline: mean ΔHHb in [MVC_REF onset − 10s, onset − 5s).
# This is before any voluntary contraction — the cleanest resting state in the session.
# Used for hhb_from_baseline_umol in all feature tables (preserves carry-over across sequences).
fresh_baseline_umol <- nirs_baseline_table |>
  dplyr::filter(.data$SEQ == "MVC_REF") |>
  dplyr::pull(.data$hhb_baseline_umol)

print(nirs_baseline_table)
cat("Fresh baseline (before MVC_REF):", round(fresh_baseline_umol, 4), "umol/L\n")
```

# 6. Reference and checkpoint table

```{r build_reference_feature_table}
# Per-VC mean ΔHHb for MVC_REF, SVC_REF (3 intensity levels), and recovery sequences.

ref_seqs <- c("MVC_REF", "SVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")

nirs_reference_feature_table <- nirs_raw |>
  dplyr::filter(.data$SEQ %in% ref_seqs, .data$VC == 1L) |>
  dplyr::group_by(.data$SEQ, .data$VC_count) |>
  # Coherence with EMG/torque feature window: keep only the middle 1 s of each
  # contraction. See `extract_middle_1s_window` in emg_individual_pipeline.Rmd.
  dplyr::mutate(
    t_mid     = (min(.data$time_ref_s) + max(.data$time_ref_s)) / 2,
    mid_start = t_mid - 0.5,
    mid_end   = t_mid + 0.5
  ) |>
  dplyr::filter(.data$time_ref_s >= mid_start, .data$time_ref_s < mid_end) |>
  dplyr::summarise(
    mean_hhb_umol  = mean(.data[[NIRS_SIGNAL_COL]], na.rm = TRUE),  # mean ΔHHb over the mid-1s window
    mean_o2hb_umol = mean(.data$nirs_o2hb_tx1, na.rm = TRUE),       # mean O2Hb
    mean_thb_umol  = mean(.data$nirs_thb_tx1, na.rm = TRUE),        # mean tHb
    time_mid_s     = mean(.data$time_ref_s, na.rm = TRUE),          # window midpoint time
    n_samples      = sum(is.finite(.data[[NIRS_SIGNAL_COL]])),      # samples in the window
    .groups        = "drop"
  ) |>
  dplyr::left_join(nirs_baseline_table, by = "SEQ") |>
  dplyr::mutate(
    RUN_ID          = .env$RUN_ID,
    participant_id  = .env$participant_id,
    source_sequence = .data$SEQ,
    reference_level = dplyr::case_when(
      .data$SEQ == "SVC_REF" & .data$VC_count == 1L ~ "30%",
      .data$SEQ == "SVC_REF" & .data$VC_count == 2L ~ "50%",
      .data$SEQ == "SVC_REF" & .data$VC_count == 3L ~ "70%",
      .data$SEQ == "MVC_REF"                         ~ "100%",
      TRUE                                            ~ NA_character_
    ),
    state_label = dplyr::case_when(
      .data$SEQ == "SVC_REF" ~ paste0("SVC_", .data$reference_level),
      TRUE                   ~ .data$SEQ
    ),
    hhb_from_baseline_umol  = .data$mean_hhb_umol - .env$fresh_baseline_umol,
    hhb_from_seq_rest_umol  = .data$mean_hhb_umol - .data$hhb_baseline_umol,
    o2hb_from_seq_rest_umol = .data$mean_o2hb_umol - .data$o2hb_baseline_umol,
    thb_from_seq_rest_umol  = .data$mean_thb_umol - .data$thb_baseline_umol,
    evidence_status = "provisional",
    .before = 1
  ) |>
  dplyr::select(
    RUN_ID, participant_id,
    source_sequence, state_label, reference_level,
    VC_count,
    mean_hhb_umol, hhb_from_baseline_umol, hhb_from_seq_rest_umol,
    mean_o2hb_umol, o2hb_from_seq_rest_umol,
    mean_thb_umol, thb_from_seq_rest_umol,
    hhb_baseline_umol, n_baseline_samples,
    time_mid_s, n_samples, evidence_status
  ) |>
  dplyr::arrange(.data$source_sequence, .data$VC_count)
```

# 7. EX_DYN feature table

```{r build_ex_dyn_feature_table}
# EX_DYN at 10 Hz: ~30 samples per contraction (3 s ON), ~20 samples per rest (2 s OFF).
#
# VC_count is 0 for all rest periods — cummax(VC_count) on time-sorted rows fills in
# the preceding contraction index, enabling rest-period features to be linked to their
# contraction.

ex_dyn_rows <- nirs_raw |>
  dplyr::filter(.data$SEQ == "EX_DYN") |>
  dplyr::arrange(.data$time_ref_s) |>
  dplyr::mutate(preceding_vc = cummax(.data$VC_count))

ex_dyn_baseline_umol <- nirs_baseline_table |>
  dplyr::filter(.data$SEQ == "EX_DYN") |>
  dplyr::pull(.data$hhb_baseline_umol)

ex_dyn_o2hb_baseline_umol <- nirs_baseline_table |>
  dplyr::filter(.data$SEQ == "EX_DYN") |>
  dplyr::pull(.data$o2hb_baseline_umol)

ex_dyn_thb_baseline_umol <- nirs_baseline_table |>
  dplyr::filter(.data$SEQ == "EX_DYN") |>
  dplyr::pull(.data$thb_baseline_umol)

# --- Contraction features (VC==1): mean, desat amplitude, onset slope ---
# `mean_hhb_umol` uses the middle 1 s of each contraction (coherence with the
# EMG/torque feature window, see `extract_middle_1s_window` in
# emg_individual_pipeline.Rmd). `hhb_desat_umol` and `hhb_slope_umol_s` keep the
# full VC==1 window: desat needs the full range (max-min) to be meaningful,
# slope is by definition computed over the onset samples.
ex_dyn_contraction_full <- ex_dyn_rows |>
  dplyr::filter(.data$VC == 1L) |>
  dplyr::group_by(.data$VC_count) |>
  dplyr::summarise(
    hhb_desat_umol = max(.data[[NIRS_SIGNAL_COL]], na.rm = TRUE) -
                     min(.data[[NIRS_SIGNAL_COL]], na.rm = TRUE),
    hhb_slope_umol_s = {
      # Slope of ΔHHb in the first EXDYN_SLOPE_SAMPLES (1 s) of each contraction.
      # Rows are time-sorted, so head() gives the onset samples.
      sig_v <- head(.data[[NIRS_SIGNAL_COL]], .env$EXDYN_SLOPE_SAMPLES)
      t_v   <- head(.data$time_ref_s, .env$EXDYN_SLOPE_SAMPLES)
      ok    <- is.finite(sig_v) & is.finite(t_v)
      if (sum(ok) >= 3L) {
        unname(coef(lm(sig_v[ok] ~ t_v[ok]))[[2]])
      } else {
        NA_real_
      }
    },
    time_mid_s     = mean(.data$time_ref_s, na.rm = TRUE),
    n_samples_cont = sum(is.finite(.data[[NIRS_SIGNAL_COL]])),
    .groups        = "drop"
  )

ex_dyn_contraction_mid1s <- ex_dyn_rows |>
  dplyr::filter(.data$VC == 1L) |>
  dplyr::group_by(.data$VC_count) |>
  dplyr::mutate(
    t_mid     = (min(.data$time_ref_s) + max(.data$time_ref_s)) / 2,
    mid_start = t_mid - 0.5,
    mid_end   = t_mid + 0.5
  ) |>
  dplyr::filter(.data$time_ref_s >= mid_start, .data$time_ref_s < mid_end) |>
  dplyr::summarise(
    mean_hhb_umol  = mean(.data[[NIRS_SIGNAL_COL]], na.rm = TRUE),
    mean_o2hb_umol = mean(.data$nirs_o2hb_tx1, na.rm = TRUE),
    mean_thb_umol  = mean(.data$nirs_thb_tx1, na.rm = TRUE),
    .groups        = "drop"
  )

ex_dyn_contraction <- ex_dyn_contraction_full |>
  dplyr::left_join(ex_dyn_contraction_mid1s, by = "VC_count") |>
  dplyr::select(
    VC_count, mean_hhb_umol, mean_o2hb_umol, mean_thb_umol,
    hhb_desat_umol, hhb_slope_umol_s,
    time_mid_s, n_samples_cont
  )

# --- Rest features (VC==0): reoxygenation amplitude ---
# hhb_sat = first(ΔHHb at rest onset) - min(ΔHHb during rest).
# HHb peaks near the end of contraction then falls during rest as O2 is resupplied;
# the amplitude of that fall is the reoxy metric.
ex_dyn_rest <- ex_dyn_rows |>
  dplyr::filter(.data$VC == 0L, .data$preceding_vc > 0L) |>
  dplyr::group_by(.data$preceding_vc) |>
  dplyr::summarise(
    hhb_sat_umol   = dplyr::first(.data[[NIRS_SIGNAL_COL]]) -
                     min(.data[[NIRS_SIGNAL_COL]], na.rm = TRUE),
    n_samples_rest = sum(is.finite(.data[[NIRS_SIGNAL_COL]])),
    .groups        = "drop"
  ) |>
  dplyr::rename(VC_count = preceding_vc)

nirs_ex_dyn_feature_table <- ex_dyn_contraction |>
  dplyr::left_join(ex_dyn_rest, by = "VC_count") |>
  dplyr::mutate(
    RUN_ID                  = .env$RUN_ID,
    participant_id          = .env$participant_id,
    source_sequence         = "EX_DYN",
    hhb_from_baseline_umol  = .data$mean_hhb_umol - .env$fresh_baseline_umol,
    hhb_from_seq_rest_umol  = .data$mean_hhb_umol - .env$ex_dyn_baseline_umol,
    o2hb_from_seq_rest_umol = .data$mean_o2hb_umol - .env$ex_dyn_o2hb_baseline_umol,
    thb_from_seq_rest_umol  = .data$mean_thb_umol - .env$ex_dyn_thb_baseline_umol,
    evidence_status         = "provisional",
    .before                 = 1
  ) |>
  dplyr::select(
    RUN_ID, participant_id, source_sequence,
    VC_count,
    mean_hhb_umol, hhb_from_baseline_umol, hhb_from_seq_rest_umol,
    mean_o2hb_umol, o2hb_from_seq_rest_umol,
    mean_thb_umol, thb_from_seq_rest_umol,
    hhb_desat_umol, hhb_slope_umol_s, hhb_sat_umol,
    time_mid_s, n_samples_cont, n_samples_rest,
    evidence_status
  ) |>
  dplyr::filter(.data$VC_count <= 60L) |>
  dplyr::arrange(.data$VC_count)
```

# 8. EX_STA feature table

```{r build_ex_sta_window_table}
# 1 s non-overlapping windows over the continuous EX_STA VC==1 signal.
# Onset is determined by torque threshold (parallel to BIA pipeline).

ex_sta_samples <- nirs_raw |>
  dplyr::filter(.data$SEQ == "EX_STA", .data$VC == 1L) |>
  dplyr::arrange(.data$time_ref_s)

if (nrow(ex_sta_samples) == 0L) {
  stop(glue("EX_STA contains no VC==1 NIRS samples for {RUN_ID}."))
}

t_sta_start    <- min(ex_sta_samples$time_ref_s)
t_sta_end      <- max(ex_sta_samples$time_ref_s)
window_start_s <- seq(t_sta_start, t_sta_end - EX_STA_WINDOW_LENGTH_S, by = EX_STA_WINDOW_STEP_S)

ex_sta_window_layout <- tibble::tibble(
  window_id      = seq_along(window_start_s),
  window_start_s = window_start_s,
  window_end_s   = window_start_s + EX_STA_WINDOW_LENGTH_S,
  window_mid_s   = window_start_s + EX_STA_WINDOW_LENGTH_S / 2
)

ex_sta_window_raw <- ex_sta_samples |>
  dplyr::mutate(
    window_id = floor((.data$time_ref_s - t_sta_start) / EX_STA_WINDOW_STEP_S) + 1L
  ) |>
  dplyr::filter(.data$window_id >= 1L, .data$window_id <= nrow(ex_sta_window_layout)) |>
  dplyr::left_join(ex_sta_window_layout, by = "window_id") |>
  dplyr::group_by(.data$window_id, .data$window_mid_s) |>
  dplyr::summarise(
    mean_hhb_umol  = mean(.data[[NIRS_SIGNAL_COL]], na.rm = TRUE),  # mean ΔHHb in this 1s window
    mean_o2hb_umol = mean(.data$nirs_o2hb_tx1, na.rm = TRUE),       # mean O2Hb
    mean_thb_umol  = mean(.data$nirs_thb_tx1, na.rm = TRUE),        # mean tHb
    n_samples      = sum(is.finite(.data[[NIRS_SIGNAL_COL]])),      # samples in this window
    .groups        = "drop"
  ) |>
  dplyr::arrange(.data$window_id)
```

```{r filter_ex_sta_onset_windows}
# Common onset keeps the EX_STA denominator aligned with EMG and BIA exports.
EX_STA_COMMON_ONSET_WINDOW_ID <- 2L

ex_sta_onset_window_id <- EX_STA_COMMON_ONSET_WINDOW_ID
ex_sta_onset_mid_s     <- ex_sta_window_layout$window_mid_s[[ex_sta_onset_window_id]]

window_index_lookup <- ex_sta_window_layout |>
  dplyr::filter(.data$window_id >= ex_sta_onset_window_id) |>
  dplyr::mutate(
    window_index             = dplyr::row_number(),
    window_time_from_start_s = .data$window_mid_s - ex_sta_onset_mid_s
  ) |>
  dplyr::select(window_id, window_index, window_time_from_start_s)

ex_sta_baseline_umol <- nirs_baseline_table |>
  dplyr::filter(.data$SEQ == "EX_STA") |>
  dplyr::pull(.data$hhb_baseline_umol)

ex_sta_o2hb_baseline_umol <- nirs_baseline_table |>
  dplyr::filter(.data$SEQ == "EX_STA") |>
  dplyr::pull(.data$o2hb_baseline_umol)

ex_sta_thb_baseline_umol <- nirs_baseline_table |>
  dplyr::filter(.data$SEQ == "EX_STA") |>
  dplyr::pull(.data$thb_baseline_umol)

nirs_ex_sta_feature_table <- ex_sta_window_raw |>
  dplyr::filter(.data$window_id >= ex_sta_onset_window_id) |>
  dplyr::left_join(window_index_lookup, by = "window_id") |>
  dplyr::mutate(
    RUN_ID                  = .env$RUN_ID,
    participant_id          = .env$participant_id,
    source_sequence         = "EX_STA",
    hhb_from_baseline_umol  = .data$mean_hhb_umol - .env$fresh_baseline_umol,
    hhb_from_seq_rest_umol  = .data$mean_hhb_umol - .env$ex_sta_baseline_umol,
    o2hb_from_seq_rest_umol = .data$mean_o2hb_umol - .env$ex_sta_o2hb_baseline_umol,
    thb_from_seq_rest_umol  = .data$mean_thb_umol - .env$ex_sta_thb_baseline_umol,
    evidence_status         = "provisional",
    .before                 = 1
  ) |>
  dplyr::select(
    RUN_ID, participant_id, source_sequence,
    window_id, window_index,
    window_mid_s, window_time_from_start_s,
    mean_hhb_umol, hhb_from_baseline_umol, hhb_from_seq_rest_umol,
    mean_o2hb_umol, o2hb_from_seq_rest_umol,
    mean_thb_umol, thb_from_seq_rest_umol,
    n_samples, evidence_status
  ) |>
  dplyr::arrange(.data$window_index)
```

# 9. Run-level summary table

```{r build_nirs_run_summary_table}
# Run-level scalars: linear slope over time, and last-minus-first delta, computed for each
# listed NIRS variable on both EX_DYN (vs VC_count) and EX_STA (vs window_index).
make_nirs_slope <- function(df, x_col, value_col, analysis_block, source_seq) {
  df_clean <- df |>
    dplyr::filter(is.finite(.data[[x_col]]), is.finite(.data[[value_col]])) |>
    dplyr::arrange(.data[[x_col]])
  if (nrow(df_clean) < 3L || dplyr::n_distinct(df_clean[[x_col]]) < 3L) return(NULL)
  fit <- lm(reformulate(x_col, response = value_col), data = df_clean)
  return(tibble::tibble(
    analysis_block   = analysis_block,
    summary_variable = value_col,
    summary_type     = paste0(x_col, "_linear_slope"),
    source_sequence  = source_seq,
    summary_value    = unname(coef(fit)[[2]]),
    n_points         = nrow(df_clean)
  ))
}

make_nirs_last_minus_first <- function(df, x_col, value_col, analysis_block, source_seq) {
  df_clean <- df |>
    dplyr::filter(is.finite(.data[[x_col]]), is.finite(.data[[value_col]])) |>
    dplyr::arrange(.data[[x_col]])
  if (nrow(df_clean) < 2L) return(NULL)
  return(tibble::tibble(
    analysis_block   = analysis_block,
    summary_variable = value_col,
    summary_type     = "last_minus_first",
    source_sequence  = source_seq,
    summary_value    = dplyr::last(df_clean[[value_col]]) - dplyr::first(df_clean[[value_col]]),
    n_points         = nrow(df_clean)
  ))
}

ex_dyn_summary_vars <- c("mean_hhb_umol", "hhb_from_baseline_umol", "hhb_from_seq_rest_umol",
                         "hhb_desat_umol", "hhb_slope_umol_s", "hhb_sat_umol",
                         "o2hb_from_seq_rest_umol", "thb_from_seq_rest_umol")
ex_sta_summary_vars <- c("mean_hhb_umol", "hhb_from_baseline_umol", "hhb_from_seq_rest_umol",
                         "o2hb_from_seq_rest_umol", "thb_from_seq_rest_umol")

nirs_run_summary_table <- dplyr::bind_rows(c(
  lapply(ex_dyn_summary_vars, function(v)
    make_nirs_slope(nirs_ex_dyn_feature_table, "VC_count", v, "nirs_ex_dyn", "EX_DYN")),
  lapply(ex_dyn_summary_vars, function(v)
    make_nirs_last_minus_first(nirs_ex_dyn_feature_table, "VC_count", v, "nirs_ex_dyn", "EX_DYN")),
  lapply(ex_sta_summary_vars, function(v)
    make_nirs_slope(nirs_ex_sta_feature_table, "window_index", v, "nirs_ex_sta", "EX_STA")),
  lapply(ex_sta_summary_vars, function(v)
    make_nirs_last_minus_first(nirs_ex_sta_feature_table, "window_index", v, "nirs_ex_sta", "EX_STA"))
)) |>
  dplyr::mutate(
    RUN_ID          = .env$RUN_ID,
    participant_id  = .env$participant_id,
    evidence_status = "provisional",
    .before         = 1
  ) |>
  dplyr::select(
    RUN_ID, participant_id,
    analysis_block, summary_variable,
    summary_type, source_sequence,
    summary_value, n_points, evidence_status
  ) |>
  dplyr::arrange(.data$analysis_block, .data$summary_variable, .data$summary_type)
```

# 9b. Signal quality QC

```{r nirs_signal_qc}
# Spike threshold (session-wide): clear motion artifact / probe displacement.
# EX_STA bad-window threshold: probe movement during sustained hold (systematic in this cohort).
NIRS_SPIKE_THRESHOLD_UMOL       <- 50L
NIRS_EXSTA_SPIKE_THRESHOLD_UMOL <- 15L

n_session_spikes <- nirs_raw |>
  dplyr::summarise(
    n = sum(abs(.data[[NIRS_SIGNAL_COL]]) > NIRS_SPIKE_THRESHOLD_UMOL, na.rm = TRUE)
  ) |>
  dplyr::pull(n)

n_exdyn_bad_windows <- if (nrow(nirs_ex_dyn_feature_table) > 0L) {
  nirs_ex_dyn_feature_table |>
    dplyr::summarise(
      n = sum(abs(.data$hhb_from_seq_rest_umol) > NIRS_EXSTA_SPIKE_THRESHOLD_UMOL, na.rm = TRUE)
    ) |>
    dplyr::pull(n)
} else { 0L }

n_exsta_bad_windows <- if (nrow(nirs_ex_sta_feature_table) > 0L) {
  nirs_ex_sta_feature_table |>
    dplyr::summarise(
      n = sum(abs(.data$hhb_from_seq_rest_umol) > NIRS_EXSTA_SPIKE_THRESHOLD_UMOL, na.rm = TRUE)
    ) |>
    dplyr::pull(n)
} else { 0L }

# tHb direction during EX_DYN contractions: mean tHb (delta from session start) across all reps.
# Healthy probe placement = blood pooling during contraction → tHb positive or near zero.
# Negative mean tHb = blood expelled from measurement volume → probe over compressible superficial
# structure (fascia / large vein), not the active capillary bed. Threshold -5 µmol/L confirmed
# on 261PoLe (mean = -13.54) vs clean run 802BlHu (mean = +5.91).
NIRS_THB_DEPLETION_THRESHOLD_UMOL <- -5L

exdyn_thb_mean <- if (nrow(nirs_ex_dyn_feature_table) > 0L) {
  round(mean(nirs_ex_dyn_feature_table$mean_thb_umol, na.rm = TRUE), 2)
} else { NA_real_ }

exdyn_thb_depletion_flag <- isTRUE(!is.na(exdyn_thb_mean) &&
                                     exdyn_thb_mean < NIRS_THB_DEPLETION_THRESHOLD_UMOL)

nirs_qc_table <- tibble::tibble(
  RUN_ID                   = RUN_ID,
  participant_id           = participant_id,
  n_session_spikes         = n_session_spikes,
  n_exdyn_bad_windows      = n_exdyn_bad_windows,
  n_exsta_bad_windows      = n_exsta_bad_windows,
  exsta_spike_flag         = n_exsta_bad_windows > 0L,
  exdyn_thb_mean_umol      = exdyn_thb_mean,
  exdyn_thb_depletion_flag = exdyn_thb_depletion_flag
)

cat("[NIRS QC]", RUN_ID,
    "| session spikes (>50 µmol/L):", n_session_spikes,
    "| EX_DYN bad windows:", n_exdyn_bad_windows,
    "| EX_STA bad windows:", n_exsta_bad_windows,
    "| EX_DYN tHb mean:", exdyn_thb_mean, "µmol/L",
    "| tHb depletion flag:", exdyn_thb_depletion_flag, "\n")
```

# 10. Diagnostic plots

```{r build_nirs_plots}
plot_export_map <- list()

# Baseline extraction windows: [onset + BASELINE_START_S, onset + BASELINE_END_S) per sequence.
# Used for geom_rect annotations — lets you visually verify the window lands in clean rest signal.
baseline_rects <- nirs_raw |>
  dplyr::filter(.data$SEQ %in% critical_seqs, .data$VC == 1L) |>
  dplyr::group_by(.data$SEQ) |>
  dplyr::summarise(onset_s = min(.data$time_ref_s, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(
    xmin = .data$onset_s + BASELINE_START_S,
    xmax = .data$onset_s + BASELINE_END_S
  )

# --- 10.1 Full session overview ---
nirs_session_plot <- ggplot(
  nirs_raw,
  aes(x = .data$time_ref_s, y = .data[[NIRS_SIGNAL_COL]], colour = .data$SEQ)
) +
  geom_rect(
    data        = baseline_rects,
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill        = "grey70", alpha = 0.35
  ) +
  geom_line(linewidth = 0.45, alpha = 0.85) +
  labs(
    title    = paste0("NIRS full session - ", RUN_ID),
    subtitle = "Shaded bands = baseline extraction windows [-10s, -5s) before each sequence onset",
    x        = "Session time (s)",
    y        = "ΔHHb (µmol/L)",
    colour   = "Sequence"
  ) +
  theme_amstrom_hgrid +
  theme(legend.position = "right")

plot_export_map[["session_overview"]] <- nirs_session_plot

# --- 10.2 Reference and checkpoint ---
ref_plot_levels <- c("SVC_30%", "SVC_50%", "SVC_70%", "MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")

nirs_ref_plot <- nirs_reference_feature_table |>
  dplyr::filter(!is.na(.data$state_label)) |>
  dplyr::group_by(.data$state_label) |>
  dplyr::summarise(
    hhb_from_baseline_umol = mean(.data$hhb_from_baseline_umol, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(state_label = factor(.data$state_label, levels = ref_plot_levels)) |>
  ggplot(aes(x = .data$state_label, y = .data$hhb_from_baseline_umol, group = 1)) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_nirs) +
  geom_point(shape = 15, size = 2.2, colour = col_nirs) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  labs(
    title = paste0("NIRS reference and checkpoint - ", RUN_ID),
    x     = NULL,
    y     = "ΔHHb from baseline (µmol/L)"
  ) +
  theme_amstrom_hgrid +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

plot_export_map[["reference_checkpoint"]] <- nirs_ref_plot

# --- 10.3 EX_DYN - ΔHHb from baseline trajectory ---
nirs_ex_dyn_plot <- ggplot(
  nirs_ex_dyn_feature_table,
  aes(x = .data$VC_count, y = .data$hhb_from_seq_rest_umol)
) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_nirs) +
  geom_point(shape = 15, size = 2.2, colour = col_nirs) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  labs(
    title = paste0("EX_DYN - ΔHHb trajectory - ", RUN_ID),
    x     = "Repetition",
    y     = "ΔHHb from seq. rest (µmol/L)"
  ) +
  theme_amstrom_hgrid

plot_export_map[["ex_dyn_hhb"]] <- nirs_ex_dyn_plot

# --- 10.4 EX_DYN - per-cycle features ---
ex_dyn_features_long <- nirs_ex_dyn_feature_table |>
  tidyr::pivot_longer(
    cols      = c("hhb_desat_umol", "hhb_sat_umol", "hhb_slope_umol_s"),
    names_to  = "feature",
    values_to = "value"
  ) |>
  dplyr::mutate(feature = factor(
    .data$feature,
    levels = c("hhb_desat_umol", "hhb_sat_umol", "hhb_slope_umol_s"),
    labels = c("Desat amplitude (µmol/L)", "Reoxy amplitude (µmol/L)", "Onset slope (µmol/L/s)")
  ))

nirs_ex_dyn_features_plot <- ggplot(
  ex_dyn_features_long,
  aes(x = .data$VC_count, y = .data$value)
) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_nirs) +
  geom_point(shape = 15, size = 2.2, colour = col_nirs) +
  facet_wrap(~ feature, ncol = 1, scales = "free_y") +
  labs(
    title = paste0("EX_DYN - per-cycle features - ", RUN_ID),
    x     = "Repetition",
    y     = NULL
  ) +
  theme_amstrom_hgrid

plot_export_map[["ex_dyn_features"]] <- nirs_ex_dyn_features_plot

# --- 10.5 EX_STA - ΔHHb trajectory ---
nirs_ex_sta_plot <- ggplot(
  nirs_ex_sta_feature_table,
  aes(x = .data$window_index, y = .data$hhb_from_seq_rest_umol)
) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_nirs) +
  geom_point(shape = 15, size = 2.2, colour = col_nirs) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +
  labs(
    title = paste0("EX_STA - ΔHHb trajectory - ", RUN_ID),
    x     = "Window index (1 s)",
    y     = "ΔHHb from seq. rest (µmol/L)"
  ) +
  theme_amstrom_hgrid

plot_export_map[["ex_sta_hhb"]] <- nirs_ex_sta_plot
```

```{r show_nirs_plots, fig.width=10, fig.height=6}
if (isTRUE(SHOW_NIRS_PLOTS)) {
  for (plot_name in names(plot_export_map)) {
    print(plot_export_map[[plot_name]])
  }
}
```

```{r export_nirs_plots, message=FALSE}
if (isTRUE(EXPORT_NIRS_PLOTS)) {
  qc_export_path <- here::here("results", "QC_EXPORT", "nirs_individual_analysis",
    paste0(RUN_ID, "_nirs_qc.pdf")
  )
  dir.create(dirname(qc_export_path), recursive = TRUE, showWarnings = FALSE)

  grDevices::pdf(file = qc_export_path, width = 10, height = 6, onefile = TRUE)
  for (plot_name in names(plot_export_map)) {
    print(plot_export_map[[plot_name]])
  }
  grDevices::dev.off()

  message(
    "\n[NIRS QC EXPORT] PDF -> ",
    normalizePath(qc_export_path, winslash = "/", mustWork = FALSE)
  )
}
```

# 11. Export analysis tables

```{r export_analysis_tables, message=FALSE}
# Write the per-run NIRS feature tables + QC table back to the run's export folder.
EXPORT_NIRS_REFERENCE <- TRUE
EXPORT_NIRS_EX_DYN    <- TRUE
EXPORT_NIRS_EX_STA    <- TRUE
EXPORT_NIRS_SUMMARY   <- TRUE
EXPORT_NIRS_QC        <- TRUE

analysis_export_map <- list(
  nirs_reference_feature_table = nirs_reference_feature_table,
  nirs_ex_dyn_feature_table    = nirs_ex_dyn_feature_table,
  nirs_ex_sta_feature_table    = nirs_ex_sta_feature_table,
  nirs_run_summary_table       = nirs_run_summary_table,
  nirs_qc_table                = nirs_qc_table
)

analysis_export_flags <- c(
  nirs_reference_feature_table = EXPORT_NIRS_REFERENCE,
  nirs_ex_dyn_feature_table    = EXPORT_NIRS_EX_DYN,
  nirs_ex_sta_feature_table    = EXPORT_NIRS_EX_STA,
  nirs_run_summary_table       = EXPORT_NIRS_SUMMARY,
  nirs_qc_table                = EXPORT_NIRS_QC
)

filename_map <- c(
  nirs_reference_feature_table = "nirs_reference_feature_table.parquet",
  nirs_ex_dyn_feature_table    = "nirs_ex_dyn_feature_table.parquet",
  nirs_ex_sta_feature_table    = "nirs_ex_sta_feature_table.parquet",
  nirs_run_summary_table       = "nirs_run_summary_table.parquet",
  nirs_qc_table                = "nirs_qc_table.parquet"
)

if (isTRUE(EXPORT_ANALYSIS_TABLES)) {
  for (tbl_name in names(analysis_export_map)) {
    if (!isTRUE(analysis_export_flags[[tbl_name]])) next
    out_path <- file.path(run_directory, filename_map[[tbl_name]])
    arrow::write_parquet(analysis_export_map[[tbl_name]], out_path)
    message(
      "[NIRS EXPORT] ", filename_map[[tbl_name]], " -> ",
      normalizePath(out_path, winslash = "/", mustWork = FALSE)
    )
  }
}
```
