---
title: "NIRS Cohort Analysis"
author: "Morgan Virolan"
date: "2026-05-16"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: cosmo
---

# Overview

Cohort NIRS analysis: pool per-run feature tables and plot ΔHHb reference, EX_DYN, and EX_STA summaries.

# 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(arrow)
library(dplyr)
library(tidyr)
library(ggplot2)
library(glue)

options(dplyr.summarise.inform = FALSE)
source(here::here("sources", "R", "config", "amstrom_style.R"))
source(here::here("sources", "R", "config", "cohort_config.R"))
```

```{r controls}
# Paths + the shared EX_STA truncation length and the summary-table print switch.
export_root          <- here::here("results", "DATA_EXPORT")
subset_registry_path <- here::here("results", "R_ANALYSIS", "subset_registry", "multimodal_subset_registry.parquet"
)

# Shared cross-modal EX_STA duration from R/config/cohort_config.R (torque-based, 10% drop threshold).
STA_N_MIN <- GLOBAL_STA_N_MIN
PRINT_SUMMARY_TABLES <- FALSE
```

```{r load_registry}
# All four cohort notebooks pool from cohort_final (four-way intersection minus excluded runs).
# See R/config/cohort_config.R for exclusion lists. N is printed below after filtering.
subset_registry <- arrow::read_parquet(subset_registry_path)

nirs_run_ids <- subset_registry |>
  dplyr::filter(.data$analysis_block == "cohort_final", .data$included == 1L) |>
  dplyr::pull(.data$RUN_ID) |>
  sort()

cat("Cohort N (cohort_final):", length(nirs_run_ids), "\n")
print(nirs_run_ids)
```

```{r pool_tables}
# pool_nirs_table: read one per-run NIRS export for every cohort run and stack them into a
# single table. Warns (does not stop) on a missing file. Called once per feature table below.
pool_nirs_table <- function(run_ids, filename) {
  paths  <- file.path(export_root, run_ids, filename)
  exists <- file.exists(paths)
  if (any(!exists)) {
    warning(glue(
      "Missing {sum(!exists)} parquet(s) for {filename}: ",
      paste(run_ids[!exists], collapse = ", ")
    ))
  }
  return(dplyr::bind_rows(lapply(paths[exists], arrow::read_parquet)))
}

nirs_ref_pooled     <- pool_nirs_table(nirs_run_ids, "nirs_reference_feature_table.parquet")
nirs_ex_dyn_pooled  <- pool_nirs_table(nirs_run_ids, "nirs_ex_dyn_feature_table.parquet")
nirs_ex_sta_pooled  <- pool_nirs_table(nirs_run_ids, "nirs_ex_sta_feature_table.parquet")
nirs_summary_pooled <- pool_nirs_table(nirs_run_ids, "nirs_run_summary_table.parquet")

cat("Pooled runs - ref:",     dplyr::n_distinct(nirs_ref_pooled$RUN_ID),
    "| ex_dyn:", dplyr::n_distinct(nirs_ex_dyn_pooled$RUN_ID),
    "| ex_sta:", dplyr::n_distinct(nirs_ex_sta_pooled$RUN_ID), "\n")
```

# 2. Reference

Cohort mean ΔHHb (from the fresh global baseline) at each reference/checkpoint state. Error
bars are mean ± SD across runs.

```{r ref_checkpoint_plot, fig.width=7, fig.height=4}
# Summarise to cohort mean/SD ΔHHb per reference state, then plot.
ref_levels <- c("SVC_30%", "SVC_50%", "SVC_70%", "MVC_REF", "MVC_RECOV_DYN", "MVC_RECOV_STA")

ref_summary <- nirs_ref_pooled |>
  dplyr::filter(.data$state_label %in% ref_levels) |>
  dplyr::mutate(state_label = factor(.data$state_label, levels = ref_levels)) |>
  dplyr::group_by(.data$state_label) |>
  dplyr::summarise(
    mean_hhb = mean(.data$hhb_from_baseline_umol, na.rm = TRUE),  # cohort mean ΔHHb (from fresh baseline)
    sd_hhb   = sd(.data$hhb_from_baseline_umol, na.rm = TRUE),    # between-run SD
    n        = dplyr::n_distinct(.data$RUN_ID),                  # runs at this state
    .groups  = "drop"
  )

nirs_ref_plot <- ggplot(ref_summary, aes(x = .data$state_label, y = .data$mean_hhb, group = 1)) +
  geom_errorbar(
    aes(ymin = mean_hhb - sd_hhb, ymax = mean_hhb + sd_hhb),
    width = 0, colour = col_nirs, linewidth = 0.4, alpha = 0.28
  ) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_nirs) +
  geom_point(shape = 15, size = 2.2, colour = col_nirs) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +  # zero = no change from baseline
  labs(
    title    = "NIRS - Reference (cohort mean +/- SD)",
    subtitle = glue("N = {max(ref_summary$n)} runs"),
    x        = NULL,
    y        = "delta HHb from baseline (umol/L)"
  ) +
  theme_amstrom_hgrid +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

print(nirs_ref_plot)
```

# 3. EX_DYN trajectories

Cohort mean per repetition (ΔHHb from sequence rest), then the per-cycle features (desat /
reoxy / onset slope) pooled across runs. Each chunk = one summarise + one plot; error bars
are mean ± SD.

```{r ex_dyn_hhb_plot, fig.width=7, fig.height=4}
# Cohort mean/SD ΔHHb (from sequence rest) per repetition, then the trajectory plot.
ex_dyn_summary <- nirs_ex_dyn_pooled |>
  dplyr::group_by(.data$VC_count) |>
  dplyr::summarise(
    mean_hhb = mean(.data$hhb_from_seq_rest_umol, na.rm = TRUE),  # cohort mean ΔHHb (from seq rest)
    sd_hhb   = sd(.data$hhb_from_seq_rest_umol, na.rm = TRUE),    # between-run SD
    n        = dplyr::n(),                                       # rows pooled at this point
    .groups  = "drop"
  )

nirs_ex_dyn_plot <- ggplot(ex_dyn_summary, aes(x = .data$VC_count, y = .data$mean_hhb)) +
  geom_errorbar(
    aes(ymin = mean_hhb - sd_hhb, ymax = mean_hhb + sd_hhb),
    width = 0, colour = col_nirs, linewidth = 0.4, alpha = 0.28
  ) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_nirs) +
  geom_point(shape = 15, size = 2.2, colour = col_nirs) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +  # zero = no change from baseline
  labs(
    title    = "EX_DYN - delta HHb from seq. rest (cohort mean +/- SD)",
    subtitle = glue("N = {max(ex_dyn_summary$n)} runs"),
    x        = "Repetition",
    y        = "delta HHb from seq. rest (umol/L)"
  ) +
  theme_amstrom_hgrid

print(nirs_ex_dyn_plot)
```

```{r ex_dyn_features_plot, fig.width=7, fig.height=7}
# Per-cycle features (desat / reoxy / onset slope): pivot long, cohort mean/SD per rep, facet.
feature_labels <- c(
  hhb_desat_umol   = "Desat amplitude (umol/L)",
  hhb_sat_umol     = "Reoxy amplitude (umol/L)",
  hhb_slope_umol_s = "Onset slope (umol/L/s)"
)

ex_dyn_features_summary <- nirs_ex_dyn_pooled |>
  tidyr::pivot_longer(
    cols      = c("hhb_desat_umol", "hhb_sat_umol", "hhb_slope_umol_s"),
    names_to  = "feature",
    values_to = "value"
  ) |>
  dplyr::mutate(
    feature = factor(.data$feature, levels = names(feature_labels), labels = unname(feature_labels))
  ) |>
  dplyr::group_by(.data$feature, .data$VC_count) |>
  dplyr::summarise(
    mean_val = mean(.data$value, na.rm = TRUE),  # cohort mean of this feature at this rep
    sd_val   = sd(.data$value, na.rm = TRUE),    # between-run SD
    .groups  = "drop"
  )

nirs_ex_dyn_features_plot <- ggplot(
  ex_dyn_features_summary,
  aes(x = .data$VC_count, y = .data$mean_val)
) +
  geom_errorbar(
    aes(ymin = mean_val - sd_val, ymax = mean_val + sd_val),
    width = 0, colour = col_nirs, linewidth = 0.4, alpha = 0.28
  ) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_nirs) +
  geom_point(shape = 15, size = 2.2, colour = col_nirs) +
  facet_wrap(~ feature, ncol = 1, scales = "free_y") +
  labs(
    title = "EX_DYN - per-cycle NIRS features (cohort mean +/- SD)",
    x     = "Repetition",
    y     = NULL
  ) +
  theme_amstrom_hgrid

print(nirs_ex_dyn_features_plot)
```

# 4. EX_STA trajectories

All runs truncated to the shared window count (`STA_N_MIN`), then cohort mean ΔHHb per 1 s
window. The first chunk just inspects per-run window counts before truncating.

```{r ex_sta_inspect_n_min}
# Inspect per-run EX_STA length (max window_index) before truncating to the shared minimum.
run_window_maxima <- nirs_ex_sta_pooled |>
  dplyr::group_by(.data$RUN_ID) |>
  dplyr::summarise(max_window_index = max(.data$window_index, na.rm = TRUE), .groups = "drop") |>
  dplyr::arrange(.data$max_window_index)

cat("Per-run EX_STA window_index maxima:\n")
print(run_window_maxima, n = Inf)
cat("\nSTA_N_MIN (GLOBAL_STA_N_MIN from cohort_config.R):", STA_N_MIN, "\n")
```

```{r ex_sta_harmonize}
# All runs, truncated at STA_N_MIN.
sta_label <- glue("All participants (N = {length(nirs_run_ids)})")

ex_sta <- nirs_ex_sta_pooled |>
  dplyr::filter(.data$window_index <= STA_N_MIN)

cat(sta_label, "| truncated at window_index <=", STA_N_MIN, "\n")
```

```{r ex_sta_plot, fig.width=6, fig.height=4}
# Cohort mean/SD ΔHHb per 1 s window, then the trajectory plot.
ex_sta_summary <- ex_sta |>
  dplyr::group_by(.data$window_index) |>
  dplyr::summarise(
    mean_hhb = mean(.data$hhb_from_seq_rest_umol, na.rm = TRUE),  # cohort mean ΔHHb (from seq rest)
    sd_hhb   = sd(.data$hhb_from_seq_rest_umol, na.rm = TRUE),    # between-run SD
    n        = dplyr::n(),                                       # rows pooled at this point
    .groups  = "drop"
  )

nirs_ex_sta_plot <- ggplot(ex_sta_summary, aes(x = .data$window_index, y = .data$mean_hhb)) +
  geom_errorbar(
    aes(ymin = mean_hhb - sd_hhb, ymax = mean_hhb + sd_hhb),
    width = 0, colour = col_nirs, linewidth = 0.4, alpha = 0.28
  ) +
  geom_line(linetype = "dotted", linewidth = 0.35, colour = col_nirs) +
  geom_point(shape = 15, size = 2.2, colour = col_nirs) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +  # zero = no change from baseline
  labs(
    title    = "EX_STA - delta HHb from seq. rest (cohort mean +/- SD)",
    subtitle = sta_label,
    x        = "Window index (1 s)",
    y        = "delta HHb from seq. rest (umol/L)"
  ) +
  theme_amstrom_hgrid

print(nirs_ex_sta_plot)
```

# 5. Run-level summary

One value per run (linear slope over time, and last-minus-first delta), shown as individual
points plus the cohort mean. Slope and last-minus-first are drawn in separate plots because
their y-scales are incompatible (µmol/L per step vs µmol/L).

```{r run_summary_plot, fig.width=9, fig.height=6}
# Split the per-run summary metrics into slope vs last-minus-first (different y-scales), then
# draw each as jittered per-run points with the cohort mean marked.
summary_plot_df <- nirs_summary_pooled |>
  dplyr::filter(is.finite(.data$summary_value)) |>
  dplyr::mutate(
    panel_label = paste0(.data$summary_variable, "\n[", .data$source_sequence, "]"),
    type_label  = dplyr::case_match(
      .data$summary_type,
      c("VC_count_linear_slope", "window_index_linear_slope") ~ "Linear slope",
      "last_minus_first"                                       ~ "Last - first",
      .default = .data$summary_type
    )
  )

# Slope and last_minus_first have incompatible y-axis scales (µmol/L/rep vs µmol/L),
# so they are shown in separate plots rather than the same facet panel.

slope_df <- summary_plot_df |> dplyr::filter(.data$type_label == "Linear slope")
delta_df <- summary_plot_df |> dplyr::filter(.data$type_label == "Last - first")

nirs_slope_plot <- ggplot(slope_df, aes(x = "", y = .data$summary_value)) +
  geom_jitter(width = 0.15, height = 0, shape = 1, size = 2, alpha = 0.75, colour = col_nirs) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3.5, colour = col_nirs) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +  # zero = no change from baseline
  facet_wrap(~ panel_label, scales = "free_y", ncol = 3) +
  labs(
    title = "NIRS - Linear slope per run (individual runs + cohort mean)",
    x     = NULL,
    y     = "µmol/L per repetition (or per window)"
  ) +
  theme_amstrom_hgrid +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

# Last - first: delta HHb from the first to the last window of each sequence.
# Analogous to Song et al. 2025 (1273) phase-1 to phase-5 total excursion — total
# net deoxygenation change across the exercise bout.
nirs_delta_plot <- ggplot(delta_df, aes(x = "", y = .data$summary_value)) +
  geom_jitter(width = 0.15, height = 0, shape = 1, size = 2, alpha = 0.75, colour = col_expected) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3.5, colour = col_expected) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey72", linewidth = 0.4) +  # zero = no change from baseline
  facet_wrap(~ panel_label, scales = "free_y", ncol = 3) +
  labs(
    title    = "NIRS - Last - first delta per run (individual runs + cohort mean)",
    subtitle = "Analogous to Song et al. 2025 phase-1 to phase-5 total excursion",
    x        = NULL,
    y        = "delta HHb (µmol/L)"
  ) +
  theme_amstrom_hgrid +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

print(nirs_slope_plot)
print(nirs_delta_plot)
```

```{r run_summary_table}
# Per-run wide summary: EX_DYN/EX_STA mean ΔHHb + slope + last-minus-first, one row per run.
per_run_dyn_mean <- nirs_ex_dyn_pooled |>
  dplyr::group_by(.data$RUN_ID) |>
  dplyr::summarise(
    exdyn_mean = round(mean(.data$hhb_from_seq_rest_umol, na.rm = TRUE), 1),
    .groups = "drop"
  )

per_run_sta_mean <- nirs_ex_sta_pooled |>
  dplyr::filter(.data$window_index <= STA_N_MIN) |>
  dplyr::group_by(.data$RUN_ID) |>
  dplyr::summarise(
    exsta_mean = round(mean(.data$hhb_from_seq_rest_umol, na.rm = TRUE), 1),
    .groups = "drop"
  )

summary_wide <- nirs_summary_pooled |>
  dplyr::filter(
    .data$summary_variable == "hhb_from_seq_rest_umol",
    .data$analysis_block   %in% c("nirs_ex_dyn", "nirs_ex_sta")
  ) |>
  dplyr::mutate(col_name = dplyr::case_when(
    analysis_block == "nirs_ex_dyn" & summary_type == "VC_count_linear_slope"    ~ "exdyn_slope",
    analysis_block == "nirs_ex_dyn" & summary_type == "last_minus_first"          ~ "exdyn_lmf",
    analysis_block == "nirs_ex_sta" & summary_type == "window_index_linear_slope" ~ "exsta_slope",
    analysis_block == "nirs_ex_sta" & summary_type == "last_minus_first"          ~ "exsta_lmf"
  )) |>
  dplyr::filter(!is.na(.data$col_name)) |>
  tidyr::pivot_wider(id_cols = "RUN_ID", names_from = "col_name", values_from = "summary_value") |>
  dplyr::mutate(dplyr::across(-RUN_ID, \(x) round(x, 2)))

per_run_dyn_mean |>
  dplyr::left_join(per_run_sta_mean, by = "RUN_ID") |>
  dplyr::left_join(summary_wide,     by = "RUN_ID") |>
  dplyr::arrange(.data$exdyn_mean) |>
  knitr::kable(
    col.names = c("Run", "EX_DYN mean", "EX_STA mean",
                  "EX_DYN slope", "EX_DYN last-first",
                  "EX_STA slope", "EX_STA last-first"),
    caption = "Per-run delta HHb from seq. rest (µmol/L). Sorted by EX_DYN mean. Slope = µmol/L per rep or window."
  )
```

# 6. Summary tables

Set `PRINT_SUMMARY_TABLES <- TRUE` in §1 controls to render these.

## Reference checkpoints

```{r summary_ref, eval = PRINT_SUMMARY_TABLES}
# Summary table: reference ΔHHb per state (cohort mean +/- SD).
ref_summary |>
  dplyr::mutate(dplyr::across(c(mean_hhb, sd_hhb), ~round(.x, 2))) |>
  knitr::kable(
    col.names = c("State", "Mean HHb (µmol/L)", "SD", "N runs"),
    caption   = "Reference sequences - delta HHb from per-sequence baseline."
  )
```

## EX_DYN trajectory (5 representative reps)

```{r summary_ex_dyn, eval = PRINT_SUMMARY_TABLES}
# Summary table: ΔHHb at 5 evenly-spaced EX_DYN reps.
idx_dyn <- round(seq(1, nrow(ex_dyn_summary), length.out = min(5L, nrow(ex_dyn_summary))))
ex_dyn_summary[idx_dyn, ] |>
  dplyr::mutate(dplyr::across(c(mean_hhb, sd_hhb), ~round(.x, 2))) |>
  knitr::kable(
    col.names = c("Rep", "Mean HHb (µmol/L)", "SD", "N"),
    caption   = "EX_DYN delta HHb - 5 evenly-spaced reps (cohort mean +/- SD)."
  )
```

## EX_STA trajectory (5 representative windows)

```{r summary_ex_sta, eval = PRINT_SUMMARY_TABLES}
# Summary table: ΔHHb at 5 evenly-spaced EX_STA windows.
idx_sta <- round(seq(1, nrow(ex_sta_summary), length.out = min(5L, nrow(ex_sta_summary))))
ex_sta_summary[idx_sta, ] |>
  dplyr::mutate(dplyr::across(c(mean_hhb, sd_hhb), ~round(.x, 2))) |>
  knitr::kable(
    col.names = c("Window", "Mean HHb (µmol/L)", "SD", "N"),
    caption   = paste0("EX_STA delta HHb - 5 evenly-spaced windows (all runs, truncated at ", STA_N_MIN, ").")
  )
```

## Run-level summary

```{r summary_runs, eval = PRINT_SUMMARY_TABLES}
# Summary table: every per-run NIRS metric (slopes + last-minus-first deltas), long format.
nirs_summary_pooled |>
  dplyr::filter(is.finite(.data$summary_value)) |>
  dplyr::mutate(summary_value = round(.data$summary_value, 4)) |>
  dplyr::select(RUN_ID, source_sequence, summary_variable, summary_type, summary_value) |>
  dplyr::arrange(.data$source_sequence, .data$summary_variable, .data$RUN_ID) |>
  knitr::kable(
    col.names = c("Run", "Sequence", "Variable", "Type", "Value"),
    caption   = "Per-run NIRS summary metrics (slopes and last-minus-first deltas)."
  )
``` 